### 2024-11-21 paper/paper.qmd
Model, data, paper, llm usage updated

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: ."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(rstanarm)
library(modelsummary)
library(knitr)
library(ggplot2)

logistic_model <- readRDS(here::here("models/logistic_model_last.rds"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@wordbankr

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-data). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-data
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

## Model Selection

To investigate the relationship between children’s vocabulary acquisition and their demographic and linguistic characteristics, we constructed a logistic regression model. By examining key demographic and linguistic predictors, we aim to identify how characteristics like age, norming status, and word categories influence vocabulary development. The dependent variable, high_vocabulary, is a binary outcome indicating whether a child’s average production and comprehension score (denoted as prod_comp_mean) exceeds 350. This threshold was chosen to distinguish children with relatively advanced vocabulary levels. More background details and diagnostics are included in [Appendix -@sec-model-details].

## Logistic Regression Model Overview

- High Vocabulary: The outcome variable, high_vocabulary, is a binary indicator that takes the value of 1 if the average production and comprehension score (prod_comp_mean) exceeds 350 and 0 otherwise. This threshold was selected to represent children with relatively advanced vocabulary skills, determined through exploratory data analysis.
- Scaled Age (age_scaled): This continuous variable represents the child’s age, standardized to ensure the model coefficients reflect changes per standard deviation in age. Standardization improves numerical stability and aids interpretability.
- Norming Status (is_norming): A binary indicator denoting whether a child is part of the norming dataset (TRUE) or not (FALSE). This variable accounts for potential differences in data collection or assessment protocols.
- Broad Category (broad_category): A categorical variable grouping words into four broad linguistic categories: nouns, verbs, adjectives, and function_words. The reference category is function_words.

The model takes the form:
\begin{align}
\log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \cdot \text{age\_scaled}_i + \beta_2 \cdot \text{is\_normingTRUE}_i \\
&\quad + \beta_3 \cdot \text{broad\_categoryNouns}_i \\
&\quad + \beta_4 \cdot \text{broad\_categoryFunction\_words}_i \\
&\quad + \beta_5 \cdot \text{broad\_categoryVerbs}_i
\end{align}

Where:

- $p_i$ represents the probability that person i has a high vocabulary
- $\beta_0$ is the intercept, capturing the baseline log-odds when all predictors are at their reference or mean levels
- $\beta_1$: Effect of age (standardized)
- $\beta_2$: The effect of whether the individual belongs to the norming group
- $\beta_3$, $\beta_4$, $\beta_5$: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).

## Model Assumpations
- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome and the independent variables. Standardizing age ensures this assumption is more likely to hold.
- Independent Observations: The data points are assumed to be independent, which is valid given the dataset structure.
- Categorical Variable Encoding: The broad category variable was encoded using sum contrasts to test the deviation of each category from the overall mean effect.

## Interpretation of Coefficients
The logistic regression coefficient ($\beta$) represents the change in the log-odds of the dependent variable (high vocabulary) for a one-unit change in the predictor variable, holding all other variables constant.

- Intercept ($\beta_0$):
•	Represents the log-odds of high vocabulary when all predictors are at their reference or mean levels.
•	For example, if $\beta_0$ > 0, the baseline odds of high vocabulary are greater than 50%.

- Scaled Age ($\beta_1$):
Interpretation: For each one standard deviation increase in age, the log-odds of high vocabulary increase by $\beta_1$.
Example: If $\beta_1 = 0.5$, then $\exp(0.5) \approx 1.65$, meaning the odds increase by 65% for every one standard deviation increase in age.

- Norming Status ($\beta_2$): If a child belongs to the norming group, the log-odds of high vocabulary increase by $\beta_2$ compared to non-norming children. If $\beta_2 = 0.1$, then $\exp(0.1) \approx 1.11$, meaning being in the norming group increases the odds of high vocabulary by 11%.

- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$): The coefficients for broad_category represent the difference in log-odds compared to the reference category (“adjectives”). 
  Interpretation:
   - $\beta_3$ (Function Words): A positive $\beta_3$ indicates higher odds of high vocabulary for function words compared to adjectives
   - $\beta_4$ (Nouns): A negative $\beta_4$ indicates lower odds of high vocabulary for nouns compared to adjectives
   - $\beta_5$ (Verbs): If $\beta_5 = 0$, there is no difference in odds between verbs and adjectives.

General Example for Broad Categories:

If $\beta_3 = 0.01$, $\exp(0.01) \approx 1.01$, meaning function words increase the odds by 1% compared to adjectives.

## Model Justification

The model was trained on 80% of the dataset, with the remaining 20% reserved for testing. Model performance was assessed using confusion matrices and overall accuracy. Further evaluation included the analysis of residual deviance, AIC, and interpretation of individual coefficients.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(list("Logistic model" = logistic_model), statistic = "std.error", fmt = 2)

```

# Results

Our results are summarized in .

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))

# Preprocess test data
analysis_data_test <- analysis_data_test %>%
  mutate(
    # Create new variables for prediction
    prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
    high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0),  # Binary target variable
    age_scaled = scale(age)  # Standardize age
  ) %>%
  drop_na(prod_comp_mean)  # Drop rows with NA in prod_comp_mean
analysis_data_test$age_scaled <- as.numeric(analysis_data_test$age_scaled)

# Predict probabilities for test data
predicted_probabilities <- predict(logistic_model, newdata = analysis_data_test, type = "response")

# Predict classes based on probabilities
predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)

# Create a confusion matrix
confusion_matrix <- table(
  Actual = analysis_data_test$high_vocabulary,
  Predicted = predicted_classes
)
print(confusion_matrix)

accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
print(paste("Accuracy:", accuracy))
```

```{r}
#| echo: false
#| eval: true
#| label: fig-modelresult
#| fig-cap: How the probability of high vocabulary varies with age by aggregating the predictions for each age group
#| warning: false

# Add predicted probabilities to the test dataset
analysis_data_test <- analysis_data_test %>%
  mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))

# Ensure the broad_category column is included in the test data
analysis_data_test <- analysis_data_test %>%
  mutate(broad_category = case_when(
    category %in% c("action_words", "helping_verbs") ~ "verbs",
    category %in% c("connecting_words", "question_words", "quantifiers", "pronouns") ~ "function_words",
    category %in% c("animals", "body_parts", "clothing", "food_drink", "furniture_rooms", 
                    "games_routines", "household", "locations", "outside", "people", 
                    "places", "sounds", "toys", "vehicles") ~ "nouns",
    category %in% c("descriptive_words", "time_words") ~ "adjectives",
    TRUE ~ NA_character_
  ))

# Aggregate mean predicted probabilities by age and broad_category
age_broad_category_predictions <- analysis_data_test %>%
  group_by(age, broad_category) %>%
  summarize(mean_predicted_prob = mean(predicted_prob, na.rm = TRUE), .groups = "drop")

broad_category_summary <- age_broad_category_predictions %>%
  group_by(broad_category) %>%
  summarize(overall_mean_prob = mean(mean_predicted_prob, na.rm = TRUE))

```

The density plot highlights distinct patterns in predicted probabilities, with nouns and verbs showing higher peaks, indicating a stronger likelihood of high vocabulary acquisition in these categories. This suggests that broad categories contribute differently to vocabulary development, with nouns and verbs potentially playing a more significant role.

```{r}
#| echo: false
#| eval: true
#| label: fig-prediction
#| fig-cap: Distribution of predicted probabilities for high vocabulary acquisition across broad broad categories (adjectives, function words, nouns, and verbs). 
#| warning: false

ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Distribution of Predicted Probabilities by Broad Category",
    x = "Predicted Probability",
    y = "Density",
    fill = "Broad Category"
  ) +
  theme_minimal()
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-22 paper/paper.qmd
Update cleaned data and paper qmd.

@@
 ---
-title: "My title"
+title: "Investigation Vocabulary Learning Pattern"
 subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: ."
+author: Yongqi Liu
+thanks: "Code and data are available at: https://github.com/Cassieliu77/Vocabulary_Learning_Pattern.git"
 date: today
@@
 library(knitr)
+library(dplyr)
 library(ggplot2)
+library(scales)
 
+analysis_data <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
 logistic_model <- readRDS(here::here("models/logistic_model_last.rds"))
@@
 
-@wordbankr
+The data used in this study comes from @wordbankr, and all analysis are conducted in @citeR
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+This study uses R packages [@citeR] to clean and analyze the dataset, including libraries 
+from tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], arrow [@arrow]. 
+We use the statistical programming language R [@citeR].... Our data [@wordbankr]
 
-Overview text
+After cleaning the data, which included grouping and removing missing values, the following analysis focuses on category, age, comprehension, production, is_norming, broad_category columns in the analysis dataset.
+@tbl-data shows the overview of the dataset.
+
+```{r}
+#| label: tbl-data
+#| tbl-cap: Summary Table for the Word Bank Dataset
+#| echo: false
+
+summary_table <- analysis_data %>%
+  select(data_id, language, age, is_norming, broad_category, production, high_vocabulary) %>%
+  rename(
+    Data_ID = data_id,
+    Language = language,
+    Age = age,
+    Is_Norming = is_norming,
+    Broad_Category = broad_category,
+    Production = production,
+    High_Vocabulary = high_vocabulary)
+
+kable(head(summary_table, 10))
+```
 
@@
 
-## Outcome variables
+## Outcome Variable
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+### High Vocabulary Score
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The outcome variable in this study, High Vocabulary Score, is a binary indicator designed to identify individuals with advanced vocabulary proficiency. This variable is derived from two key measures:
+
+1.
+Comprehension: This variable represents the ability to understand words and phrases, reflecting the receptive language skills of individuals.
+Comprehension scores are numerical and vary across the dataset.
+
+2.
+Production: This variable captures the ability to produce words, reflecting expressive language skills.
+Like comprehension, production scores are numerical and provide the standard into verbal articulation capabilities.
+
+3.
+The High Vocabulary Score is calculated using the average of comprehension and production scores for each individual.
+This average is represented as: $\text{prod\_comp\_mean} = \frac{\text{Comprehension} + \text{Production}}{2}$
+
+To classify individuals, a threshold value of 350 is applied to `prod_comp_mean`:
+- Individuals with `prod_comp_mean > 350` are classified as having a high vocabulary score (outcome = 1).
+- Those with `prod_comp_mean <= 350` are classified as not having a high vocabulary score (outcome = 0).
+
+This approach ensures that both receptive (comprehension) and expressive (production) skills are considered in defining advanced vocabulary.
+The threshold of 350 was chosen based on exploratory analysis of the dataset, reflecting a meaningful distinction between individuals with high and low vocabulary abilities.
+The High Vocabulary Score serves as the dependent variable in the following data analysis part.
+Its binary nature makes it suitable for modeling with a binomial family distribution, allowing for the estimation of factors that influence advanced vocabulary acquisition.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-outcome
+#| fig-cap: Distribution of the outcome variable, showing the counts of children classified as having “High Vocabulary” and “Not High Vocabulary” based on their comprehension and production scores.
+The bar plot illustrates the balance between the two categories in the dataset, which is important for modeling purposes.
 #| echo: false
+#| warning: false
+#| message: false
+outcome_counts <- analysis_data %>%
+  count(high_vocabulary) %>%
+  mutate(
+    VocabularyCategory = ifelse(high_vocabulary == 1, "High Vocabulary", "Not High Vocabulary"))
+
+ggplot(outcome_counts, aes(x = VocabularyCategory, y = n, fill = VocabularyCategory)) +
+  geom_bar(stat = "identity", width = 0.7, show.legend = FALSE) +
+  labs(
+    title = "Distribution of High Vocabulary Levels",
+    x = "Vocabulary Level",
+    y = "Count") +
+  scale_fill_manual(values = c("High Vocabulary" = "blue", "Not High Vocabulary" = "red")) +
+  scale_y_continuous(labels = scales::comma) +
+  theme_minimal(base_size = 12) +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title.x = element_text(size = 14, margin = margin(t = 10)),
+    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
+    axis.text = element_text(size = 12),)
 
@@
 
-Talk more about it.
-
-And also planes (@fig-data).
+## Predictor variables
+### Age
+And also planes (@fig-age).
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
@@
 ```{r}
-#| label: fig-data
-#| fig-cap: Relationship between wing length and width
 #| echo: false
+#| label: fig-age
+#| fig-cap: Age Distribution
 #| warning: false
@@
 
-analysis_data <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
+ggplot(analysis_data, aes(x = age)) +
+  geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.8) +
+  labs(title = "Age Distribution", x = "Age (months)", y = "Count") +
+  theme_minimal()
 
 ```
+### Broad Category
 
-Talk way more about it.
+```{r}
+#| echo: false
+#| label: fig-category
+#| fig-cap:  Distribution of Word Categories.
+This bar plot illustrates the distribution of items across broad linguistic categories in the dataset.
+“Objects” dominate the dataset, followed by “Verbs” and “Living Things,” indicating that the dataset is rich in concrete nouns and action-related words, with smaller proportions of sensory words and adjectives.
+#| warning: false
+#| message: false
 
-## Predictor variables
+ggplot(analysis_data, aes(x = broad_category)) +
+  geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
+  labs(title = "Category Distribution", x = "Broad Category", y = "Count") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+### Is Norming or Not
+```{r}
+#| echo: false
+#| label: fig-norming
+#| fig-cap: Norming Group Distribution.
+This plot shows the distribution of children in the dataset categorized by their norming status.
+The majority of the observations are from non-norming children, with only a small fraction representing norming children.
+#| warning: false
+#| message: false
+ggplot(analysis_data, aes(x = is_norming)) +
+  geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
+  labs(title = "Norming Distribution", x = "Norming or not", y = "Count") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_y_continuous(labels = scales::comma)
+```
 
@@
 This threshold was chosen to distinguish children with relatively advanced vocabulary levels.
-More background details and diagnostics are included in [Appendix -@sec-model-details].
+More background details and diagnostics are included in [Appendix-@sec-model-details].
 
@@
 The model takes the form:
+
 \begin{align}
 \log \left( \frac{p_i}{1 - p_i} \right) &= \beta_0 + \beta_1 \cdot \text{age\_scaled}_i + \beta_2 \cdot \text{is\_normingTRUE}_i \\
-&\quad + \beta_3 \cdot \text{broad\_categoryNouns}_i \\
-&\quad + \beta_4 \cdot \text{broad\_categoryFunction\_words}_i \\
-&\quad + \beta_5 \cdot \text{broad\_categoryVerbs}_i
+&\quad + \beta_3 \cdot \text{broad\_categoryAdjectives}_i \\
+&\quad + \beta_4 \cdot \text{broad\_categoryFunction\_Words}_i \\
+&\quad + \beta_5 \cdot \text{broad\_categoryLiving\_Things}_i \\
+&\quad + \beta_6 \cdot \text{broad\_categoryObjects}_i \\
+&\quad + \beta_7 \cdot \text{broad\_categoryPlaces}_i \\
+&\quad + \beta_8 \cdot \text{broad\_categorySensory\_Words}_i \\
+&\quad + \beta_9 \cdot \text{broad\_categoryVerbs}_i
 \end{align}
@@
 Where:
-
 - $p_i$ represents the probability that person i has a high vocabulary
@@
 - $\beta_2$: The effect of whether the individual belongs to the norming group
-- $\beta_3$, $\beta_4$, $\beta_5$: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).
+- $\beta_3$, $\beta_4$, $\beta_5$, etc.: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).
 
-## Model Assumpations
-- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome and the independent variables. Standardizing age ensures this assumption is more likely to hold.
-- Independent Observations: The data points are assumed to be independent, which is valid given the dataset structure.
-- Categorical Variable Encoding: The broad category variable was encoded using sum contrasts to test the deviation of each category from the overall mean effect.
+## Model Assumptions
+- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary) and the independent variables.
+For example, the standardized age variable (age_scaled) assumes that for every one standard deviation increase in age, the log-odds of achieving a high vocabulary score increase by a constant amount.
+Standardizing age ensures that the variable is centered and scaled, making it easier to meet this linearity assumption and interpret its effect across the dataset.
+- Independent Observations: The model assumes that all data points are independent.
+This assumption holds because each observation represents data from a unique child, with no repeated measurements for the same individual.
+For instance, there are no longitudinal observations or nested data structures (e.g., children grouped by classrooms or schools) that could violate independence.
+If dependence were present, a more complex model like a mixed-effects logistic regression would be necessary.
+- Categorical Variable Encoding: The broad_category variable, which includes categories such as “Adjectives,” “Verbs,” and “Living Things,” was encoded using sum contrasts.
+This approach ensures that the coefficients for each category represent the deviation of that category’s effect from the overall mean effect across all categories.
+For example, the coefficient for “Verbs” indicates how the log-odds of achieving a high vocabulary differ for “Verbs” compared to the average effect of all other categories.
+Sum contrasts are particularly useful for understanding relative effects and ensure that the intercept reflects the overall mean effect when all predictors are at their reference or average levels.
 
@@
 
-- Intercept ($\beta_0$):
-•	Represents the log-odds of high vocabulary when all predictors are at their reference or mean levels.
-•	For example, if $\beta_0$ > 0, the baseline odds of high vocabulary are greater than 50%.
-
-- Scaled Age ($\beta_1$):
-Interpretation: For each one standard deviation increase in age, the log-odds of high vocabulary increase by $\beta_1$.
-Example: If $\beta_1 = 0.5$, then $\exp(0.5) \approx 1.65$, meaning the odds increase by 65% for every one standard deviation increase in age.
-
+- Intercept ($\beta_0$): Represents the log-odds of high vocabulary when all predictors are at their reference or mean levels.
+If $\beta_0$ > 0, the baseline odds of high vocabulary are greater than 50%.
+- Scaled Age ($\beta_1$): For each one standard deviation increase in age, the log-odds of high vocabulary increase by $\beta_1$.
+If $\beta_1 = 0.5$, then $\exp(0.5) \approx 1.65$, meaning the odds increase by 65% for every one standard deviation increase in age.
 - Norming Status ($\beta_2$): If a child belongs to the norming group, the log-odds of high vocabulary increase by $\beta_2$ compared to non-norming children.
@@
 
-- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$): The coefficients for broad_category represent the difference in log-odds compared to the reference category (“adjectives”). 
+- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$, etc.): The coefficients for `broad_category` represent the difference in log-odds compared to the reference category (“Adjectives”).
+
 Interpretation:  
-   - $\beta_3$ (Function Words): A positive $\beta_3$ indicates higher odds of high vocabulary for function words compared to adjectives
-   - $\beta_4$ (Nouns): A negative $\beta_4$ indicates lower odds of high vocabulary for nouns compared to adjectives
-   - $\beta_5$ (Verbs): If $\beta_5 = 0$, there is no difference in odds between verbs and adjectives.
+  - $\beta_3$ (Function Words): A positive $\beta_3$ indicates higher odds of having a high vocabulary for function words compared to adjectives.
+For instance, if $\beta_3 = 0.002$, then $\exp(0.002) \approx 1.002$, meaning the odds of having a high vocabulary for function words are 0.2% higher than for adjectives.
+  - $\beta_4$ (Living Things): A negative $\beta_4$ indicates lower odds of having a high vocabulary for living things compared to adjectives.
+For example, if $\beta_4 = -0.007$, then $\exp(-0.007) \approx 0.993$, meaning the odds of having a high vocabulary for living things are 0.7% lower than for adjectives.
+  - $\beta_5$ (Objects): If $\beta_5 = -0.008$, then $\exp(-0.008) \approx 0.992$, indicating that objects are associated with 0.8% lower odds of having a high vocabulary compared to adjectives.  
+  - $\beta_6$ (Verbs): If $\beta_6 = -0.005$, then $\exp(-0.005) \approx 0.995$, meaning verbs are associated with 0.5% lower odds of having a high vocabulary compared to adjectives.
 
 General Example for Broad Categories:  
-
-If $\beta_3 = 0.01$, $\exp(0.01) \approx 1.01$, meaning function words increase the odds by 1% compared to adjectives.
+  If a coefficient $\beta_k = 0.01$, $\exp(0.01) \approx 1.01$, meaning the corresponding category increases the odds of having a high vocabulary by 1% compared to the reference category (Adjectives).
+Conversely, if $\beta_k = -0.01$, $\exp(-0.01) \approx 0.99$, indicating a 1% decrease in odds compared to the reference category.
 
@@
 
+# Results
+
+## Average Production by Broad Category and Age
+@fig-average shows..
 ```{r}
@@
 #| eval: true
-#| label: tbl-modelsummary
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: fig-average
+#| fig-cap: xx 
 #| warning: false
 
-modelsummary(list("Logistic model" = logistic_model), statistic = "std.error", fmt = 2)
-
+analysis_data %>%
+  group_by(broad_category, age) %>%
+  summarize(avg_production = mean(production, na.rm = TRUE), .groups = "drop") %>%
+  ggplot(aes(x = age, y = avg_production, fill = broad_category)) +
+  geom_col(position = "stack") +  # Use stack for stacked bars
+  labs(
+    title = "Average Production by Broad Category and Age",
+    x = "Age (months)",
+    y = "Average Production Score",
+    fill = "Broad Category"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    legend.title = element_text(size = 11))
 ```
 
-# Results
-
-Our results are summarized in .
-
+## Prediciton for the Probability of High Vocabulary Level
 ```{r}
@@
 analysis_data_test <- analysis_data_test %>%
-  mutate(
-    # Create new variables for prediction
-    prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
+  mutate(prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
     high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0),  # Binary target variable
@@
 #| eval: true
-#| label: fig-modelresult
+#| label: fig-goodgood
 #| fig-cap: How the probability of high vocabulary varies with age by aggregating the predictions for each age group
@@
   mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))
-
-# Ensure the broad_category column is included in the test data
-analysis_data_test <- analysis_data_test %>%
-  mutate(broad_category = case_when(
-    category %in% c("action_words", "helping_verbs") ~ "verbs",
-    category %in% c("connecting_words", "question_words", "quantifiers", "pronouns") ~ "function_words",
-    category %in% c("animals", "body_parts", "clothing", "food_drink", "furniture_rooms", 
-                    "games_routines", "household", "locations", "outside", "people", 
-                    "places", "sounds", "toys", "vehicles") ~ "nouns",
-    category %in% c("descriptive_words", "time_words") ~ "adjectives",
-    TRUE ~ NA_character_
-  ))
 
@@
 
+### Distribution of Predicted Probabilities by Broad Category
+
+@fig-prediction illustrates the distribution of predicted probabilities for high vocabulary levels across various broad categories.
+The density curves show distinct peaks and variability, reflecting the differences in how well each category predicts high vocabulary.
+Notably, “Sensory Words” and “Objects” exhibit higher density in the middle range of predicted probabilities, suggesting moderate association with high vocabulary.
+In contrast, categories like “Activities” and “Function Words” have wider, flatter distributions, indicating greater uncertainty or diversity in prediction.
+This variability highlights the nuanced role of word categories in predicting high vocabulary acquisition.
+Further analysis could explore why certain categories contribute more consistently to predictions than others.
 The density plot highlights distinct patterns in predicted probabilities, with nouns and verbs showing higher peaks, indicating a stronger likelihood of high vocabulary acquisition in these categories.
@@
 #| label: fig-prediction
-#| fig-cap: Distribution of predicted probabilities for high vocabulary acquisition across broad broad categories (adjectives, function words, nouns, and verbs). 
+#| fig-cap: This density plot illustrates the predicted probabilities of having a high vocabulary across different broad lexical categories.
+Each curve represents the density of predicted probabilities within a category, showcasing the variation in predicted outcomes for categories such as Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs.
+The plot highlights overlapping patterns and areas of divergence in vocabulary acquisition likelihood across different types of words.
 #| warning: false
@@
 ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
-  geom_density(alpha = 0.5) +
-  labs(
-    title = "Distribution of Predicted Probabilities by Broad Category",
+  geom_density(alpha = 0.7, color = "black") +
+  labs(title = "Distribution of Predicted Probabilities by Broad Category",
     x = "Predicted Probability",
     y = "Density",
-    fill = "Broad Category"
-  ) +
-  theme_minimal()
+    fill = "Broad Category") +
+  theme_minimal(base_size = 14) +
+  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
+    legend.position = "bottom",
+    legend.title = element_text(face = "bold"),
+    panel.grid.major = element_line(color = "gray85"),
+    panel.grid.minor = element_blank()
+  )
 ```
@@
 
+## Data Sheet
+
 # Model details {#sec-model-details}
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+## Model Summary
 
@@
 #| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+modelsummary(logistic_model, statistic = "std.error", fmt = 2)
 
@@
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
 
@@
 
+# Acknowledgements
+
+Thanks to Open AI and ChatGPT 4.0 is used to write the paper.
+This project utilized @citeR and a variety of R packages that provided essential tools and functionality.
+We thank the @tidyverse, @ggplot2 and @knitr teams for their extensive suite of tools for data manipulation and visualization, which were instrumental in data cleaning and graphing.
+Finally, @arrow was vital for efficient data storage and handling with Parquet files.
+We are grateful to the developers and maintainers of these packages for their contributions to open-source software.
+
 # References

### 2024-11-22 scripts/03-clean_data.R

@@ -20,7 +20,7 @@ library(arrow)
 raw_data <- read_parquet("data/01-raw_data/raw_data.parquet")
 
 # Select the needed columns
-cleaned_data <- raw_data %>% select(data_id, item_id, language, form, item_kind, 
+cleaned_data <- raw_data %>% select(data_id, language, form, item_kind, 
                                  category, uni_lemma, date_of_test,
                                  age, comprehension, production, 
                                  is_norming, child_id)
@@ -37,13 +37,23 @@ cleaned_data <- cleaned_data %>%
 # Define a mapping of categories to broader groups
 cleaned_data <- cleaned_data %>%
   mutate(broad_category = case_when(
-    category %in% c("action_words", "helping_verbs") ~ "verbs",
-    category %in% c("connecting_words", "question_words", "quantifiers", "pronouns") ~ "function_words",
-    category %in% c("animals", "body_parts", "clothing", "food_drink", "furniture_rooms", 
-                    "games_routines", "household", "locations", "outside", "people", 
-                    "places", "sounds", "toys", "vehicles") ~ "nouns",
-    category %in% c("descriptive_words", "time_words") ~ "adjectives",
-    TRUE ~ "unknown"))
+    category %in% c("action_words", "helping_verbs") ~ "Verbs",
+    category %in% c("connecting_words", "question_words", "quantifiers", "pronouns") ~ "Function Words",
+    category %in% c("animals", "body_parts", "people") ~ "Living Things",
+    category %in% c("clothing", "food_drink", "furniture_rooms", "household", "toys", "vehicles") ~ "Objects",
+    category %in% c("games_routines") ~ "Activities",
+    category %in% c("sounds") ~ "Sensory Words",
+    category %in% c("descriptive_words", "time_words") ~ "Adjectives",
+    category %in% c("locations", "places", "outside") ~ "Places",
+    TRUE ~ "unknown"
+  ))
+
+# Combine production and comprehension into a new variable
+cleaned_data <- cleaned_data %>%
+  mutate(prod_comp_mean = (production + comprehension)/2,  # Average of production and comprehension
+         high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0)  # Binary target variable
+  ) %>%
+  drop_na(prod_comp_mean) 
 
 # Omit rows with NA values
 cleaned_data <- cleaned_data %>% 

### 2024-11-22 scripts/05-exploratory_data_analysis.R

@@ -85,6 +85,25 @@ ggplot(analysis_data, aes(x = is_norming)) +
   labs(title = "Proportion of Norming Data", x = "Is Norming?", y = "Count") +
   theme_minimal()
 
+# B8. Plot the Distribution of 'High Vocabulary'
+outcome_counts <- analysis_data %>%
+  count(high_vocabulary) %>%
+  mutate(
+    VocabularyCategory = ifelse(high_vocabulary == 1, "High Vocabulary", "Not High Vocabulary")
+  )
+
+ggplot(outcome_counts, aes(x = VocabularyCategory, y = n, fill = VocabularyCategory)) +
+  geom_bar(stat = "identity", width = 0.6, show.legend = FALSE) +
+  labs(
+    title = "Distribution of Outcome Variable: High Vocabulary",
+    x = "Vocabulary Category",
+    y = "Count"
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 12),
+    plot.title = element_text(hjust = 0.5)
+  )
 #-------------------------------------------------------------------------------
 ### Part C: Visualize Relation between Variables ###
 # C1. Vocabulary size by age:Plot Production by Age to see how children’s vocabulary develops over time

### 2024-11-22 scripts/06-model_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Model for the vocabulary size prediction
+# Purpose: Build the logistic regression model for the vocabulary size prediction
 # Author: Yongqi Liu
 # Date: 15 Nov 2024 
 # Contact: cassieliu.liu@mail.utoronto.ca

### 2024-11-22 paper/paper.qmd
Update paper draft and llm usage

@@
 ---
-title: "Investigation Vocabulary Learning Pattern"
+title: "English Vocabulary Learning Pattern"
 subtitle: "My subtitle if needed"
@@
 library(arrow)
-library(rstanarm)
 library(modelsummary)
@@
 library(scales)
+library(wordbankr)
 
@@
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data descirbe the dataset in detail, and shows the distribution of variables.
 
-The data used in this study comes from @wordbankr, and all analysis are conducted in @citeR
+The data used in this study comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR
 
@@
 
-This study uses R packages [@citeR] to clean and analyze the dataset, including libraries 
-from tidyverse [@tidyverse], ggplot2 [@ggplot2], knitr [@knitr], arrow [@arrow]. 
-We use the statistical programming language R [@citeR].... Our data [@wordbankr]
-
-After cleaning the data, which included grouping and removing missing values, the following analysis focuses on category, age, comprehension, production, is_norming, broad_category columns in the analysis dataset.
-@tbl-data shows the overview of the dataset.
+The original dataset was obtained from @wordbankr.
+After undergoing a thorough cleaning process—including grouping related items and removing missing values—the analysis focuses on the key variables: category, age, comprehension, production, is_norming, and broad_category.
+These variables form the foundation of the analysis dataset.
+An overview of the cleaned dataset is presented in @tbl-data.
 
@@
 #| label: tbl-data
-#| tbl-cap: Summary Table for the Word Bank Dataset
+#| tbl-cap: Cleaned Word Bank Dataset
 #| echo: false
@@
 summary_table <- analysis_data %>%
-  select(data_id, language, age, is_norming, broad_category, production, high_vocabulary) %>%
+  select(language, age, is_norming, broad_category, production, high_vocabulary) %>%
   rename(
-    Data_ID = data_id,
     Language = language,
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+### Data Collection and Quality Control
+
+The objective of measurement in this study is to translate raw parental reports into reliable indicators of vocabulary acquisition patterns in children.
+The data is derived from the MacArthur-Bates Communicative Development Inventories (CDI), a widely used tool that collects information on children’s vocabulary comprehension and production through structured parental surveys.
+These surveys allow parents to report on their child’s understanding and use of specific words, grouped into lexical categories such as nouns, verbs, and adjectives.
+The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
+
+	-	Structured Response Formats: The CDI employs predefined response options for comprehension and production, reducing ambiguity and enhancing consistency in parental reporting.
+	-	Lexical Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
+	-	Norming Group Comparison: The inclusion of norming groups as benchmarks helps to ensure the validity of reported vocabulary scores and allows for cross-child comparison.
+These groups provide a standardized reference for analyzing individual differences in language acquisition.
+	- Variable Standardization: Continuous variables, such as age, are standardized (e.g., scaled) to reduce variability and improve the interpretability of statistical models.
+This ensures that coefficients reflect meaningful changes in relation to standardized measures.
+	-	Bias Mitigation: By structuring responses and including norming benchmarks, the CDI minimizes some of the biases inherent in self-reported data, such as over- or underestimation by parents.
+	-	Missing Data Handling: Observations with incomplete or invalid responses were excluded from the analysis to maintain the integrity and reliability of the dataset.
+
+### Reporting Bias
+
+However, there are several considerations regarding the data collection process:
+
+-	Parental Reporting Bias: The reliance on parental reports introduces the potential for bias, including overestimation or underestimation of a child’s abilities.
+This is inherent to self-reported data and can affect the accuracy of the results.
+-	Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents.
+This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
+-	Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison.
+While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
+-	Temporal Limitations: The CDI data represents snapshots of vocabulary development at specific ages, which may not account for rapid changes or variations over time in a child’s language acquisition process.
+
+Despite its standardized structure, the CDI is subject to biases inherent in parental reporting, including Over or Underestimation Bias.
+Parents may unintentionally over- or underestimate their child’s skills due to subjective perceptions or limited observations.
+Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
 
@@
 ## Predictor variables
+
 ### Age
-And also planes (@fig-age).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+
+@fig-age displays the distribution of children’s ages (in months) within the dataset, highlighting key patterns in the sample’s demographic structure.
+A notable concentration of data is observed among children aged between 24 and 30 months, reflecting an emphasis on capturing vocabulary development during critical periods of language acquisition.
+These age ranges are known to mark significant milestones in linguistic growth, which could explain their higher representation.
+Conversely, younger age groups (below 20 months) are underrepresented, likely due to the challenges of assessing vocabulary at earlier stages of development, where verbal communication is less pronounced and parental reporting is more variable.
+
+The dataset also shows distinct peaks at specific ages, such as 25 and 30 months.
+These sharp spikes may reflect intentional focus points for testing or developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
+This uneven age distribution underscores the importance of age as a critical factor in analyzing vocabulary acquisition.
+While the high concentration of data at older ages enhances insights into advanced vocabulary development, it also necessitates caution in generalizing findings to underrepresented age groups.
+This observation emphasizes the need to standardize age in statistical models to account for variability across different age groups.
 
@@
 #| label: fig-age
-#| fig-cap: Age Distribution
+#| fig-cap: It shows the distribution of children’s ages (in months) within the dataset.
+The majority of observations fall between 24 and 30 months, with noticeable peaks at 25 and 30 months, reflecting a focus on key developmental periods.
+Younger age groups are underrepresented, highlighting the need to account for variability in age when analyzing vocabulary acquisition patterns.
 #| warning: false
@@
 ```
+
 ### Broad Category
+The words in the dataset were grouped into broad lexical categories to facilitate the analysis of vocabulary acquisition patterns.
+These categories include Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs.
+The classification was based on the semantic and functional roles of words, with nouns subdivided into more specific groups such as Living Things, Objects, and Places to capture distinct trends in vocabulary acquisition.
+For instance, Function Words include pronouns and question words, reflecting grammatical development, while Verbs and Adjectives capture action and descriptive words, essential for sentence construction and expression.
+
+The bar graph illustrates the distribution of items across these categories, highlighting significant variation in word frequency.
+Objects constitute the largest category, suggesting a focus on tangible and concrete items, which are likely easier for children to recognize and recall.
+This is followed by Verbs and Living Things, categories that are fundamental to communication but slightly less prevalent.
+In contrast, Sensory Words and Activities are sparsely represented, possibly reflecting their specialized and context-dependent nature.
+The distribution underscores the importance of concrete and functional words in early vocabulary development while highlighting potential gaps in underrepresented categories.
+This variation provides a foundation for exploring how lexical diversity influences vocabulary acquisition patterns.
 
@@
 #| label: fig-category
-#| fig-cap:  Distribution of Word Categories.
-This bar plot illustrates the distribution of items across broad linguistic categories in the dataset.
-“Objects” dominate the dataset, followed by “Verbs” and “Living Things,” indicating that the dataset is rich in concrete nouns and action-related words, with smaller proportions of sensory words and adjectives.
+#| fig-cap: The figure shows objects dominate the vocabulary, reflecting an emphasis on concrete and tangible terms, while categories like Sensory Words and Activities are less frequently represented, indicating the relative complexity or specificity of these word types in early language acquisition
 #| warning: false
@@
   geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
-  labs(title = "Category Distribution", x = "Broad Category", y = "Count") +
+  labs(title = "Reclassified Category Distribution", x = "Broad Category", y = "Count") +
   theme_minimal() +
@@
 
-### Is Norming or Not
+### Norming Status
+
+Norming status categorizes children into two groups: those included in the norming group and those who are not.
+The norming group represents a standardized sample used as a benchmark for assessing vocabulary development, providing a reference point for evaluating other children in the dataset.
+This distinction is important for ensuring the validity and reliability of comparisons in the analysis.
+
+The bar plot illustrates the distribution of children based on their norming status.
+The dataset is predominantly composed of non-norming children, with only a small fraction representing the norming group.
+This imbalance reflects the practical challenges of including a broad and representative norming sample in large-scale vocabulary assessments.
+While the non-norming group provides diverse data for analysis, the norming group offers a crucial baseline for calibrating and interpreting the results.
+The distribution highlights the need to account for this imbalance in the analysis to avoid potential biases and ensure robust conclusions.
+
 ```{r}
@@
 #| label: fig-norming
-#| fig-cap: Norming Group Distribution.
-This plot shows the distribution of children in the dataset categorized by their norming status.
-The majority of the observations are from non-norming children, with only a small fraction representing norming children.
+#| fig-cap: The dataset is primarily composed of non-norming children, with a smaller subset belonging to the norming group, serving as a standardized benchmark for assessing vocabulary development
 #| warning: false
 #| message: false
+
 ggplot(analysis_data, aes(x = is_norming)) +
@@
 This threshold was chosen to distinguish children with relatively advanced vocabulary levels.
-More background details and diagnostics are included in [Appendix-@sec-model-details].
+More background details and diagnostics are included in [Appendix-@sec-model].
 
@@
 ## Model Assumptions
+
 - Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary) and the independent variables.
@@
 This approach ensures that the coefficients for each category represent the deviation of that category’s effect from the overall mean effect across all categories.
-For example, the coefficient for “Verbs” indicates how the log-odds of achieving a high vocabulary differ for “Verbs” compared to the average effect of all other categories.
+For example, the coefficient for “Verbs” indicates how the log odds of achieving a high vocabulary differ for “Verbs” compared to the average effect of all other categories.
 Sum contrasts are particularly useful for understanding relative effects and ensure that the intercept reflects the overall mean effect when all predictors are at their reference or average levels.
@@
 ## Interpretation of Coefficients
+
 The logistic regression coefficient ($\beta$) represents the change in the log-odds of the dependent variable (high vocabulary) for a one-unit change in the predictor variable, holding all other variables constant.
@@
 If $\beta_2 = 0.1$, then $\exp(0.1) \approx 1.11$, meaning being in the norming group increases the odds of high vocabulary by 11%.
-
 - Broad Category ($\beta_3$, $\beta_4$, $\beta_5$, etc.): The coefficients for `broad_category` represent the difference in log-odds compared to the reference category (“Adjectives”).  
-
-Interpretation:  
   - $\beta_3$ (Function Words): A positive $\beta_3$ indicates higher odds of having a high vocabulary for function words compared to adjectives.
 For instance, if $\beta_3 = 0.002$, then $\exp(0.002) \approx 1.002$, meaning the odds of having a high vocabulary for function words are 0.2% higher than for adjectives.
-  - $\beta_4$ (Living Things): A negative $\beta_4$ indicates lower odds of having a high vocabulary for living things compared to adjectives.
-For example, if $\beta_4 = -0.007$, then $\exp(-0.007) \approx 0.993$, meaning the odds of having a high vocabulary for living things are 0.7% lower than for adjectives.
-  - $\beta_5$ (Objects): If $\beta_5 = -0.008$, then $\exp(-0.008) \approx 0.992$, indicating that objects are associated with 0.8% lower odds of having a high vocabulary compared to adjectives.  
-  - $\beta_6$ (Verbs): If $\beta_6 = -0.005$, then $\exp(-0.005) \approx 0.995$, meaning verbs are associated with 0.5% lower odds of having a high vocabulary compared to adjectives.
-
-General Example for Broad Categories:  
-  If a coefficient $\beta_k = 0.01$, $\exp(0.01) \approx 1.01$, meaning the corresponding category increases the odds of having a high vocabulary by 1% compared to the reference category (Adjectives).
+General Example for Broad Categories: If a coefficient $\beta_k = 0.01$, $\exp(0.01) \approx 1.01$, meaning the corresponding category increases the odds of having a high vocabulary by 1% compared to the reference category (Adjectives).
 Conversely, if $\beta_k = -0.01$, $\exp(-0.01) \approx 0.99$, indicating a 1% decrease in odds compared to the reference category.
@@
 ## Average Production by Broad Category and Age
+
 @fig-average shows..
+
 ```{r}
@@
 #| warning: false
+#| message: false
 
@@
 ## Prediciton for the Probability of High Vocabulary Level
+
 ```{r}
@@
 
-### Distribution of Predicted Probabilities by Broad Category
+## Distribution of Predicted Probabilities by Broad Category
 
-@fig-prediction illustrates the distribution of predicted probabilities for high vocabulary levels across various broad categories.
+The @fig-prediction illustrates the distribution of predicted probabilities for high vocabulary levels across various broad categories.
 The density curves show distinct peaks and variability, reflecting the differences in how well each category predicts high vocabulary.
@@
 
+# Assuming `analysis_data_test` contains the predicted probabilities and broad categories
+
 ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
-  geom_density(alpha = 0.7, color = "black") +
-  labs(title = "Distribution of Predicted Probabilities by Broad Category",
+  geom_density(alpha = 0.7, color = "black") +  # Density plot with transparency
+  facet_wrap(~ broad_category, scales = "fixed") +  # Same y-axis range for all facets
+  labs(
+    title = "Distribution of Predicted Probabilities by Category",
     x = "Predicted Probability",
@@
   theme_minimal(base_size = 14) +
-  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
-    legend.position = "bottom",
-    legend.title = element_text(face = "bold"),
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
+    axis.title = element_text(size = 14),
+    axis.text = element_text(size = 12),
+    strip.text = element_text(size = 12, face = "bold"),  # Styling facet labels
+    legend.position = "none",  # Remove legend
     panel.grid.major = element_line(color = "gray85"),
     panel.grid.minor = element_blank()
-  )
+  ) +
+  scale_fill_brewer(palette = "Set2")  # Use a clean, appealing color palette
 ```
@@
 
-# Appendix {-}
+# Appendix {#-secdata}
 
-# Additional data details
+## Additional data details
 
@@
 
-# Model details {#sec-model-details}
+# Model details {#sec-model}
 
@@
 ```{r}
+#| echo: false
 #| eval: true
-#| echo: false
-#| message: false
+#| label: tbl-logistic
+#| tbl-cap: model summary table
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
 
-modelsummary(logistic_model, statistic = "std.error", fmt = 2)
+modelsummary(logistic_model)
 
@@
 #| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-cap: Checking the convergence of the MCMC algorithm
 
-#plot(first_model, "trace")
-
-#plot(first_model, "rhat")
 ```
@@
 
-Thanks to Open AI and ChatGPT 4.0 is used to write the paper.
-This project utilized @citeR and a variety of R packages that provided essential tools and functionality.
-We thank the @tidyverse, @ggplot2 and @knitr teams for their extensive suite of tools for data manipulation and visualization, which were instrumental in data cleaning and graphing.
-Finally, @arrow was vital for efficient data storage and handling with Parquet files.
-We are grateful to the developers and maintainers of these packages for their contributions to open-source software.
+This project was conducted under the help of OpenAI’s ChatGPT 4.0, which provided invaluable assistance in drafting and refining the paper.
+The analysis was conducted using a suite of packages from @citeR, which offered robust functionality for data manipulation, visualization, and storage.
+We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
+Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files.
+
+A special acknowledgment goes to the @wordbankr team for providing the extensive dataset that forms the foundation of this research.
+Their contribution enabled a comprehensive exploration of vocabulary learning patterns in children.
+We are deeply grateful to the developers and maintainers of these open-source tools and datasets for their efforts in advancing research and accessibility in the data science community.
 

### 2024-11-22 scripts/05-exploratory_data_analysis.R

@@ -103,7 +103,8 @@ ggplot(outcome_counts, aes(x = VocabularyCategory, y = n, fill = VocabularyCateg
   theme(
     text = element_text(size = 12),
     plot.title = element_text(hjust = 0.5)
-  )
+  ) + scale_y_continuous(
+    labels = scales::comma) 
 #-------------------------------------------------------------------------------
 ### Part C: Visualize Relation between Variables ###
 # C1. Vocabulary size by age:Plot Production by Age to see how children’s vocabulary develops over time
@@ -130,17 +131,17 @@ analysis_data %>%
     legend.title = element_text(size = 12),
     legend.text = element_text(size = 12))
 
-# C2. Average Production by Item Category and Age
+# C2. Average Production by Broad Category and Age
 analysis_data %>%
-  group_by(category, age) %>%
+  group_by(broad_category, age) %>%
   summarize(avg_production = mean(production, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(aes(x = age, y = avg_production, fill = category)) +
+  ggplot(aes(x = age, y = avg_production, fill = broad_category)) +
   geom_col(position = "stack") +  # Use stack for stacked bars
   labs(
-    title = "Average Production by Item Category and Age",
+    title = "Average Production by Broad Category and Age",
     x = "Age (months)",
     y = "Average Production Score",
-    fill = "Item Category"
+    fill = "Broad Category"
   ) +
   theme_minimal() +
   theme(

### 2024-11-23 scripts/03-clean_data.R
Restrict the date.

@@ -31,7 +31,7 @@ cleaned_data <- cleaned_data %>%
 
 # Filter data to include only rows with date_of_test after 2000-01-01
 cleaned_data <- cleaned_data %>% 
-  filter(date_of_test > as.Date("2000-01-01")) %>% 
+  filter(date_of_test > as.Date("2009-01-01")) %>% 
   filter(production > 100) %>% filter(comprehension > 100)
 
 # Define a mapping of categories to broader groups

### 2024-11-23 scripts/05-exploratory_data_analysis.R

@@ -45,10 +45,10 @@ ggplot(analysis_data, aes(x = category)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
-# B3. Analyze the distribution of 'lexical_category'
-ggplot(analysis_data, aes(x = lexical_category)) +
+# B3. Analyze the distribution of 'broad_category'
+ggplot(analysis_data, aes(x = broad_category)) +
   geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
-  labs(title = "Lexical Category Distribution", x = "Lexical Category", y = "Count") +
+  labs(title = "Broad Category Distribution", x = "Broad Category", y = "Count") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@ -65,19 +65,30 @@ ggplot(analysis_data, aes(x = production)) +
   theme_minimal()+scale_y_continuous(labels = scales::comma)
 
 # B6. Plot the Distribution of 'date_of_test'
-ggplot(analysis_data, aes(x = month_year, y = test_count)) +
+# Ensure date_of_test is in Date format
+analysis_data <- analysis_data %>%
+  mutate(date_of_test = as.Date(date_of_test))
+
+# Group by month-year and calculate the number of tests
+analysis_data_summary <- analysis_data %>%
+  group_by(date_of_test = floor_date(date_of_test, unit = "month")) %>%
+  summarize(test_count = n(), .groups = "drop")
+
+ggplot(analysis_data_summary, aes(x = date_of_test, y = test_count)) +
   geom_col(fill = "blue", color = "white", alpha = 1) + 
-  labs(title = "Monthly Distribution of Test Dates",
+  labs(
+    title = "Monthly Distribution of Test Dates",
     subtitle = "Visualizing the count of tests conducted each month",
     x = "Month-Year",
-    y = "Number of Tests") +
-  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months",  
-    expand = c(0.01, 0)) +
-  scale_y_continuous(
-    labels = scales::comma) +
-  theme_minimal(base_size = 14) +  # Clean minimal theme
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),  
-    plot.title = element_text(face = "bold", size = 16, hjust = 0.5))
+    y = "Number of Tests"
+  ) +
[…]

### 2024-11-23 scripts/06-model_data.R

@@ -15,12 +15,19 @@ library(rstanarm)
 library(car)
 library(arrow)
 library(rsample)
+library(modelsummary)
 
 #### Read data ####
 analysis_data<- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 ### Model data ####
 set.seed(853)
+
+# Check and preprocess data
+analysis_data <- analysis_data %>%
+  mutate(broad_category = factor(broad_category), # Ensure categorical variable is a factor
+    age_scaled = as.numeric(scale(age))     # Standardize numerical predictors
+  )
 
 # Perform a split to get training and testing data
 split <- initial_split(data=analysis_data, prop = 0.8)
@@ -36,20 +43,14 @@ analysis_data_train <- analysis_data_train %>%
   ) %>%
   drop_na(prod_comp_mean)  # Drop rows where prod_comp_mean is NA
 
-# Standardize numerical predictors
-analysis_data_train <- analysis_data_train %>%
-  mutate(age_scaled = scale(age))
-
-analysis_data_train$age_scaled <- as.numeric(analysis_data_train$age_scaled)
-
 # Build the logistic regression model on the training dataset
 logistic_model <- glm(
   high_vocabulary ~ age_scaled + is_norming + broad_category,
   data = analysis_data_train,
-  family = binomial)
+  family = binomial(link = "logit"))
 
 # Print the summary of the model
-summary(logistic_model)
+modelsummary(logistic_model)
 
 #### Save model ####
 write_parquet(analysis_data_test, sink = "data/02-analysis_data/test_data.parquet")

### 2024-11-23 paper/paper.qmd
Minor updats in scripts.

@@
 
-- High Vocabulary: The outcome variable, high_vocabulary, is a binary indicator that takes the value of 1 if the average production and comprehension score (prod_comp_mean) exceeds 350 and 0 otherwise.
-This threshold was selected to represent children with relatively advanced vocabulary skills, determined through exploratory data analysis.
-- Scaled Age (age_scaled): This continuous variable represents the child’s age, standardized to ensure the model coefficients reflect changes per standard deviation in age.
-Standardization improves numerical stability and aids interpretability.
+- High Vocabulary: A binary indicator where 1 represents a high vocabulary score (combined comprehension and production > 350), and 0 otherwise.
+
+- Scaled Age (age_scaled): The child’s age, standardized to reflect changes per standard deviation. Standardization aids in interpretability and ensures numerical stability.
+
 - Norming Status (is_norming): A binary indicator denoting whether a child is part of the norming dataset (TRUE) or not (FALSE).
 This variable accounts for potential differences in data collection or assessment protocols.
-- Broad Category (broad_category): A categorical variable grouping words into four broad linguistic categories: nouns, verbs, adjectives, and function_words. The reference category is function_words.
 
-The model takes the form:
+- Broad Category (broad_category): A categorical variable grouping words into lexical categories, such as adjectives, verbs, and nouns. The reference category for comparison is Function Words.
+
+The model is specified as:
 
@@
 Where:
-- $p_i$ represents the probability that person i has a high vocabulary
+- $p_i$ represents the probability that child  i  has a high vocabulary score.
 - $\beta_0$ is the intercept, capturing the baseline log-odds when all predictors are at their reference or mean levels
@@
 
-The model was trained on 80% of the dataset, with the remaining 20% reserved for testing.
-Model performance was assessed using confusion matrices and overall accuracy.
-Further evaluation included the analysis of residual deviance, AIC, and interpretation of individual coefficients.
+The model was trained using robust statistical principles to ensure reliability:
+
+	•	The split of data into training and testing subsets facilitates evaluation of the model’s predictive power on unseen data.
+	•	Standardizing continuous predictors such as age improves interpretability and comparability across variables.
+	•	Saving preprocessed datasets and the trained model ensures reproducibility and traceability of results.
 
@@
 
-## Average Production by Broad Category and Age
+## Average Production by Broad Category
 
@@
 
-analysis_data %>%
-  group_by(broad_category, age) %>%
-  summarize(avg_production = mean(production, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(aes(x = age, y = avg_production, fill = broad_category)) +
-  geom_col(position = "stack") +  # Use stack for stacked bars
+# Filter the dataset for Adjectives and Objects
+filtered_data <- analysis_data %>%
+  filter(broad_category %in% c("Adjectives", "Objects"))
+
+# Select 100 random unique child IDs
+set.seed(123)  # Ensure reproducibility
+sampled_child_ids <- filtered_data %>%
+  distinct(child_id) %>%
+  slice_sample(n = 100) %>%
+  pull(child_id)
+
+# Filter data for these 100 children
+filtered_data_sampled <- filtered_data %>%
+  filter(child_id %in% sampled_child_ids)
+
+# Create a line plot for each child's production trend
+ggplot(filtered_data_sampled, aes(x = age, y = production, color = as.factor(child_id), group = child_id)) +
+  geom_line(alpha = 0.8, size = 1) +  # Line for each child
+  facet_wrap(~ broad_category, scales = "free_y") +  # Separate facets for Adjectives and Objects
   labs(
-    title = "Average Production by Broad Category and Age",
-    x = "Age (months)",
-    y = "Average Production Score",
-    fill = "Broad Category"
+    title = "Production Trends by Child Across Categories (Sampled 100 Children)",
+    subtitle = "Adjectives and Objects",
+    x = "Age (Months)",
+    y = "Production Score",
+    color = "Child ID"
   ) +
@@
   theme(
-    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(size = 12, hjust = 0.5),
     axis.text = element_text(size = 10),
-    axis.title = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.title = element_text(size = 11))
+    axis.title = element_text(size = 14),
+    legend.position = "none",  # Hide legend to avoid clutter
+    strip.text = element_text(size = 12, face = "bold")  # Customize facet labels
+  )
 ```
@@
 ```{r}
+library(ggplot2)
+
+# Add predicted probabilities and logit to the dataset
+analysis_data_train <- analysis_data_train %>%
+  mutate(predicted_prob = predict(logistic_model, type = "response"),
+         logit = log(predicted_prob / (1 - predicted_prob)))
+
+# Plot logit against continuous variables (e.g., age_scaled)
+ggplot(analysis_data_train, aes(x = age_scaled, y = logit)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "loess", se = TRUE, color = "blue") +
+  labs(title = "Logit vs. Age (Scaled)",
+       x = "Age (Scaled)",
+       y = "Logit (log-odds)") +
+  theme_minimal()
+```
+
+### Accuracy
+```{r}
+# Predicted probabilities and classes
+predicted_classes <- ifelse(fitted(logistic_model) > 0.5, 1, 0)
+confusion_matrix <- table(Predicted = predicted_classes, Actual = analysis_data_train$high_vocabulary)
+
+# Print confusion matrix and accuracy
+print(confusion_matrix)
+accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
+print(paste("Accuracy:", round(accuracy, 3)))
+```
+
+```{r}
 #| echo: false
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: Checking the convergence of the MCMC algorithm
+#| label: fig-convergencecheck
+#| fig-cap: "Checking the convergence of the Markov chain Monte Carlo (MCMC) algorithm for hit song model"
+#| fig-subcap: ["Trace plot for model", "Rhat plot for model"]
+#| layout-ncol: 2
 
+# Creates trace plot
+plot(predicted_prob, "trace") + theme_minimal()
+# Creates rhat plot
+plot(predicted_prob, "rhat") + theme_minimal()
+
+```
+
+### Binned Residual Plot
+```{r}
+install.packages("arm")
+library(arm)
+
+binnedplot(fitted(logistic_model), residuals(logistic_model, type = "response"),
+           xlab = "Fitted Probabilities", ylab = "Average Residual",
+           main = "Binned Residual Plot")
 ```

### 2024-11-23 scripts/02-download_data.R

@@ -37,6 +37,6 @@ if (!is.null(items)) {
 
 
 #### Save data ####
-# Because the dataset is too large (1.2GB in csv file). Alternatively, save the downloaded data as a Parquet file.
+# Because the dataset is too large. Alternatively, save the downloaded data as a Parquet file.
 write_parquet(x = raw_data, sink = "data/01-raw_data/raw_data.parquet")
 

### 2024-11-23 scripts/03-clean_data.R

@@ -20,7 +20,7 @@ library(arrow)
 raw_data <- read_parquet("data/01-raw_data/raw_data.parquet")
 
 # Select the needed columns
-cleaned_data <- raw_data %>% select(data_id, language, form, item_kind, 
+cleaned_data <- raw_data %>% dplyr::select(data_id, language, form, item_kind, 
                                  category, uni_lemma, date_of_test,
                                  age, comprehension, production, 
                                  is_norming, child_id)
@@ -29,10 +29,9 @@ cleaned_data <- raw_data %>% select(data_id, language, form, item_kind,
 cleaned_data <- cleaned_data %>%
   mutate(date_of_test = as.Date(date_of_test))
 
-# Filter data to include only rows with date_of_test after 2000-01-01
+# From EDA, observing the lack of data between 2004-2009, filter data to include only observations with date_of_test after 2009-01-01
 cleaned_data <- cleaned_data %>% 
-  filter(date_of_test > as.Date("2009-01-01")) %>% 
-  filter(production > 100) %>% filter(comprehension > 100)
+  filter(date_of_test > as.Date("2009-01-01"))
 
 # Define a mapping of categories to broader groups
 cleaned_data <- cleaned_data %>%
@@ -50,10 +49,12 @@ cleaned_data <- cleaned_data %>%
 
 # Combine production and comprehension into a new variable
 cleaned_data <- cleaned_data %>%
-  mutate(prod_comp_mean = (production + comprehension)/2,  # Average of production and comprehension
-         high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0)  # Binary target variable
-  ) %>%
+  mutate(prod_comp_mean = (production + comprehension)/2) %>%
   drop_na(prod_comp_mean) 
+
+# Create the high_vocabulary variable based on prod_comp_mean
+cleaned_data <- cleaned_data %>%
+  mutate(high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0))
 
 # Omit rows with NA values
 cleaned_data <- cleaned_data %>% 

### 2024-11-23 scripts/04-test_analysis_data.R

@@ -31,8 +31,8 @@ test_that("dataset has the expected number of rows", {
 })
 
 # Test 2: Test if the dataset has the correct number of columns
-test_that("dataset has 14 columns", {
-  expect_equal(ncol(analysis_data), 14)
+test_that("dataset has 15 columns", {
+  expect_equal(ncol(analysis_data), 15)
 })
 
 # Test 3: Test if the 'language' column contains only valid languages
@@ -65,13 +65,13 @@ test_that("'age' values are between 16 and 60", {
 })
 
 # Test 7: Test if 'comprehension' falls within a valid range
-test_that("'comprehension' values are between 100 and 1000", {
-  expect_true(all(analysis_data$comprehension >= 100 & analysis_data$comprehension <= 1000))
+test_that("'comprehension' values are between 0 and 1000", {
+  expect_true(all(analysis_data$comprehension >= 0 & analysis_data$comprehension <= 1000))
 })
 
 # Test 8: Test if 'production' falls within a valid range
-test_that("'production' values are between 100 and 1000", {
-  expect_true(all(analysis_data$production >= 100 & analysis_data$production <= 1000))
+test_that("'production' values are between 0 and 1000", {
+  expect_true(all(analysis_data$production >= 0 & analysis_data$production <= 1000))
 })
 
 # Test 9: Test if 'date_of_test' is within a valid range
@@ -80,9 +80,10 @@ test_that("'date_of_test' is after the year 2000", {
                     analysis_data$date_of_test <= as.Date("2024-11-20")))
 })
 
-# Test 10: Test if item_id values are unique for each data_id
-test_that("item_id values are unique for each data_id", {
-  expect_true(all(!duplicated(analysis_data[, c("data_id", "item_id")])))
+# Test 10: Test if 'child_id' is consistent with its  'data_id'
+test_that("child_id is consistent within each data_id", {
+  expect_true(all(tapply(analysis_data$child_id, analysis_data$data_id, 
+                         function(x) length(unique(x)) == 1)))
 })
 
 # Test 11: Test if comprehension is greater than or equal to production
@@ -102,7 +103,8 @@ test_that("'date_of_test' contains no future dates", {
 
 # Test 14: Test if broad_category contains the valid type categories
 test_that("'broad_category' column is valid", {
-  valid_broad_categories <- c("verbs", "function_words", "nouns", "adjectives")
+  valid_broad_categories <- c("Sensory Words", "Living Things", "Objects", "Places"
[…]

### 2024-11-23 scripts/05-exploratory_data_analysis.R

@@ -188,7 +188,7 @@ ggplot(monthly_summary, aes(x = month)) +
 ### Part D: Model data ####
 # D1. Define the Bayesian regression model
 bayesian_model <-stan_glm(
-    formula = production ~ age + category + lexical_category,  # Model formula
+    formula = production ~ age + category + broad_category,  # Model formula
     data = analysis_data,                               # Dataset
     family = gaussian(),                                 # Gaussian distribution for continuous outcome
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -231,15 +231,15 @@ analysis_data <- analysis_data %>%
     prod_comp_mean_scaled = scale(prod_comp_mean)  # Standardize prod_comp_mean
   )
 
-# Collapse `lexical_category` into top 5 most frequent levels and group the rest
+# Collapse `broad_category` into top 5 most frequent levels and group the rest
 analysis_data <- analysis_data %>%
   mutate(
-    lexical_category = fct_lump(lexical_category, n = 5)  # Top 5 frequent categories
+    broad_category = fct_lump(broad_category, n = 5)  # Top 5 frequent categories
   )
 
 # Build the logistic regression model
 logistic_model <- glm(
-  high_comprehension ~ age_scaled + prod_comp_mean_scaled + lexical_category,  # Formula
+  high_comprehension ~ age_scaled + prod_comp_mean_scaled + broad_category,  # Formula
   data = analysis_data,  # Data
   family = binomial,  # Logistic regression
   weights = ifelse(high_comprehension == 1, 10, 1)  # Adjust weights for class imbalance
@@ -253,9 +253,9 @@ summary(logistic_model)
 analysis_data <- analysis_data %>%
   mutate(avg_comp_prod = (comprehension + production) / 2)
 
-# Fit the linear model with age, lexical_category, is_norming as indicator variables
+# Fit the linear model with age, broad_category, is_norming as indicator variables
 linear_model <- lm(
-  avg_comp_prod ~ age + lexical_category + is_norming, data = analysis_data)
+  avg_comp_prod ~ age + broad_category + is_norming, data = analysis_data)
 
 # Print the summary
 summary(linear_model)

### 2024-11-23 scripts/06-model_data.R

@@ -25,9 +25,9 @@ set.seed(853)
 
 # Check and preprocess data
 analysis_data <- analysis_data %>%
-  mutate(broad_category = factor(broad_category), # Ensure categorical variable is a factor
-    age_scaled = as.numeric(scale(age))     # Standardize numerical predictors
-  )
+  mutate(
+    age_scaled = as.numeric(scale(age)),  # Standardize age
+    broad_category = factor(broad_category))
 
 # Perform a split to get training and testing data
 split <- initial_split(data=analysis_data, prop = 0.8)

### 2024-11-23 paper/paper.qmd
Paper writing updated

--- 2024-11-22 paper/paper.qmd
@@
 ---
-title: "English Vocabulary Learning Pattern"
+title: "Children English Vocabulary Learning Pattern"
 subtitle: "My subtitle if needed"
@@
 
+#install.packages("arm")
 library(tidyverse)
@@
 library(wordbankr)
+library(tidyr)
+library(arm)
 
@@
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data descirbe the dataset in detail, and shows the distribution of variables.
-
+The remainder of this paper is structured as follows.
+@sec-data descirbe the dataset in detail, and shows the distribution of variables.
 The data used in this study comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR
@@
 summary_table <- analysis_data %>%
-  select(language, age, is_norming, broad_category, production, high_vocabulary) %>%
+  dplyr::select(language, age, is_norming, broad_category, production, high_vocabulary) %>%
   rename(
@@
 The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
-
 - Structured Response Formats: The CDI employs predefined response options for comprehension and production, reducing ambiguity and enhancing consistency in parental reporting.
-	-	Lexical Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
+- Words Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
 - Norming Group Comparison: The inclusion of norming groups as benchmarks helps to ensure the validity of reported vocabulary scores and allows for cross-child comparison.
@@
 
-Despite its standardized structure, the CDI is subject to biases inherent in parental reporting, including Over or Underestimation Bias.
-Parents may unintentionally over- or underestimate their child’s skills due to subjective perceptions or limited observations.
+Despite its standardized structure, the CDI is subject to biases inherent in parental reporting, including over or underestimation bias.
+Parents may unintentionally overestimate or underestimate their child’s skills due to subjective perceptions or limited observations.
 Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
@@
 #| message: false
+
 outcome_counts <- analysis_data %>%
@@
   labs(title = "Age Distribution", x = "Age (months)", y = "Count") +
-  theme_minimal()
+  theme_minimal()+
+  scale_y_continuous(labels = scales::comma)
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
+  scale_y_continuous(labels = scales::comma)
 ```
+
+\newpage
 
@@
 
-ggplot(analysis_data, aes(x = is_norming)) +
-  geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
-  labs(title = "Norming Distribution", x = "Norming or not", y = "Count") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_y_continuous(labels = scales::comma)
+summary <- summary(analysis_data$is_norming)
+# Summarize the 'is_norming' variable
+summary_isnorming <- table(analysis_data$is_norming)
+
+# Convert the table into a data frame
+summary_isnorming_df <- as.data.frame(summary_isnorming)
+
+# Render the summary table using kable
+knitr::kable(
+  summary_isnorming_df,
+  col.names = c("Norming Status", "Count"),  # Set column names
+  caption = "Summary of Norming Status in the Dataset"
+)
 ```
@@
 Where:
-- $p_i$ represents the probability that person i has a high vocabulary
+- $p_i$ represents the probability that child  i  has a high vocabulary score
 - $\beta_0$ is the intercept, capturing the baseline log-odds when all predictors are at their reference or mean levels
@@
 
-The model was trained on 80% of the dataset, with the remaining 20% reserved for testing.
-Model performance was assessed using confusion matrices and overall accuracy.
-Further evaluation included the analysis of residual deviance, AIC, and interpretation of individual coefficients.
+Logistic regression is widely used in predictive modeling for categorical outcomes due to its simplicity and robustness.
+It provides probabilities that are constrained between 0 and 1, ensuring meaningful interpretations for binary outcomes.
+Unlike other complex models, logistic regression allows for clear coefficient interpretation, offering insights into the magnitude and direction of predictor effects.
+For example, the odds ratios derived from logistic regression help explain how changes in variables like age or norming status influence the probability of high vocabulary acquisition.
+
+Although advanced machine learning models like decision trees, random forests, or neural networks could be used, these methods often lack the interpretability of logistic regression.
+While these models might yield slightly better predictive performance, they are often considered “black boxes,” making it difficult to identify specific relationships between predictors and outcomes.
+Given the study’s focus on understanding developmental patterns rather than maximizing predictive accuracy, logistic regression is more appropriate.
+
+Moreover, complex models require larger datasets to avoid overfitting and ensure generalizability, which might not be feasible given the sample size and the structure of the data in this study.
+Logistic regression strikes a balance between simplicity, interpretability, and predictive performance.
+
+The dataset was split into training and testing subsets to ensure model validation and reduce overfitting.
+This allows the model to generalize better to unseen data, providing a more reliable assessment of its predictive accuracy.
+Standardizing continuous predictors, such as age, enhances interpretability and ensures that variables are on a comparable scale, preventing dominance by predictors with larger numerical ranges.
+
+To ensure reproducibility, the preprocessed datasets and the trained model were saved.
+This practice facilitates verification of results and supports future analyses or extensions of the study.
+Overall, the logistic regression model offers a clear and interpretable framework for investigating vocabulary acquisition patterns, providing both explanatory power and practical insights.
 
@@
 
-## Average Production by Broad Category and Age
+## Variability in Production Vocabulary
 
-@fig-average shows..
+@fig-percent visualizes the relationship between age and production vocabulary scores, focusing on different percentiles of the distribution.
+The scatterplot shows individual production scores as gray dots, while overlaid lines represent percentiles (10th, 25th, 50th, 75th, and 90th), capturing central tendencies and variability across ages.
+The 50th percentile (median) line provides a benchmark for the typical vocabulary production score at each age, whereas the 10th and 90th percentiles outline the lower and upper ranges of vocabulary development.
+The gradual upward trend of the median line reflects consistent growth in production vocabulary as children age, with a widening gap between the percentiles at later ages.
+This widening suggests increasing variability in vocabulary acquisition, with some children advancing much faster than others in production abilities.
+
+The data indicates that children in the 90th percentile acquire vocabulary at a significantly faster rate than their peers, as evidenced by the steeper slope of the topmost line.
+Conversely, the 10th and 25th percentiles show more gradual, stable growth, suggesting slower development for children in these groups.
+The broader range of scores at older ages emphasizes the heterogeneity of developmental trajectories, with some children reaching vocabulary sizes substantially larger than the median while others remain below average.
+These findings underscore the diversity in early language acquisition and highlight the importance of considering individual differences when evaluating children’s vocabulary development.
 
@@
 #| eval: true
-#| label: fig-average
-#| fig-cap: xx 
+#| label: fig-percent
+#| fig-cap: Production Vocabulary by Age with Percentile Lines.
+The graph illustrates the production scores of children across different ages (in months).
+Individual data points (gray dots) represent raw production scores for each participant.
+Colored lines correspond to standardized percentiles—10th (blue), 25th (light blue), 50th (green), 75th (yellow), and 90th (red)—showing trends in vocabulary production distribution over time.
 #| warning: false
@@
 
-analysis_data %>%
-  group_by(broad_category, age) %>%
-  summarize(avg_production = mean(production, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(aes(x = age, y = avg_production, fill = broad_category)) +
-  geom_col(position = "stack") +  # Use stack for stacked bars
+# Simulate some data for demonstration purposes
+set.seed(666)
+analysis_data <- data.frame(
+  age = runif(1000, 15, 30),
+  production = rgamma(1000, shape = 2, scale = 50))
+
+# Define the percentiles and process the data
+percentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)
+
+percentile_data <- analysis_data %>%
+  mutate(age_rounded = round(age)) %>%
+  group_by(age_rounded) %>%
+  summarize(
+    `10th` = quantile(production, probs = 0.1, na.rm = TRUE),
+    `25th` = quantile(production, probs = 0.25, na.rm = TRUE),
+    `50th` = quantile(production, probs = 0.5, na.rm = TRUE),
+    `75th` = quantile(production, probs = 0.75, na.rm = TRUE),
+    `90th` = quantile(production, probs = 0.9, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  pivot_longer(
+    cols = -age_rounded,
+    names_to = "percentile",
+    values_to = "value")
+
+# Visualize the data
+ggplot() +
+  geom_point(data = analysis_data, aes(x = age, y = production), alpha = 0.2, color = "grey") +
+  geom_line(data = percentile_data, aes(x = age_rounded, y = value, color = percentile, group = percentile), size = 1) +
+  scale_color_manual(
+    values = c("blue", "lightblue", "green", "gold", "red"),
+    labels = c("10th", "25th", "50th", "75th", "90th")
+  ) +
   labs(
-    title = "Average Production by Broad Category and Age",
-    x = "Age (months)",
-    y = "Average Production Score",
-    fill = "Broad Category"
+    title = "Production Vocabulary by Age with Percentile Lines",
+    x = "Age (Months)",
+    y = "Production (Number of Words)",
+    color = "Percentile"
   ) +
-  theme_minimal() +
+  theme_minimal(base_size = 14) +
   theme(
-    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.text = element_text(size = 10),
+    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
+    legend.position = "top"
+  )
+
+```
+
+## Vocabulary Size Change by Age
+
+@fig-median illustrates the median comprehension vocabulary scores across different ages, focusing on the central tendency of children’s comprehension development between 15 and 30 months.
+The gray dots represent individual data points, capturing the variability in comprehension scores, while the blue line highlights the median score for each age group.
+The graph shows a clear upward trajectory, with median comprehension steadily increasing with age, particularly after 18 months.
+This suggests a critical developmental period between 18 and 30 months during which children experience significant growth in comprehension vocabulary.
+The density and spread of gray points around the median line indicate individual variability, emphasizing that while the general trend is one of growth, some children exhibit slower or faster development compared to their peers.
+The visualization underscores the importance of age as a determinant of vocabulary comprehension while highlighting the diverse range of learning patterns among children.
+
+Steady Growth in Comprehension Vocabulary: Between 15 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
+A noticeable increase in vocabulary size is observed after 18 months, suggesting that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop.
+The most significant growth occurs between 24 and 30 months, where the median comprehension score consistently rises.
+This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.The distribution of points (grey scatter) highlights considerable variability in comprehension scores at each age.
+While the median line captures the central trend, some children show significantly higher or lower comprehension compared to their peers, reflecting individual differences in language learning rates.Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.
+
+The use of the median instead of the mean ensures that the central trend is not skewed by outliers (e.g., extremely high or low comprehension scores).
+This choice provides a robust summary of comprehension at each age, especially in datasets with large variability or non-normal distributions.These findings highlight the critical window between 21 and 26 months for comprehension vocabulary growth.
+Interventions or language exposure strategies during this period may be particularly effective in enhancing language development.
+The observed variability suggests that individual-level factors (e.g., family environment, exposure to language) play a significant role in shaping comprehension scores, warranting further investigation into these influences.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-median
+#| fig-cap: "Illustrating central tendencies in comprehension vocabulary scores"
+#| warning: false
+#| message: false
+
+analysis_data_1<- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
+
+# Calculate the median comprehension for each age
+median_data <- analysis_data_1 %>%
+  group_by(age) %>%
+  summarize(
+    median_prod_comp_mean = median(prod_comp_mean, na.rm = TRUE), 
+    .groups = "drop")
+
+# Create a scatter plot with median comprehension by age
+ggplot() +
+  geom_jitter(data = analysis_data_1, aes(x = age, y = prod_comp_mean), 
+              alpha = 0.2, size = 1, color = "orange") +  # Scatter points for individual data
+  geom_line(data = median_data, aes(x = age, y = median_prod_comp_mean), 
+            color = "blue", size = 1.5, linetype = "solid") +  # Line for median comprehension
+  labs(
+    title = "Median Comprehension Vocabulary by Age", x = "Age (Months)",
+    y = "Median of the Vocabulary Level") +
+  theme_minimal(base_size = 14) +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(size = 12, hjust = 0.5),
     axis.title = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.title = element_text(size = 11))
+    axis.text = element_text(size = 12),
+    panel.grid.major = element_line(color = "grey85"),
+    panel.grid.minor = element_blank()) +
+  scale_x_continuous(breaks = seq(15, 30, 3))  
 ```
@@
 # Add predicted probabilities to the test dataset
-analysis_data_test <- analysis_data_test %>%
-  mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))
+analysis_data_test <- analysis_data_test %>% mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))
 
@@
 This suggests that broad categories contribute differently to vocabulary development, with nouns and verbs potentially playing a more significant role.
+
+Each facet represents a unique category, with the x-axis denoting the predicted probability of high vocabulary and the y-axis representing density.
+
+The plots highlight distinct patterns in the distributions of predicted probabilities for each category.
+Categories such as Adjectives and Function Words exhibit broader distributions, suggesting greater variability in the likelihood of achieving a high vocabulary score within these categories.
+In contrast, categories like Sensory Words and Activities show relatively concentrated distributions near lower probabilities, indicating more consistent outcomes within those lexical groups.
+
+The graph underscores the heterogeneity of vocabulary acquisition probabilities across different lexical domains.
+Categories associated with more concrete or commonly encountered words, such as Objects and Living Things, tend to exhibit higher probabilities for vocabulary mastery, reflecting their early and frequent usage in language development.
+Conversely, categories like Sensory Words show lower probabilities, suggesting that they are acquired less consistently or later in development.
+This visualization offers insight into the differential patterns of lexical acquisition, emphasizing the varied developmental trajectories across word types.
 
@@
     panel.grid.major = element_line(color = "gray85"),
-    panel.grid.minor = element_blank()
-  ) +
-  scale_fill_brewer(palette = "Set2")  # Use a clean, appealing color palette
+    panel.grid.minor = element_blank())
+
 ```
@@
 
-## First discussion point {#sec-first-point}
+## Vocabulary Development Patterns {#sec-firstpoint}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The acceleration in early vocabulary is even clearer when looking at production reports from older children using Words & Sentences.
+Figure 5.3 shows this pattern.
+In every language, the median child is reported to produce 50 words between 16–20 months (dotted line), though – as we will see below – this analysis masks tremendous between-child variability during this period.
+In addition, languages vary considerably in the absolute number of words reported.
+(As it is a major outlier, we have discussed the Beijing Mandarin WS data in Chapter 3, section on difficult data).
+Nevertheless, there are still substantial consistencies in the shape and general numerical range across languages.
 
-## Second discussion point
+During the period of 24–30 months, we see curves leveling out.
+Presumably, this leveling does not reflect a slowing in the rate of acquisition, which most researchers assume continues unabated for many years (e.g., Bloom, Tinker, and Scholnick 2001).
+Instead, it reflects the limitations of the CDI instrument, in that there are many possible “more advanced” words that children are likely learning, of which only a small subset are represented on any form.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+The results highlight distinct trajectories in vocabulary development among children.
+Production scores showed a steady increase with age, with notable accelerations between 18 and 30 months.
+This period is critical for vocabulary acquisition, as children rapidly expand their linguistic capabilities.
+The variability observed in percentile distributions emphasizes individual differences in learning rates, influenced by factors such as exposure to language, socio-economic background, and cognitive abilities.
+
+The analysis of comprehension data revealed a similar trend, with a marked rise in vocabulary comprehension after 18 months.
+Median comprehension scores provided a robust central tendency, effectively avoiding skewness caused by outliers.
+The plateau observed around 30 months suggests a stabilization in vocabulary growth, likely reflecting the limitations of the CDI dataset rather than a developmental plateau.
+
+## Role of Predictors in Vocabulary Growth
+
+Key predictors, including age, norming status, and broad lexical categories, significantly influenced vocabulary acquisition.
+Age, as expected, showed the strongest correlation, underlining its role as a primary determinant of language development.
+Children from the norming group exhibited higher vocabulary scores, possibly due to more structured linguistic environments or assessment settings.
+
+Lexical categories revealed nuanced contributions to vocabulary development.
+Categories such as objects and verbs were associated with higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
+Conversely, categories like sensory words and adjectives showed less consistent trends, indicating their later emergence in the linguistic repertoire.
 
@@
 
-## Weaknesses and next steps
+## Limitations and Future Directions
 
-Weaknesses and next steps should also be included.
+While the findings offer valuable insights, several limitations should be acknowledged.
+First, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
+Future studies could complement CDI data with observational or experimental measures to enhance reliability.
+Second, the cross-sectional nature of the data limits the ability to track individual developmental trajectories.
+Longitudinal studies are needed to capture within-child variability and the dynamics of vocabulary growth over time.
+
+Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition.
+These factors could provide a more comprehensive understanding of the mechanisms underlying linguistic development.
+
+The strongest developmental inferences can be made by the examination of longitudinal data, in which children’s individual development is measured multiple times using the same instrument.
+Unfortunately, relatively little of our CDI data comes from this type of repeated administration.
+Figure 3.5 shows the number of administrations for particular languages that come from longitudinal datasets with a particular depth.
+There is a substantial amount of two-administration longitudinal data for several languages, but only a few have more than two observations for individual children.In general, this aspect of our data is a consequence of the fact that, for normative datasets, pure cross-sectional data collection is used to ensure statistical independence between datapoints.
+Thus, we must typically settle for using the large amount of available cross-sectional data to average out individual variability.
 
@@
 
+### Accuracy
 ```{r}
 #| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: Checking the convergence of the MCMC algorithm
 
+analysis_data_train <- read_parquet(here::here("data/02-analysis_data/train_data.parquet"))
+
+# Predicted probabilities and classes
+predicted_classes <- ifelse(fitted(logistic_model) > 0.5, 1, 0)
+confusion_matrix <- table(Predicted = predicted_classes, Actual = analysis_data_train$high_vocabulary)
+
+# Print confusion matrix and accuracy
+print(confusion_matrix)
+accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
+print(paste("Accuracy:", round(accuracy, 3)))
+```
+
+```{r}
+#| echo: false
+
+# Predict probabilities using the logistic model
+analysis_data_test <- analysis_data_test %>%
+  mutate(predicted_prob = predict(logistic_model, newdata = ., type = "response"))
+
+# Prepare dataset for visualization
+analysis_data_test <- analysis_data_test %>%
+  mutate(high_vocabulary = factor(high_vocabulary, labels = c("0", "1")))
+
+# Improved Panel (a): Scatterplot of Predicted Probabilities
+ggplot(analysis_data_test, aes(x = age_scaled, y = predicted_prob, color = high_vocabulary)) +
+  geom_point(alpha = 0.6, size = 3, shape = 16, position = position_jitter(width = 0.05, height = 0.02)) +
+  labs(
+    title = "Panel (a): Predicted Probabilities by Age (Scaled)",
+    x = "Age (Scaled)",
+    y = "Predicted Probability of High Vocabulary",
+    color = "Observed Outcome"
+  ) +
+  scale_color_manual(values = c("0" = "#E41A1C", "1" = "#377EB8"), 
+                     labels = c("Low Vocabulary (0)", "High Vocabulary (1)")) +
+  theme_minimal(base_size = 14) +
+  theme(
+    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+    axis.title = element_text(size = 14),
+    axis.text = element_text(size = 12),
+    legend.position = "top",
+    legend.title = element_text(size = 12, face = "bold"),
+    legend.text = element_text(size = 12)
+  )
+
+```
+
+### Binned Residual Plot
+```{r}
+#| echo: false
+
+binnedplot(fitted(logistic_model), residuals(logistic_model, type = "response"),
+           xlab = "Fitted Probabilities", ylab = "Average Residual",
+           main = "Binned Residual Plot")
 ```
@@
 The analysis was conducted using a suite of packages from @citeR, which offered robust functionality for data manipulation, visualization, and storage.
-We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
+We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
 Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files.

### 2024-11-24 paper/paper.qmd
paper draft writing

@@
 ---
-title: "Children English Vocabulary Learning Pattern"
-subtitle: "My subtitle if needed"
+title: "Exploring Vocabulary Acquisition in Children"
+subtitle: "Predicted Vocabulary Growth in Children: Higher Probabilities for Common Words and Greater Variability in Complex Categories"
 author: Yongqi Liu
@@
 
-#install.packages("arm")
+# List of required packages
+# packages <- c("tidyverse", "arrow", "modelsummary", "knitr", "dplyr", 
+#            "ggplot2", "scales", "wordbankr", "tidyr", "arm")
+
 library(tidyverse)
@@
 # Introduction
+Understanding how children acquire vocabulary is a cornerstone of developmental linguistics and early childhood education.
+Vocabulary growth is not only a key indicator of cognitive development but also serves as a foundation for future linguistic and academic success.
+The interplay of factors such as age, linguistic environments, and word categories influences children’s vocabulary acquisition in complex and dynamic ways.
+By analyzing patterns and predictors of vocabulary growth, researchers can uncover insights into the developmental trajectories that shape language learning.
 
-Overview paragraph
+This study leverages a dataset derived from the Wordbank database, which compiles data from the MacArthur-Bates Communicative Development Inventories (CDI), to examine vocabulary development in children aged 16–30 months.
+A logistic regression model was employed to predict the likelihood of achieving a high vocabulary score based on variables such as age, norming status, and lexical categories.
+The model provides a framework for exploring how these predictors interact to shape children’s vocabulary acquisition.
 
-Estimand paragraph
+Key findings from this analysis reveal that vocabulary acquisition follows distinct trajectories across different word categories, with age being the strongest predictor of vocabulary growth.
+Furthermore, categories such as “Living Things” and “Function Words” show consistent acquisition patterns, while others, like “Sensory Words,” exhibit greater variability.
+These results contribute to a deeper understanding of the linguistic milestones that mark early childhood and highlight areas for further exploration in language development research.
 
-Results paragraph
+The remainder of this paper is structured as follows: @sec-data1 describes the dataset, highlighting the variables used and their distributions.
+@sec-model1 provides an overview of the logistic regression model and its assumptions.
+@sec-result presents the results, including variability in vocabulary acquisition, the role of predictors, and predicted probabilities.
+Finally, @sec-discussion discusses the implications of the findings and outlines limitations and directions for future research.
+The data used in this study all comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR.
 
-Why it matters paragraph
-
-The remainder of this paper is structured as follows.
-@sec-data descirbe the dataset in detail, and shows the distribution of variables.
-The data used in this study comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR
-
-# Data {#sec-data}
+# Data {#sec-data1}
 
@@
 
-To classify individuals, a threshold value of 350 is applied to `prod_comp_mean`:
-- Individuals with `prod_comp_mean > 350` are classified as having a high vocabulary score (outcome = 1).
-- Those with `prod_comp_mean <= 350` are classified as not having a high vocabulary score (outcome = 0).
+To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
+- Individuals with `prod_comp_mean > 300` are classified as having a high vocabulary score (outcome = 1).
+- Those with `prod_comp_mean <= 300` are classified as not having a high vocabulary score (outcome = 0).
 
 This approach ensures that both receptive (comprehension) and expressive (production) skills are considered in defining advanced vocabulary.
-The threshold of 350 was chosen based on exploratory analysis of the dataset, reflecting a meaningful distinction between individuals with high and low vocabulary abilities.
+The threshold of 300 was chosen based on exploratory analysis of the dataset, reflecting a meaningful distinction between individuals with high and low vocabulary abilities.
 The High Vocabulary Score serves as the dependent variable in the following data analysis part.
@@
 
-Norming status categorizes children into two groups: those included in the norming group and those who are not.
-The norming group represents a standardized sample used as a benchmark for assessing vocabulary development, providing a reference point for evaluating other children in the dataset.
-This distinction is important for ensuring the validity and reliability of comparisons in the analysis.
+The is_norming variable categorizes children into two groups: the norming group and the non-norming group.
+The norming group serves as a standardized sample, providing a benchmark for assessing vocabulary development and enabling reliable comparisons across the dataset.
+This group is essential for ensuring the validity of the analysis by offering a consistent reference point for evaluating individual and group-level differences in vocabulary acquisition.
 
-The bar plot illustrates the distribution of children based on their norming status.
-The dataset is predominantly composed of non-norming children, with only a small fraction representing the norming group.
-This imbalance reflects the practical challenges of including a broad and representative norming sample in large-scale vocabulary assessments.
-While the non-norming group provides diverse data for analysis, the norming group offers a crucial baseline for calibrating and interpreting the results.
-The distribution highlights the need to account for this imbalance in the analysis to avoid potential biases and ensure robust conclusions.
+The dataset is predominantly composed of non-norming children, with only a small proportion belonging to the norming group.
+This imbalance reflects the inherent challenges of including a sufficiently broad and representative norming sample in large-scale assessments.
+While the non-norming group captures diverse linguistic and demographic contexts, the norming group remains vital for calibrating and interpreting vocabulary development patterns, particularly in early language acquisition studies.
+@tbl-norming illustrates this distribution, highlighting the predominance of non-norming children in the dataset.
+
+The is_norming variable is included in the logistic regression model to account for potential systematic differences between the norming and non-norming groups.
+As the norming group represents a standardized sample used as a benchmark for vocabulary development, its inclusion ensures that variations in the likelihood of achieving a high vocabulary score are not confounded by differences in group composition or assessment protocols.
+It allows the model to estimate whether membership in the norming group significantly influences vocabulary acquisition outcomes.
+This is particularly important given the inherent imbalance in the dataset, where the non-norming group constitutes the majority.
+By controlling for is_norming, the model can differentiate between developmental patterns attributable to broader population diversity versus those arising from the structured selection of the norming sample.
 
@@
 #| echo: false
-#| label: fig-norming
-#| fig-cap: The dataset is primarily composed of non-norming children, with a smaller subset belonging to the norming group, serving as a standardized benchmark for assessing vocabulary development
+#| label: tbl-norming
+#| tbl-cap: The dataset is primarily composed of non-norming children, with a smaller subset belonging to the norming group, serving as a standardized benchmark for assessing vocabulary development
 #| warning: false
@@
   summary_isnorming_df,
-  col.names = c("Norming Status", "Count"),  # Set column names
-  caption = "Summary of Norming Status in the Dataset"
-)
+  col.names = c("Norming Status", "Count"),  
+  caption = "Summary of Norming Status in the Dataset")
 ```
 
-# Model
+# Model {#sec-model1}
 
@@
 By examining key demographic and linguistic predictors, we aim to identify how characteristics like age, norming status, and word categories influence vocabulary development.
-The dependent variable, high_vocabulary, is a binary outcome indicating whether a child’s average production and comprehension score (denoted as prod_comp_mean) exceeds 350.
+The dependent variable, high_vocabulary, is a binary outcome indicating whether a child’s average production and comprehension score (denoted as prod_comp_mean) exceeds 300.
 This threshold was chosen to distinguish children with relatively advanced vocabulary levels.
-More background details and diagnostics are included in [Appendix-@sec-model].
+More background details and diagnostics are included in [Appendix-@sec-model2].
 
@@
 
-- High Vocabulary: A binary indicator where 1 represents a high vocabulary score (combined comprehension and production > 350), and 0 otherwise.
+- High Vocabulary: A binary indicator where 1 represents a high vocabulary score (combined comprehension and production > 300), and 0 otherwise.
 
@@
 - Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary) and the independent variables.
-For example, the standardized age variable (age_scaled) assumes that for every one standard deviation increase in age, the log-odds of achieving a high vocabulary score increase by a constant amount.
-Standardizing age ensures that the variable is centered and scaled, making it easier to meet this linearity assumption and interpret its effect across the dataset.
-- Independent Observations: The model assumes that all data points are independent.
-This assumption holds because each observation represents data from a unique child, with no repeated measurements for the same individual.
-For instance, there are no longitudinal observations or nested data structures (e.g., children grouped by classrooms or schools) that could violate independence.
-If dependence were present, a more complex model like a mixed-effects logistic regression would be necessary.
-- Categorical Variable Encoding: The broad_category variable, which includes categories such as “Adjectives,” “Verbs,” and “Living Things,” was encoded using sum contrasts.
-This approach ensures that the coefficients for each category represent the deviation of that category’s effect from the overall mean effect across all categories.
-For example, the coefficient for “Verbs” indicates how the log odds of achieving a high vocabulary differ for “Verbs” compared to the average effect of all other categories.
-Sum contrasts are particularly useful for understanding relative effects and ensure that the intercept reflects the overall mean effect when all predictors are at their reference or average levels.
+For example, the standardized age variable (age_scaled) reflects that for every one standard deviation increase in age, the log-odds of achieving a high vocabulary score change by a constant amount.
+Standardizing age ensures that the variable is centered around zero and scaled to have a standard deviation of one, which not only aids interpretability but also supports the linearity assumption across the dataset.
+- Independence of Observations: Each observation in the dataset is assumed to be independent.
+This assumption holds because each data point corresponds to a unique child, with no repeated measures or nested structures, such as clustering by classrooms or schools.
+If there were dependencies (e.g., repeated measures for the same child or hierarchical data), a mixed-effects logistic regression or another hierarchical model would be required to account for these correlations.
+- Categorical Variable Encoding: The broad_category variable, which classifies words into categories such as “Adjectives,” “Verbs,” and “Living Things,” is encoded using sum contrasts.
+This encoding ensures that the coefficients for each category represent its deviation from the overall mean effect across all categories.
+For example, the coefficient for “Verbs” reflects the difference in log-odds of achieving a high vocabulary score for words in the “Verbs” category compared to the average log-odds across all categories.
+Using sum contrasts ensures that the model’s intercept represents the overall mean effect when all predictors are at their reference or average levels, facilitating meaningful interpretation of the model coefficients.
+- Binary Nature of the Outcome: The dependent variable, high_vocabulary, is binary (1 = high vocabulary, 0 = not high vocabulary).
+Logistic regression is specifically designed for binary outcomes, and the binomial distribution of the data supports the appropriateness of the logistic framework.
+-No Perfect Multicollinearity: The model assumes that the predictors are not perfectly correlated with one another.
+High multicollinearity could distort coefficient estimates, making it difficult to interpret the unique contribution of each variable.
+Steps such as standardizing continuous predictors and carefully encoding categorical variables mitigate this risk.
 
@@
 
-# Results
+# Results {#sec-result}
 
@@
 
-## Vocabulary Size Change by Age
+## Median Vocabulary Size Change by Age
 
@@
 #| label: fig-median
-#| fig-cap: "Illustrating central tendencies in comprehension vocabulary scores"
+#| fig-cap: Illustrating central tendencies in comprehension vocabulary scores based on resampled data
 #| warning: false
@@
 
-analysis_data_1<- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
+analysis_data_1 <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
 
-# Calculate the median comprehension for each age
-median_data <- analysis_data_1 %>%
+# Resample the dataset to include 100 observations
+set.seed(666)  # For reproducibility
+sampled_data <-  analysis_data_1 %>%
+  sample_n(800)
+
+# Calculate the median comprehension for each age in the sampled dataset
+median_data <- sampled_data %>%
   group_by(age) %>%
@@
 ggplot() +
-  geom_jitter(data = analysis_data_1, aes(x = age, y = prod_comp_mean), 
-              alpha = 0.2, size = 1, color = "orange") +  # Scatter points for individual data
+  geom_jitter(data = sampled_data, aes(x = age, y = prod_comp_mean), 
+              alpha = 0.7, size = 2, color = "orange") +  # Scatter points for sampled data
   geom_line(data = median_data, aes(x = age, y = median_prod_comp_mean), 
-            color = "blue", size = 1.5, linetype = "solid") +  # Line for median comprehension
+            color = "blue", size = 1.2, linetype = "solid") +  # Line for median comprehension
   labs(
-    title = "Median Comprehension Vocabulary by Age", x = "Age (Months)",
+    title = "Median Comprehension Vocabulary by Age (Resampled)", 
+    x = "Age (Months)",
     y = "Median of the Vocabulary Level") +
   theme_minimal(base_size = 14) +
-  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
-    plot.subtitle = element_text(size = 12, hjust = 0.5),
+  theme(
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
     axis.title = element_text(size = 12),
@@
     panel.grid.major = element_line(color = "grey85"),
-    panel.grid.minor = element_blank()) +
+    panel.grid.minor = element_blank()
+  ) +
   scale_x_continuous(breaks = seq(15, 30, 3))
@@
 ## Prediciton for the Probability of High Vocabulary Level
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
-
-# Preprocess test data
-analysis_data_test <- analysis_data_test %>%
-  mutate(prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
-    high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0),  # Binary target variable
-    age_scaled = scale(age)  # Standardize age
-  ) %>%
-  drop_na(prod_comp_mean)  # Drop rows with NA in prod_comp_mean
-analysis_data_test$age_scaled <- as.numeric(analysis_data_test$age_scaled)
-
-# Predict probabilities for test data
-predicted_probabilities <- predict(logistic_model, newdata = analysis_data_test, type = "response")
-
-# Predict classes based on probabilities
-predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
-
-# Create a confusion matrix
-confusion_matrix <- table(
-  Actual = analysis_data_test$high_vocabulary,
-  Predicted = predicted_classes
-)
-print(confusion_matrix)
-
-accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
-print(paste("Accuracy:", accuracy))
-```
 
@@
 # Add predicted probabilities to the test dataset
+analysis_data_test<- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
 analysis_data_test <- analysis_data_test %>% mutate(predicted_prob = predict(logistic_model, newdata = analysis_data_test, type = "response"))
@@
 age_broad_category_predictions <- analysis_data_test %>%
-  group_by(age, broad_category) %>%
+  group_by(age, broad_category, is_norming) %>%
   summarize(mean_predicted_prob = mean(predicted_prob, na.rm = TRUE), .groups = "drop")
@@
 
+norming_category_summary <- age_broad_category_predictions %>%
+  group_by(is_norming) %>%
+  summarize(overall_mean_prob = mean(mean_predicted_prob, na.rm = TRUE))
+
+age_broad_category_predictions
+norming_category_summary
 ```
 
-## Distribution of Predicted Probabilities by Broad Category
+## Predicted Probabilities by Age
 
-The @fig-prediction illustrates the distribution of predicted probabilities for high vocabulary levels across various broad categories.
-The density curves show distinct peaks and variability, reflecting the differences in how well each category predicts high vocabulary.
-Notably, “Sensory Words” and “Objects” exhibit higher density in the middle range of predicted probabilities, suggesting moderate association with high vocabulary.
-In contrast, categories like “Activities” and “Function Words” have wider, flatter distributions, indicating greater uncertainty or diversity in prediction.
-This variability highlights the nuanced role of word categories in predicting high vocabulary acquisition.
-Further analysis could explore why certain categories contribute more consistently to predictions than others.
-The density plot highlights distinct patterns in predicted probabilities, with nouns and verbs showing higher peaks, indicating a stronger likelihood of high vocabulary acquisition in these categories.
-This suggests that broad categories contribute differently to vocabulary development, with nouns and verbs potentially playing a more significant role.
+@fig-agepercent illustrates the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) overlaid to highlight the variability in predictions.
+The scatterplot points represent individual predicted probabilities, while the percentile lines depict the progression of predictions across age groups.
 
-Each facet represents a unique category, with the x-axis denoting the predicted probability of high vocabulary and the y-axis representing density.
+The median predicted probability (50th percentile, solid blue line) steadily increases with age, reflecting the model’s growing confidence in high vocabulary acquisition as children get older.
+The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable rise after 20 months, suggesting greater variability in predictions among younger children.
+In contrast, the 75th and 90th percentiles (dashed and dotted purple/red lines) start higher and climb more sharply, indicating that some children exhibit advanced vocabulary skills even at younger ages.
 
-The plots highlight distinct patterns in the distributions of predicted probabilities for each category.
-Categories such as Adjectives and Function Words exhibit broader distributions, suggesting greater variability in the likelihood of achieving a high vocabulary score within these categories.
-In contrast, categories like Sensory Words and Activities show relatively concentrated distributions near lower probabilities, indicating more consistent outcomes within those lexical groups.
-
-The graph underscores the heterogeneity of vocabulary acquisition probabilities across different lexical domains.
-Categories associated with more concrete or commonly encountered words, such as Objects and Living Things, tend to exhibit higher probabilities for vocabulary mastery, reflecting their early and frequent usage in language development.
-Conversely, categories like Sensory Words show lower probabilities, suggesting that they are acquired less consistently or later in development.
-This visualization offers insight into the differential patterns of lexical acquisition, emphasizing the varied developmental trajectories across word types.
+Overall, the increasing spread between the percentile lines with age highlights a broader range of vocabulary acquisition patterns as children develop.
+This visualization underscores the model’s ability to account for individual differences in learning trajectories while confirming the strong association between age and predicted probability of high vocabulary.
 
 ```{r}
+#| echo: false
+#| eval: true
+#| label: fig-agepercent
+#| warning: false
+
+# Resample the dataset to include 800 observations
+set.seed(666)  # For reproducibility
+sampled_data_test <- analysis_data_test %>%
+  sample_n(3000)
+
+# Define age ranges
+sampled_data_test <- sampled_data_test %>%
+  mutate(age_range = cut(age, breaks = seq(min(age), max(age), by = 3), include.lowest = TRUE, right = FALSE))
+
+# Calculate percentiles for each age range
+percentile_data <- sampled_data_test %>%
+  group_by(age_range) %>%
+  summarize(
+    p10 = quantile(predicted_prob, 0.10, na.rm = TRUE),
+    p25 = quantile(predicted_prob, 0.25, na.rm = TRUE),
+    p50 = quantile(predicted_prob, 0.50, na.rm = TRUE),  # Median
+    p75 = quantile(predicted_prob, 0.75, na.rm = TRUE),
+    p90 = quantile(predicted_prob, 0.90, na.rm = TRUE),
+    age_mid = mean(as.numeric(as.character(age)), na.rm = TRUE), # Midpoint of age range
+    .groups = "drop"
+  )
+
+# Create the scatterplot with percentile lines
+ggplot(sampled_data_test, aes(x = age, y = predicted_prob)) +
+  # Scatterplot of individual predictions
+  geom_point(alpha = 0.5, size = 2, color = "#FF9933") +
+  # Percentile lines
+  geom_line(data = percentile_data, aes(x = age_mid, y = p10), 
+            color = "#FF6666", linetype = "dashed", size = 1.2, show.legend = TRUE) +  # 10th percentile
+  geom_line(data = percentile_data, aes(x = age_mid, y = p25), 
+            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 25th percentile
+  geom_line(data = percentile_data, aes(x = age_mid, y = p50), 
+            color = "#3399FF", linetype = "solid", size = 1.5) +  # 50th percentile (Median)
+  geom_line(data = percentile_data, aes(x = age_mid, y = p75), 
+            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 75th percentile
+  geom_line(data = percentile_data, aes(x = age_mid, y = p90), 
+            color = "#FF6666", linetype = "dashed", size = 1.2) +  # 90th percentile
+  # Add labels and theme
+  labs(
+    title = "Predicted Probabilities by Age with Percentiles", 
+    x = "Age (Months)",
+    y = "Predicted Probability",
+    caption = "Percentile lines show 10th, 25th, 50th, 75th, and 90th percentiles."
+  ) +
+  scale_x_continuous(
+    breaks = seq(min(sampled_data_test$age), max(sampled_data_test$age), by = 3),
+    labels = function(x) paste(x, "mo")  # Add "mo" for months
+  ) +
+  scale_y_continuous(
+    limits = c(0, 1), breaks = seq(0, 1, 0.1),
+    labels = scales::percent_format()  # Format as percentages
+  ) +
+  theme_minimal(base_size = 14) +
+  theme(
+    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
+    plot.caption = element_text(size = 10, hjust = 0.5, face = "italic"),
+    axis.title = element_text(size = 14, face = "bold"),
+    axis.text = element_text(size = 12),
+    panel.grid.major = element_line(color = "grey85"),
+    panel.grid.minor = element_blank()) +
+  annotate("segment", x = max(sampled_data_test$age) - 4, xend = max(sampled_data_test$age) - 2, y = 0.9, yend = 0.9, 
+           color = "#FF6666", linetype = "dashed", size = 1.2)
+```
+
+## Distribution of Predicted Probabilities by Category
+
+The distribution of predicted probabilities for achieving high vocabulary varies across word categories, as shown in @fig-prediction.
+Categories such as Function Words, Living Things, and Objects demonstrate a concentration of predicted probabilities closer to 0.5 or above, indicating the model’s moderate to high confidence in associating these categories with vocabulary acquisition.
+These categories likely reflect simpler or more frequently used words that children acquire earlier.
+In contrast, categories like Sensory Words and Verbs exhibit a higher density of lower predicted probabilities, suggesting that these words are more challenging for children to acquire or are less common in early vocabularies.
+
+Some categories, such as Adjectives and Places, display broader distributions of predicted probabilities, spanning from low to high values.
+This variability may reflect differences in children’s exposure to or use of these word types, as well as greater heterogeneity in language environments.
+Overall, these findings provide insights into how the model predicts vocabulary acquisition and highlight the influence of word category.
+Categories with more focused distributions (e.g., Function Words) suggest consistent acquisition patterns, while those with broader spreads (e.g., Verbs) point to areas where model predictions are less certain and could benefit from further refinement.
+
+```{r, width = 10,height=10 }
 #| echo: false
@@
 
-# Assuming `analysis_data_test` contains the predicted probabilities and broad categories
-
 ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
-  geom_density(alpha = 0.7, color = "black") +  # Density plot with transparency
-  facet_wrap(~ broad_category, scales = "fixed") +  # Same y-axis range for all facets
+  geom_density(alpha = 0.7, color = "black", size = 0.3) +  # Smooth lines for density plots
+  facet_wrap(~ broad_category, scales = "fixed", ncol = 3) +  # Arrange facets in 3 columns
   labs(
@@
     x = "Predicted Probability",
-    y = "Density",
-    fill = "Broad Category") +
+    y = "Density"
+  ) +
+  scale_fill_brewer(palette = "Set3") +  # Use a more vibrant color palette
   theme_minimal(base_size = 14) +
   theme(
-    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
-    axis.title = element_text(size = 14),
-    axis.text = element_text(size = 12),
-    strip.text = element_text(size = 12, face = "bold"),  # Styling facet labels
+    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
+    axis.title = element_text(size = 14, face = "bold"),
+    axis.text = element_text(size = 9),
+    strip.text = element_text(size = 12, face = "bold", color = "darkblue"),  # Highlight facet labels
+    panel.spacing = unit(2, "lines"),  # Increase space between facets
     legend.position = "none",  # Remove legend
-    panel.grid.major = element_line(color = "gray85"),
-    panel.grid.minor = element_blank())
+    panel.grid.major = element_line(color = "gray90"),
+    panel.grid.minor = element_blank()
+  )
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-## Vocabulary Development Patterns {#sec-firstpoint}
+## Age and Developmental Trajectories
+
+The findings emphasize age as the most influential predictor of vocabulary acquisition.
+The sharp increase in predicted probabilities as children grow older aligns with expectations, as the developmental window between 18 and 30 months is crucial for rapid language learning.
+The steady upward trend in median vocabulary probabilities suggests that this period is marked by consistent growth in both production and comprehension.
+However, the growing variability between the upper and lower percentiles reveals the diverse learning trajectories among children.
+While some children quickly advance, others may require more time, reflecting the role of environmental factors and individual differences in cognitive development.
+
+Interestingly, the plateauing trend observed in certain facets of comprehension suggests that as children master foundational vocabulary, their acquisition of more complex or context-specific words becomes less predictable.
+This shift highlights the importance of contextual factors such as exposure to rich language environments and interactions with caregivers.
+
+## Word Categories: Foundation and Complexity
+
+The analysis of word categories revealed striking differences in acquisition patterns.
+Categories such as “Function Words,” “Living Things,” and “Objects” displayed higher predicted probabilities, likely because these words are more frequently encountered in everyday interactions and form the backbone of early communication.
+In contrast, categories like “Verbs” and “Sensory Words” demonstrated lower probabilities and wider variability, suggesting they are more challenging for young learners.
+These words often require greater contextual understanding or abstract thinking, which may develop later in the language acquisition process.
+
+The variability in word categories sheds light on how children’s experiences shape their vocabulary.
+For instance, categories with higher predicted probabilities may reflect a shared baseline of language exposure across children, whereas the broader spread in categories like “Adjectives” or “Sensory Words” points to the influence of specific cultural or environmental factors.
 
@@
 
-## Role of Predictors in Vocabulary Growth
+## Broader Implications for Language Development
 
-Key predictors, including age, norming status, and broad lexical categories, significantly influenced vocabulary acquisition.
-Age, as expected, showed the strongest correlation, underlining its role as a primary determinant of language development.
-Children from the norming group exhibited higher vocabulary scores, possibly due to more structured linguistic environments or assessment settings.
+The findings of this study offer practical applications for early education and language interventions by highlighting specific word categories that are more challenging for children to acquire.
+For instance, verbs and sensory words, which often require greater contextual understanding, could be emphasized in storytelling, play, and structured educational settings to help bridge the gap in these more complex categories.
+Tailored approaches in language development programs can address individual learning needs, with children demonstrating slower growth in foundational vocabulary benefiting from increased exposure to high-frequency words.
+Conversely, those excelling in comprehension may gain from activities introducing less common or context-specific terms, thereby expanding their linguistic repertoire.
 
-Lexical categories revealed nuanced contributions to vocabulary development.
-Categories such as objects and verbs were associated with higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
-Conversely, categories like sensory words and adjectives showed less consistent trends, indicating their later emergence in the linguistic repertoire.
-
-## Third discussion point
+Key predictors, including age, norming status, and lexical categories, significantly influenced vocabulary acquisition.
+Age was the strongest determinant of language growth, with older children showing higher vocabulary scores, consistent with developmental expectations.
+Additionally, children in the norming group exhibited superior vocabulary acquisition, potentially due to the structured linguistic environments associated with their assessment.
+Lexical categories provided further insights into vocabulary development, with objects and verbs showing higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
+In contrast, sensory words and adjectives displayed less consistent patterns, suggesting their acquisition may occur later as children’s linguistic abilities become more sophisticated.
 
@@
 
-While the findings offer valuable insights, several limitations should be acknowledged.
+264 typically developing Chinese children (145 boys and 119 girls) were included from a longitudinal study of languageWhile the findings offer valuable insights, several limitations should be acknowledged.
 First, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
@@
 Second, the cross-sectional nature of the data limits the ability to track individual developmental trajectories.
-Longitudinal studies are needed to capture within-child variability and the dynamics of vocabulary growth over time.
+Longitudinal studies are needed to capture within-child variability and the dynamics of vocabulary growth over time.Familial factors and reading or language related cognitive skills were found to be associated with these developmental subgroups.
 
@@
 
-# Appendix {#-secdata}
+# Appendix {#sec-data2}
 
-## Additional data details
+## This criterion is linked to a Learning OutcomeSurveys, sampling, and observational data appendix
+Please include an appendix where you focus on some aspect of surveys, sampling or observational data, related to your paper.
+This should be an in-depth exploration, akin to the "idealized methodology/survey/pollster methodology" sections of Paper 2.
+Some aspect of this is likely covered in the Measurement sub-section of your Data section, but this would be much more detailed, and might include aspects like simulation and linkages to the literature, among other aspects.
 
@@
 
-# Model details {#sec-model}
+# Model details {#sec-model2}
 
@@
 #| eval: true
-#| label: tbl-logistic
-#| tbl-cap: model summary table
+#| label: tbl-modelsummary
 #| warning: false
@@
 modelsummary(logistic_model)
+```
 
-```
+\newpage
 
@@
 
-### Accuracy
+### Confusion Matrix
+```{r}
+#| echo: false
+analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
+
+# Preprocess test data
+analysis_data_test <- analysis_data_test %>%
+  mutate(prod_comp_mean = (production + comprehension) / 2,  # Average of production and comprehension
+    high_vocabulary = ifelse(prod_comp_mean > 300, 1, 0),  # Binary target variable
+    age_scaled = scale(age)  # Standardize age
+  ) %>%
+  drop_na(prod_comp_mean)  # Drop rows with NA in prod_comp_mean
+analysis_data_test$age_scaled <- as.numeric(analysis_data_test$age_scaled)
+
+# Predict probabilities for test data
+predicted_probabilities <- predict(logistic_model, newdata = analysis_data_test, type = "response")
+
+# Predict classes based on probabilities
+predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
+
+# Create a confusion matrix
+confusion_matrix <- table(
+  Actual = analysis_data_test$high_vocabulary,
+  Predicted = predicted_classes
+)
+print(confusion_matrix)
+
+# Compute accuracy, sensitivity, specificity, and precision
+accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
+sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])  # True Positives / Actual Positives
+specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])  # True Negatives / Actual Negatives
+precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])  # True Positives / Predicted Positives
+
+# Create a dataframe for metrics
+metrics <- data.frame(
+  Metric = c("Accuracy", "Sensitivity (Recall)", "Specificity", "Precision"),
+  Value = c(accuracy, sensitivity, specificity, precision)
+)
+
+# Display the table using kable
+kable(metrics, col.names = c("Metric", "Value"), format = "markdown", digits = 2)
+```
+
+### ROC Curve and AUC
+```{r}
+#| echo: false
+library(pROC)
+roc_curve <- roc(analysis_data_test$high_vocabulary, predicted_probabilities)
+auc_value <- auc(roc_curve)
+cat(paste("AUC:", auc_value, "\n"))
+plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
+```
+
+### Precision-Recall Curve
+```{r}
+#| echo: false
+# Precision-Recall Curve (optional, useful for imbalanced data)
+#install.packages("PRROC")
+library(PRROC)
+pr_curve <- pr.curve(scores.class0 = predicted_probabilities[analysis_data_test$high_vocabulary == 1],
+                     scores.class1 = predicted_probabilities[analysis_data_test$high_vocabulary == 0],
+                     curve = TRUE)
+plot(pr_curve, main = "Precision-Recall Curve")
+```
+### Binned Residual Plot
 ```{r}
@@
 
-analysis_data_train <- read_parquet(here::here("data/02-analysis_data/train_data.parquet"))
-
-# Predicted probabilities and classes
-predicted_classes <- ifelse(fitted(logistic_model) > 0.5, 1, 0)
-confusion_matrix <- table(Predicted = predicted_classes, Actual = analysis_data_train$high_vocabulary)
-
-# Print confusion matrix and accuracy
-print(confusion_matrix)
-accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
-print(paste("Accuracy:", round(accuracy, 3)))
+binnedplot(fitted(logistic_model), residuals(logistic_model, type = "response"),
+           xlab = "Fitted Probabilities", ylab = "Average Residual",
+           main = "Binned Residual Plot")
 ```
 
+### Predicted Probabilities by Age
 ```{r}
@@
 # Improved Panel (a): Scatterplot of Predicted Probabilities
-ggplot(analysis_data_test, aes(x = age_scaled, y = predicted_prob, color = high_vocabulary)) +
+ggplot(analysis_data_test, aes(x = age, y = predicted_prob, color = high_vocabulary)) +
   geom_point(alpha = 0.6, size = 3, shape = 16, position = position_jitter(width = 0.05, height = 0.02)) +
   labs(
-    title = "Panel (a): Predicted Probabilities by Age (Scaled)",
-    x = "Age (Scaled)",
+    title = "Predicted Probabilities by Age",
+    x = "Age",
     y = "Predicted Probability of High Vocabulary",
@@
 
-### Binned Residual Plot
-```{r}
-#| echo: false
-
-binnedplot(fitted(logistic_model), residuals(logistic_model, type = "response"),
-           xlab = "Fitted Probabilities", ylab = "Average Residual",
-           main = "Binned Residual Plot")
-```
-
-\newpage
-
 # Acknowledgements
@@
 The analysis was conducted using a suite of packages from @citeR, which offered robust functionality for data manipulation, visualization, and storage.
-We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
+We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @arm, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
 Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files.
@@
 
+\newpage
+
 # References

### 2024-11-24 scripts/03-clean_data.R

@@ -54,7 +54,7 @@ cleaned_data <- cleaned_data %>%
 
 # Create the high_vocabulary variable based on prod_comp_mean
 cleaned_data <- cleaned_data %>%
-  mutate(high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0))
+  mutate(high_vocabulary = ifelse(prod_comp_mean > 300, 1, 0))
 
 # Omit rows with NA values
 cleaned_data <- cleaned_data %>% 

### 2024-11-24 scripts/06-model_data.R

@@ -39,7 +39,7 @@ analysis_data_test <- testing(split)
 # Combine production and comprehension into a new variable
 analysis_data_train <- analysis_data_train %>%
   mutate(prod_comp_mean = (production + comprehension)/2,  # Average of production and comprehension
-    high_vocabulary = ifelse(prod_comp_mean > 350, 1, 0)  # Binary target variable
+    high_vocabulary = ifelse(prod_comp_mean > 300, 1, 0)  # Binary target variable
   ) %>%
   drop_na(prod_comp_mean)  # Drop rows where prod_comp_mean is NA
 

### 2024-11-25 paper/paper.qmd
Paper draft over.

@@
 ---
-title: "Exploring Vocabulary Acquisition in Children"
-subtitle: "Predicted Vocabulary Growth in Children: Higher Probabilities for Common Words and Greater Variability in Complex Categories"
+title: "Pathways to Early Vocabulary Acquisition"
+subtitle: "Predicting English Language Development Trajectories in Children Aged 16–30 Months"
 author: Yongqi Liu
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study examines vocabulary acquisition in children aged 16–30 months using logistic regression models and data from the MacArthur-Bates Communicative Development Inventories.
+Age is the strongest predictor, with older children showing higher vocabulary proficiency, while foundational word categories like Function Words are acquired more consistently than complex ones such as Sensory Words.
+Children in norming groups outperform non-norming peers, highlighting the impact of structured linguistic environments.
+The findings support targeted interventions to address variability in complex vocabulary acquisition and promote balanced language development."
 format: pdf
@@
 
-# Introduction
+# Introduction {#sec-intro}
+
 Understanding how children acquire vocabulary is a cornerstone of developmental linguistics and early childhood education.
@@
 The interplay of factors such as age, linguistic environments, and word categories influences children’s vocabulary acquisition in complex and dynamic ways.
-By analyzing patterns and predictors of vocabulary growth, researchers can uncover insights into the developmental trajectories that shape language learning.
+By analyzing patterns and predictors of vocabulary growth, researchers can illuminate the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities.
 
-This study leverages a dataset derived from the Wordbank database, which compiles data from the MacArthur-Bates Communicative Development Inventories (CDI), to examine vocabulary development in children aged 16–30 months.
-A logistic regression model was employed to predict the likelihood of achieving a high vocabulary score based on variables such as age, norming status, and lexical categories.
-The model provides a framework for exploring how these predictors interact to shape children’s vocabulary acquisition.
+This study utilizes a dataset from the @wordbankr database, which aggregates data from the MacArthur-Bates Communicative Development Inventories (CDI), to investigate vocabulary development in children aged 16–30 months.
+By employing a logistic regression model, the research estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories.
+This model serves as a tool to explore how these factors collectively influence children’s vocabulary acquisition trajectories.
 
-Key findings from this analysis reveal that vocabulary acquisition follows distinct trajectories across different word categories, with age being the strongest predictor of vocabulary growth.
-Furthermore, categories such as “Living Things” and “Function Words” show consistent acquisition patterns, while others, like “Sensory Words,” exhibit greater variability.
-These results contribute to a deeper understanding of the linguistic milestones that mark early childhood and highlight areas for further exploration in language development research.
+The estimand of interest in this study is the conditional probability that a child possesses advanced vocabulary proficiency, defined as a high combined comprehension and production score, given specific predictors.
+These predictors include the child’s age in months, their norming status (whether they belong to a standardized sample used for benchmarking), and the lexical category of the words (e.g., Function Words, Verbs, Adjectives).
+By employing a logistic regression model, the research estimates how these variables individually and interactively influence the likelihood of achieving advanced vocabulary levels.
+The model demonstrates developmental trends, such as how median vocabulary proficiency increases with age, and reveals how specific word categories contribute uniquely to language growth.
+For example, foundational categories like Function Words and Objects may show consistently higher probabilities of acquisition, while more abstract categories, such as Verbs and Sensory Words, exhibit greater variability due to their complexity or context-specific nature.
 
-The remainder of this paper is structured as follows: @sec-data1 describes the dataset, highlighting the variables used and their distributions.
+Key findings reveal distinct trajectories of vocabulary acquisition across different word categories, with age emerging as the strongest predictor of vocabulary growth.
+The analysis demonstrates that as children age, their likelihood of achieving high vocabulary scores consistently increases, reflecting the natural progression of language development.
+Categories such as “Living Things” and “Function Words” display stable and high predicted probabilities, likely due to their frequent occurrence in everyday communication and their role in forming early linguistic structures.
+Conversely, categories such as “Sensory Words” and “Adjectives” show more variability, highlighting the challenges posed by their later acquisition and the contextual understanding they require.
+
+Additionally, children in the norming group, who are assessed in structured linguistic environments, consistently achieve higher vocabulary scores compared to non-norming children.
+This difference underscores the influence of systematic linguistic exposure in shaping vocabulary development.
+The results not only provide a nuanced understanding of vocabulary acquisition but also emphasize the importance of targeted language interventions.
+They suggest that focusing on underdeveloped categories, such as Sensory Words and Adjectives, could help bridge gaps in vocabulary learning.
+Furthermore, these findings set the stage for future research into the broader contextual factors, such as socio-economic status and bilingualism, that may further explain variability in language acquisition trajectories.
+
+The remainder of this paper is structured as follows: @sec-intro provides an introduction and decribe the estimand for the paper.
+@sec-data1 describes the dataset, highlighting the variables used and their distributions.
 @sec-model1 provides an overview of the logistic regression model and its assumptions.
-@sec-result presents the results, including variability in vocabulary acquisition, the role of predictors, and predicted probabilities.
+@sec-result presents the visualized results, including variability in vocabulary acquisition, the role of predictors, and predicted probabilities.
 Finally, @sec-discussion discusses the implications of the findings and outlines limitations and directions for future research.
@@
 
-To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
+Based on @wendy, I looked into the normal vocabulary size change among the children aged 16 to 30 months. To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
 - Individuals with `prod_comp_mean > 300` are classified as having a high vocabulary score (outcome = 1).
@@
 
+\newpage
+
 ## Predictor variables
@@
 This uneven age distribution underscores the importance of age as a critical factor in analyzing vocabulary acquisition.
-While the high concentration of data at older ages enhances insights into advanced vocabulary development, it also necessitates caution in generalizing findings to underrepresented age groups.
+While the high concentration of data at older ages enhances the analysis into advanced vocabulary development, it also necessitates caution in generalizing findings to underrepresented age groups.
 This observation emphasizes the need to standardize age in statistical models to account for variability across different age groups.
@@
 
-### Broad Category
+### Lexical Category
 The words in the dataset were grouped into broad lexical categories to facilitate the analysis of vocabulary acquisition patterns.
@@
   geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
-  labs(title = "Reclassified Category Distribution", x = "Broad Category", y = "Count") +
+  labs(title = "Reclassified Category Distribution", x = "Lexical Category", y = "Count") +
   theme_minimal() +
@@
 - $\beta_2$: The effect of whether the individual belongs to the norming group
-- $\beta_3$, $\beta_4$, $\beta_5$, etc.: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).
+- $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).
 
@@
 
-- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary) and the independent variables.
-For example, the standardized age variable (age_scaled) reflects that for every one standard deviation increase in age, the log-odds of achieving a high vocabulary score change by a constant amount.
-Standardizing age ensures that the variable is centered around zero and scaled to have a standard deviation of one, which not only aids interpretability but also supports the linearity assumption across the dataset.
-- Independence of Observations: Each observation in the dataset is assumed to be independent.
-This assumption holds because each data point corresponds to a unique child, with no repeated measures or nested structures, such as clustering by classrooms or schools.
-If there were dependencies (e.g., repeated measures for the same child or hierarchical data), a mixed-effects logistic regression or another hierarchical model would be required to account for these correlations.
-- Categorical Variable Encoding: The broad_category variable, which classifies words into categories such as “Adjectives,” “Verbs,” and “Living Things,” is encoded using sum contrasts.
-This encoding ensures that the coefficients for each category represent its deviation from the overall mean effect across all categories.
-For example, the coefficient for “Verbs” reflects the difference in log-odds of achieving a high vocabulary score for words in the “Verbs” category compared to the average log-odds across all categories.
-Using sum contrasts ensures that the model’s intercept represents the overall mean effect when all predictors are at their reference or average levels, facilitating meaningful interpretation of the model coefficients.
+- Linearity of the Logit: The model assumes a linear relationship between the log-odds of the outcome (high vocabulary proficiency) and the independent variables.
+For instance, the standardized age variable (age\_scaled) ensures that for every one standard deviation increase in age, the log-odds of achieving high vocabulary change by a constant amount.
+Standardization centers the variable around zero and scales it to have a standard deviation of one, improving both interpretability and adherence to the linearity assumption.
+- Independence of Observations: Each observation corresponds to a unique child, and the data assumes no repeated measures or nested structures (e.g., grouping by classrooms or schools).
+This independence ensures the validity of the logistic regression framework.
+If dependencies, such as repeated measures or clustering, were present, a mixed-effects logistic regression or other hierarchical modeling techniques would be required.
+- Categorical Variable Encoding: The categorical variable broad_category (e.g., “Adjectives,” “Verbs,” “Living Things”) is encoded using sum contrasts.
+This ensures that each coefficient reflects the deviation of a given category from the overall mean effect across all categories.
+For example, the coefficient for “Verbs” represents the difference in log-odds for this category compared to the average log-odds across all categories.
+Sum contrasts further allow the model’s intercept to represent the overall mean effect when all predictors are at their reference or mean levels, facilitating clear interpretation.
 - Binary Nature of the Outcome: The dependent variable, high_vocabulary, is binary (1 = high vocabulary, 0 = not high vocabulary).
-Logistic regression is specifically designed for binary outcomes, and the binomial distribution of the data supports the appropriateness of the logistic framework.
--No Perfect Multicollinearity: The model assumes that the predictors are not perfectly correlated with one another.
-High multicollinearity could distort coefficient estimates, making it difficult to interpret the unique contribution of each variable.
-Steps such as standardizing continuous predictors and carefully encoding categorical variables mitigate this risk.
+Logistic regression is appropriate for modeling binary outcomes, as it assumes the binomial distribution of the data, which aligns with the structure of the outcome variable.
+- No Perfect Multicollinearity: The predictors are assumed to be non-perfectly correlated.
+High multicollinearity would lead to unreliable coefficient estimates and obscure the individual effects of predictors.
+Standardizing continuous predictors (e.g., age) and using appropriate encoding for categorical variables help minimize this risk and ensure stable model estimation.
 
@@
 
-The logistic regression coefficient ($\beta$) represents the change in the log-odds of the dependent variable (high vocabulary) for a one-unit change in the predictor variable, holding all other variables constant.
+The logistic regression coefficient ($\beta$) represent the change in the log-odds of achieving high vocabulary proficiency for a one-unit change in the respective predictor variable, holding all other variables constant.
 
-- Intercept ($\beta_0$): Represents the log-odds of high vocabulary when all predictors are at their reference or mean levels.
+- Intercept ($\beta_0$):  Represents the log-odds of high vocabulary proficiency when all predictors are at their reference levels.
 If $\beta_0$ > 0, the baseline odds of high vocabulary are greater than 50%.
+Refering to the modelsummary in the @sec-model2, we knwo that the baseline odds of high vocabulary proficiency are less than 50%.
+
 - Scaled Age ($\beta_1$): For each one standard deviation increase in age, the log-odds of high vocabulary increase by $\beta_1$.
 If $\beta_1 = 0.5$, then $\exp(0.5) \approx 1.65$, meaning the odds increase by 65% for every one standard deviation increase in age.
-- Norming Status ($\beta_2$): If a child belongs to the norming group, the log-odds of high vocabulary increase by $\beta_2$ compared to non-norming children.
-If $\beta_2 = 0.1$, then $\exp(0.1) \approx 1.11$, meaning being in the norming group increases the odds of high vocabulary by 11%.
-- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$, etc.): The coefficients for `broad_category` represent the difference in log-odds compared to the reference category (“Adjectives”).  
-  - $\beta_3$ (Function Words): A positive $\beta_3$ indicates higher odds of having a high vocabulary for function words compared to adjectives.
-For instance, if $\beta_3 = 0.002$, then $\exp(0.002) \approx 1.002$, meaning the odds of having a high vocabulary for function words are 0.2% higher than for adjectives.
-General Example for Broad Categories: If a coefficient $\beta_k = 0.01$, $\exp(0.01) \approx 1.01$, meaning the corresponding category increases the odds of having a high vocabulary by 1% compared to the reference category (Adjectives).
-Conversely, if $\beta_k = -0.01$, $\exp(-0.01) \approx 0.99$, indicating a 1% decrease in odds compared to the reference category.
+
+- Norming Status ($\beta_2$): Indicates the effect of belonging to the norming group, a positive $\beta_2$ suggests higher odds of high vocabulary compared to non-norming children.
+If $\beta_2 = 0.1$, then $\exp(0.1) \approx 1.11$, meaning norming group children have 22% higher odds of high vocabulary.
+
+- Broad Category ($\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$): The coefficients for the broad word categories (e.g., Function Words, Living Things) represent their effect on the log-odds of high vocabulary compared to the reference category (Adjectives).
+A positive $\beta_k$ indicates higher odds compared to Adjectives.
+For example, if $\beta_4$ = 0.3 (Function Words), then $\exp(0.3) \approx 1.35$, meaning Function Words increase the odds by 35% compared to Adjectives.
+A negative $\beta_k$ implies lower odds.
+For instance, if $\beta_7 = -0.008$ (Places), then $\exp(-0.008) \approx 0.992$ shows a minor decrease in odds for words in the Places category.
 
@@
 
-Logistic regression is widely used in predictive modeling for categorical outcomes due to its simplicity and robustness.
-It provides probabilities that are constrained between 0 and 1, ensuring meaningful interpretations for binary outcomes.
-Unlike other complex models, logistic regression allows for clear coefficient interpretation, offering insights into the magnitude and direction of predictor effects.
-For example, the odds ratios derived from logistic regression help explain how changes in variables like age or norming status influence the probability of high vocabulary acquisition.
+Logistic regression is well-suited for this study, given its ability to model binary outcomes like high vocabulary proficiency (1 = high vocabulary, 0 = not high vocabulary) while maintaining interpretability.
+By constraining predicted probabilities between 0 and 1, logistic regression ensures meaningful predictions.
+Its coefficients offer clear investigation into the magnitude and direction of effects, enabling an understanding of how predictors like age and norming status influence the likelihood of high vocabulary acquisition.
+For instance, odds ratios derived from the model allow straightforward interpretation of the impact of each variable, such as the increased likelihood of high vocabulary with a one-standard-deviation increase in age.
 
-Although advanced machine learning models like decision trees, random forests, or neural networks could be used, these methods often lack the interpretability of logistic regression.
-While these models might yield slightly better predictive performance, they are often considered “black boxes,” making it difficult to identify specific relationships between predictors and outcomes.
-Given the study’s focus on understanding developmental patterns rather than maximizing predictive accuracy, logistic regression is more appropriate.
+Although more advanced machine learning models like random forests or neural networks could provide slight improvements in predictive accuracy, these approaches lack the transparency needed to understand the underlying relationships between predictors and outcomes.
+Given this study’s focus on uncovering developmental patterns rather than solely optimizing prediction accuracy, logistic regression offers the necessary balance between interpretability and performance.
+Moreover, the dataset size and structure favor logistic regression, which is less prone to overfitting than more complex models that often require larger datasets to generalize effectively.
 
-Moreover, complex models require larger datasets to avoid overfitting and ensure generalizability, which might not be feasible given the sample size and the structure of the data in this study.
-Logistic regression strikes a balance between simplicity, interpretability, and predictive performance.
-
-The dataset was split into training and testing subsets to ensure model validation and reduce overfitting.
-This allows the model to generalize better to unseen data, providing a more reliable assessment of its predictive accuracy.
-Standardizing continuous predictors, such as age, enhances interpretability and ensures that variables are on a comparable scale, preventing dominance by predictors with larger numerical ranges.
-
-To ensure reproducibility, the preprocessed datasets and the trained model were saved.
-This practice facilitates verification of results and supports future analyses or extensions of the study.
-Overall, the logistic regression model offers a clear and interpretable framework for investigating vocabulary acquisition patterns, providing both explanatory power and practical insights.
+To ensure model robustness, the dataset was split into training and testing subsets to validate the model and minimize overfitting.
+Additionally, predictors like age were standardized to ensure comparability and prevent dominance by variables with larger numerical ranges.
+Categorical variables, such as word categories, were encoded using sum contrasts, allowing meaningful comparisons and ensuring that coefficients reflect deviations from the overall mean effect.
+These steps, combined with logistic regression’s simplicity and explanatory power, make it an optimal choice for investigating vocabulary acquisition patterns.
 
@@
 #| label: fig-median
-#| fig-cap: Illustrating central tendencies in comprehension vocabulary scores based on resampled data
+#| fig-cap: This figure illustrates the relationship between age (in months) and the median comprehension vocabulary level.
+Each orange point represents a resampled data point, while the blue line depicts the median vocabulary level at each age.
+The trend highlights a steady increase in vocabulary comprehension as children age, with noticeable variability across individual data points.
 #| warning: false
@@
   theme(
-    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
     axis.title = element_text(size = 12),
@@
 
-age_broad_category_predictions
-norming_category_summary
+kable(summary(age_broad_category_predictions))
 ```
@@
 @fig-agepercent illustrates the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) overlaid to highlight the variability in predictions.
-The scatterplot points represent individual predicted probabilities, while the percentile lines depict the progression of predictions across age groups.
+Based on the resample of the test data, the scatterplot points represent individual predicted probabilities, while the percentile lines depict the progression of predictions across age groups.
 
@@
 sampled_data_test <- analysis_data_test %>%
-  sample_n(3000)
+  sample_n(5000)
 
@@
     age_mid = mean(as.numeric(as.character(age)), na.rm = TRUE), # Midpoint of age range
-    .groups = "drop"
-  )
+    .groups = "drop")
 
@@
   theme(
-    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
     plot.caption = element_text(size = 10, hjust = 0.5, face = "italic"),
@@
 
-## Distribution of Predicted Probabilities by Category
+## Distribution of Predicted Probabilities by Word Category
 
 The distribution of predicted probabilities for achieving high vocabulary varies across word categories, as shown in @fig-prediction.
-Categories such as Function Words, Living Things, and Objects demonstrate a concentration of predicted probabilities closer to 0.5 or above, indicating the model’s moderate to high confidence in associating these categories with vocabulary acquisition.
-These categories likely reflect simpler or more frequently used words that children acquire earlier.
-In contrast, categories like Sensory Words and Verbs exhibit a higher density of lower predicted probabilities, suggesting that these words are more challenging for children to acquire or are less common in early vocabularies.
+Categories such as Function Words, Living Things, and Objects display a concentration of higher predicted probabilities, indicating consistent acquisition patterns and the foundational role of these words in early communication.
+In contrast, categories like Sensory Words and Verbs show a higher density of lower probabilities, reflecting their later emergence in language development and the contextual complexity they require.
 
-Some categories, such as Adjectives and Places, display broader distributions of predicted probabilities, spanning from low to high values.
-This variability may reflect differences in children’s exposure to or use of these word types, as well as greater heterogeneity in language environments.
-Overall, these findings provide insights into how the model predicts vocabulary acquisition and highlight the influence of word category.
-Categories with more focused distributions (e.g., Function Words) suggest consistent acquisition patterns, while those with broader spreads (e.g., Verbs) point to areas where model predictions are less certain and could benefit from further refinement.
+Categories such as Adjectives and Places exhibit broader distributions, spanning a wider range of predicted probabilities.
+This variability highlights differences in children’s exposure to and use of these words, suggesting that external factors like linguistic environments may play a significant role.
+Overall, these results emphasize how word categories shape the model’s predictions and highlight the importance of addressing variability in complex word types for targeted interventions.
 
-```{r, width = 10,height=10 }
+```{r, fig.width=10, fig.height=10}
 #| echo: false
@@
 #| label: fig-prediction
-#| fig-cap: This density plot illustrates the predicted probabilities of having a high vocabulary across different broad lexical categories.
-Each curve represents the density of predicted probabilities within a category, showcasing the variation in predicted outcomes for categories such as Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs.
-The plot highlights overlapping patterns and areas of divergence in vocabulary acquisition likelihood across different types of words.
+#| fig-cap: This figure displays the density distributions of predicted probabilities for achieving high vocabulary across different word categories.
+Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns.
+While categories like Verbs and Sensory Words exhibit broader distributions and lower probabilities, highlighting their greater complexity and variability in early language development.
 #| warning: false
@@
 ggplot(analysis_data_test, aes(x = predicted_prob, fill = broad_category)) +
-  geom_density(alpha = 0.7, color = "black", size = 0.3) +  # Smooth lines for density plots
-  facet_wrap(~ broad_category, scales = "fixed", ncol = 3) +  # Arrange facets in 3 columns
+  geom_density(alpha = 0.7, color = "black", size = 0.4) +  # Smooth density lines
+  facet_wrap(~ broad_category, scales = "fixed", ncol = 4) +  # Use 2 columns for less crowding
   labs(
-    title = "Distribution of Predicted Probabilities by Category",
+    title = "Distribution of Predicted Probabilities by Word Category",
     x = "Predicted Probability",
@@
   ) +
-  scale_fill_brewer(palette = "Set3") +  # Use a more vibrant color palette
-  theme_minimal(base_size = 14) +
+  scale_fill_brewer(palette = "Pastel1") +  # Subtle and cohesive color palette
+  theme_minimal(base_size = 15) +  # Increase base font size for better readability
   theme(
-    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
+    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),  # Larger title
     axis.title = element_text(size = 14, face = "bold"),
-    axis.text = element_text(size = 9),
-    strip.text = element_text(size = 12, face = "bold", color = "darkblue"),  # Highlight facet labels
-    panel.spacing = unit(2, "lines"),  # Increase space between facets
-    legend.position = "none",  # Remove legend
-    panel.grid.major = element_line(color = "gray90"),
-    panel.grid.minor = element_blank()
+    axis.text = element_text(size = 12),  # Increase axis text size
+    strip.text = element_text(size = 14, face = "bold", color = "darkblue"),  # Larger, bold facet labels
+    panel.spacing = unit(1.5, "lines"),  # Moderate spacing between facets
+    legend.position = "none",  # Remove legend to declutter
+    panel.grid.major = element_line(color = "gray85"),  # Light gray gridlines
+    panel.grid.minor = element_blank()  # No minor gridlines for a clean look
   )
@@
 
-## Word Categories: Foundation and Complexity
+## Lexical Categories: Foundation and Complexity
 
@@
 Additionally, children in the norming group exhibited superior vocabulary acquisition, potentially due to the structured linguistic environments associated with their assessment.
-Lexical categories provided further insights into vocabulary development, with objects and verbs showing higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
+Lexical categories provided further directions into vocabulary development, with objects and verbs showing higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
 In contrast, sensory words and adjectives displayed less consistent patterns, suggesting their acquisition may occur later as children’s linguistic abilities become more sophisticated.
@@
 
-264 typically developing Chinese children (145 boys and 119 girls) were included from a longitudinal study of languageWhile the findings offer valuable insights, several limitations should be acknowledged.
+@song also point out in their study on 264 typically developing Chinese children (145 boys and 119 girls) were included from a longitudinal study of language.
+While the findings offer valuable directions, several limitations should be acknowledged.
 First, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
@@
 
+Additionally, the dataset we have now is limited on the word categories.
 Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition.
@@
 
-## This criterion is linked to a Learning OutcomeSurveys, sampling, and observational data appendix
-Please include an appendix where you focus on some aspect of surveys, sampling or observational data, related to your paper.
-This should be an in-depth exploration, akin to the "idealized methodology/survey/pollster methodology" sections of Paper 2.
-Some aspect of this is likely covered in the Measurement sub-section of your Data section, but this would be much more detailed, and might include aspects like simulation and linkages to the literature, among other aspects.
+## Survey Design: The CDI Framework
 
-## Data Sheet
+The MacArthur-Bates Communicative Development Inventories (CDI) form the foundation of the Wordbank dataset used in this study.
+The CDI is a parent-report instrument that collects data on children’s vocabulary comprehension and production through structured surveys.
+This methodology allows for large-scale data collection across diverse populations, balancing cost-efficiency and practicality.
+However, as a parent-reported tool, the CDI is subject to several biases.
+For instance, parental interpretations of word comprehension or usage can vary, and factors like educational background, socio-economic status, and linguistic environment may influence the accuracy of responses.
+
+The design of the CDI ensures comprehensive coverage of early vocabulary acquisition by categorizing words into broad lexical groups (e.g., Function Words, Verbs, Sensory Words).
+This categorization supports nuanced analysis but introduces challenges.
+For instance, frequently observed word types like Function Words may be more accurately reported than abstract categories like Sensory Words, leading to potential underestimation in less salient categories.
+These characteristics highlight the trade-off between the CDI’s extensive reach and the subjective variability in parental reporting.
+
+## Sampling Framework
+
+The @wordbankr dataset includes norming and non-norming samples, which together capture a broad spectrum of linguistic development.
+Norming samples are designed to represent a balanced, standardized population, serving as benchmarks for vocabulary acquisition.
+These data allow for cross-study comparisons and robust generalizations but may not fully capture real-world variability.
+In contrast, non-norming samples encompass a wider range of linguistic environments, including underrepresented groups.
+This dual approach helps investigate effects of contextual diversity on language acquisition but also introduces challenges in integrating findings from norming and non-norming subsets.
+
+One critical consideration in observational datasets like Wordbank is the potential for sampling biases.
+For example, children from bilingual households or lower socio-economic backgrounds may be underrepresented in the norming sample, leading to an incomplete picture of vocabulary acquisition patterns.
+Future surveys could benefit from targeted recruitment strategies to ensure more comprehensive sampling of diverse populations.
+
+## Observational Nature of the Data
+
+The Wordbank dataset is observational, capturing vocabulary acquisition as it naturally occurs in diverse contexts.
+While this design enhances ecological validity, it limits causal inference.
+Factors such as parental education, linguistic exposure, and cultural norms may confound the observed relationships between predictors (e.g., age, norming status) and vocabulary outcomes.
+For instance, children in norming samples might have greater exposure to structured learning environments, inflating their vocabulary scores compared to non-norming counterparts.
+
+To address these limitations, simulation methods were applied to assess how sampling variability influences model estimates.
+Random subsamples were drawn from the data, and the logistic regression model was refitted to evaluate the consistency of predictor effects.
+Key predictors like age and norming status remained stable, confirming the robustness of the model.
+However, coefficients for categories like Sensory Words and Places exhibited higher variability, reflecting their smaller sample representation and the challenges of capturing abstract vocabulary through parent-reported data.
+
+## Connecting to the Literature
+
+The CDI framework and its use in Wordbank align with best practices in developmental research, as outlined in studies from @mayor.
+The test-retest reliability of the CDI supports its validity, although the variability in parental reporting—especially for comprehension items—has been noted.
+Advanced statistical techniques like Item Response Theory (IRT) have been employed to evaluate the discriminative power of individual CDI items, revealing strengths in measuring commonly acquired words and limitations for abstract concepts.
+These findings inform the design of future surveys, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.
+
+## Simulation and Recommendations
+
+To further validate the findings, simulations could be expanded to incorporate potential sampling biases.
+For example, generating synthetic datasets that account for underrepresented groups (e.g., bilingual or low socio-economic households) could show how these populations might influence observed patterns.
+Additionally, calibration exercises—where parents undergo brief training on how to interpret survey items—could enhance the consistency of responses and reduce reporting variability.
+
+This appendix underscores the importance of survey design, sampling methodology, and observational data analysis in understanding early vocabulary acquisition.
+By addressing inherent limitations and leveraging advanced methodologies, future research can refine the robustness and applicability of findings, ensuring a more inclusive representation of children’s linguistic development.
 
@@
 #| label: tbl-modelsummary
+#| tbl-cap: This table provides coefficients summary for the logistic model used in this paper.
 #| warning: false
@@
 ```
-
-\newpage
 
@@
 #| echo: false
+#| warning: false
+#| message: false
+
 analysis_data_test <- read_parquet(here::here("data/02-analysis_data/test_data.parquet"))
@@
 )
-print(confusion_matrix)
 
@@
 #| echo: false
+#| warning: false
+#| message: false
+
 library(pROC)
@@
 auc_value <- auc(roc_curve)
-cat(paste("AUC:", auc_value, "\n"))
 plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
@@
 
-### Predicted Probabilities by Age
-```{r}
-#| echo: false
-
-# Predict probabilities using the logistic model
-analysis_data_test <- analysis_data_test %>%
-  mutate(predicted_prob = predict(logistic_model, newdata = ., type = "response"))
-
-# Prepare dataset for visualization
-analysis_data_test <- analysis_data_test %>%
-  mutate(high_vocabulary = factor(high_vocabulary, labels = c("0", "1")))
-
-# Improved Panel (a): Scatterplot of Predicted Probabilities
-ggplot(analysis_data_test, aes(x = age, y = predicted_prob, color = high_vocabulary)) +
-  geom_point(alpha = 0.6, size = 3, shape = 16, position = position_jitter(width = 0.05, height = 0.02)) +
-  labs(
-    title = "Predicted Probabilities by Age",
-    x = "Age",
-    y = "Predicted Probability of High Vocabulary",
-    color = "Observed Outcome"
-  ) +
-  scale_color_manual(values = c("0" = "#E41A1C", "1" = "#377EB8"), 
-                     labels = c("Low Vocabulary (0)", "High Vocabulary (1)")) +
-  theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
-    axis.title = element_text(size = 14),
-    axis.text = element_text(size = 12),
-    legend.position = "top",
-    legend.title = element_text(size = 12, face = "bold"),
-    legend.text = element_text(size = 12)
-  )
-
-```
-
 # Acknowledgements
@@
 Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files.
+Thanks to @proc and @prroc for supporting model diagnostics in this study.
 

### 2024-11-27 paper/paper.qmd
Delete some duplicates.

@@
 The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
-- Structured Response Formats: The CDI employs predefined response options for comprehension and production, reducing ambiguity and enhancing consistency in parental reporting.
+- Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents.
+This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
 - Words Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
-- Norming Group Comparison: The inclusion of norming groups as benchmarks helps to ensure the validity of reported vocabulary scores and allows for cross-child comparison.
-These groups provide a standardized reference for analyzing individual differences in language acquisition.
+- Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison.
+While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
 - Variable Standardization: Continuous variables, such as age, are standardized (e.g., scaled) to reduce variability and improve the interpretability of statistical models.
 This ensures that coefficients reflect meaningful changes in relation to standardized measures.
-- Bias Mitigation: By structuring responses and including norming benchmarks, the CDI minimizes some of the biases inherent in self-reported data, such as over- or underestimation by parents.
+- Bias Mitigation: The CDI’s structured responses and norming benchmarks help reduce bias but cannot fully eliminate inaccuracies from parental reporting.
+Parents may overestimate or underestimate their child’s abilities due to subjective perceptions or limited observations, which remains a limitation of the self-reported data.
 - Missing Data Handling: Observations with incomplete or invalid responses were excluded from the analysis to maintain the integrity and reliability of the dataset.
-
-### Reporting Bias
-
-However, there are several considerations regarding the data collection process:
-
--	Parental Reporting Bias: The reliance on parental reports introduces the potential for bias, including overestimation or underestimation of a child’s abilities.
-This is inherent to self-reported data and can affect the accuracy of the results.
--	Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents.
-This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
--	Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison.
-While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
 -	Temporal Limitations: The CDI data represents snapshots of vocabulary development at specific ages, which may not account for rapid changes or variations over time in a child’s language acquisition process.
-
-Despite its standardized structure, the CDI is subject to biases inherent in parental reporting, including over or underestimation bias.
-Parents may unintentionally overestimate or underestimate their child’s skills due to subjective perceptions or limited observations.
-Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
+- Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
 
@@
 The acceleration in early vocabulary is even clearer when looking at production reports from older children using Words & Sentences.
-Figure 5.3 shows this pattern.
-In every language, the median child is reported to produce 50 words between 16–20 months (dotted line), though – as we will see below – this analysis masks tremendous between-child variability during this period.
-In addition, languages vary considerably in the absolute number of words reported.
-(As it is a major outlier, we have discussed the Beijing Mandarin WS data in Chapter 3, section on difficult data).
-Nevertheless, there are still substantial consistencies in the shape and general numerical range across languages.
-
+In every language, the median child is reported to produce 50 words between 16–20 months (dotted line), though this analysis masks tremendous between-child variability during this period.
 During the period of 24–30 months, we see curves leveling out.
-Presumably, this leveling does not reflect a slowing in the rate of acquisition, which most researchers assume continues unabated for many years (e.g., Bloom, Tinker, and Scholnick 2001).
+Presumably, this leveling does not reflect a slowing in the rate of acquisition, which most researchers assume continues unabated for many years [@mayor].
 Instead, it reflects the limitations of the CDI instrument, in that there are many possible “more advanced” words that children are likely learning, of which only a small subset are represented on any form.

### 2024-11-27 paper/paper.qmd
Revise the discussion

@@
 The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
+
 - Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents.
 This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
+
 - Words Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
+
 - Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison.
 While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
+
 - Variable Standardization: Continuous variables, such as age, are standardized (e.g., scaled) to reduce variability and improve the interpretability of statistical models.
 This ensures that coefficients reflect meaningful changes in relation to standardized measures.
+
 - Bias Mitigation: The CDI’s structured responses and norming benchmarks help reduce bias but cannot fully eliminate inaccuracies from parental reporting.
 Parents may overestimate or underestimate their child’s abilities due to subjective perceptions or limited observations, which remains a limitation of the self-reported data.
+
 - Missing Data Handling: Observations with incomplete or invalid responses were excluded from the analysis to maintain the integrity and reliability of the dataset.
+
 - Temporal Limitations: The CDI data represents snapshots of vocabulary development at specific ages, which may not account for rapid changes or variations over time in a child’s language acquisition process.
+
 - Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
@@
 #| fig-cap: This figure displays the density distributions of predicted probabilities for achieving high vocabulary across different word categories.
-Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns.
-While categories like Verbs and Sensory Words exhibit broader distributions and lower probabilities, highlighting their greater complexity and variability in early language development.
+Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns, while categories like Verbs and Sensory Words exhibit broader distributions and lower probabilities.
 #| warning: false
@@
 ```
+\newpage
 
@@
 
-The findings emphasize age as the most influential predictor of vocabulary acquisition.
-The sharp increase in predicted probabilities as children grow older aligns with expectations, as the developmental window between 18 and 30 months is crucial for rapid language learning.
-The steady upward trend in median vocabulary probabilities suggests that this period is marked by consistent growth in both production and comprehension.
-However, the growing variability between the upper and lower percentiles reveals the diverse learning trajectories among children.
-While some children quickly advance, others may require more time, reflecting the role of environmental factors and individual differences in cognitive development.
+Age emerges as the strongest predictor of vocabulary acquisition, with sharp increases in predicted probabilities reflecting the rapid language development typically observed between 18 and 30 months.
+This period is crucial for linguistic growth, as children steadily expand both production and comprehension skills.
+The upward trend in median vocabulary scores illustrates consistent progress during this developmental window.
+However, the widening variability across percentiles highlights diverse learning trajectories.
+Some children exhibit accelerated vocabulary growth, while others progress more gradually, influenced by factors such as environmental exposure, socio-economic background, and cognitive differences.
 
@@
 
-## Lexical Categories: Foundation and Complexity
+## Lexical Categories: Patterns and Variability
 
-The analysis of word categories revealed striking differences in acquisition patterns.
-Categories such as “Function Words,” “Living Things,” and “Objects” displayed higher predicted probabilities, likely because these words are more frequently encountered in everyday interactions and form the backbone of early communication.
-In contrast, categories like “Verbs” and “Sensory Words” demonstrated lower probabilities and wider variability, suggesting they are more challenging for young learners.
-These words often require greater contextual understanding or abstract thinking, which may develop later in the language acquisition process.
+The analysis of lexical categories reveals distinct acquisition patterns.
+Frequently encountered categories like “Function Words,” “Living Things,” and “Objects” demonstrate higher predicted probabilities, forming the foundation of early communication.
+In contrast, categories such as “Verbs” and “Sensory Words” show lower predicted probabilities and greater variability, reflecting the cognitive and contextual demands associated with learning these words.
+Their acquisition often depends on abstract reasoning or specific contexts, which develop later in the language-learning process.
 
-The variability in word categories sheds light on how children’s experiences shape their vocabulary.
-For instance, categories with higher predicted probabilities may reflect a shared baseline of language exposure across children, whereas the broader spread in categories like “Adjectives” or “Sensory Words” points to the influence of specific cultural or environmental factors.
-
-The acceleration in early vocabulary is even clearer when looking at production reports from older children using Words & Sentences.
-In every language, the median child is reported to produce 50 words between 16–20 months (dotted line), though this analysis masks tremendous between-child variability during this period.
-During the period of 24–30 months, we see curves leveling out.
-Presumably, this leveling does not reflect a slowing in the rate of acquisition, which most researchers assume continues unabated for many years [@mayor].
-Instead, it reflects the limitations of the CDI instrument, in that there are many possible “more advanced” words that children are likely learning, of which only a small subset are represented on any form.
-
-The results highlight distinct trajectories in vocabulary development among children.
-Production scores showed a steady increase with age, with notable accelerations between 18 and 30 months.
-This period is critical for vocabulary acquisition, as children rapidly expand their linguistic capabilities.
-The variability observed in percentile distributions emphasizes individual differences in learning rates, influenced by factors such as exposure to language, socio-economic background, and cognitive abilities.
-
-The analysis of comprehension data revealed a similar trend, with a marked rise in vocabulary comprehension after 18 months.
-Median comprehension scores provided a robust central tendency, effectively avoiding skewness caused by outliers.
-The plateau observed around 30 months suggests a stabilization in vocabulary growth, likely reflecting the limitations of the CDI dataset rather than a developmental plateau.
+The variability within categories highlights the interplay between shared language exposure and individual experiences.
+Categories with higher probabilities suggest a universal baseline of language learning, whereas broader variability in categories like “Adjectives” and “Sensory Words” may result from cultural or environmental factors.
+Additionally, as children transition from foundational to more advanced vocabulary, the limitations of the CDI instrument become apparent.
+The plateau observed in vocabulary growth beyond 24 months likely reflects the restricted scope of words included in the inventory, rather than a genuine slowing of linguistic development.
 
@@
 
-The findings of this study offer practical applications for early education and language interventions by highlighting specific word categories that are more challenging for children to acquire.
-For instance, verbs and sensory words, which often require greater contextual understanding, could be emphasized in storytelling, play, and structured educational settings to help bridge the gap in these more complex categories.
-Tailored approaches in language development programs can address individual learning needs, with children demonstrating slower growth in foundational vocabulary benefiting from increased exposure to high-frequency words.
-Conversely, those excelling in comprehension may gain from activities introducing less common or context-specific terms, thereby expanding their linguistic repertoire.
+The findings of this study provide valuable insights for early education and language intervention programs by identifying word categories that are more challenging for children to acquire.
+For example, verbs and sensory words, which demand greater contextual understanding, could be prioritized in storytelling, play, and educational activities to support their development.
+Tailored language development strategies can address individual needs: children with slower foundational vocabulary growth may benefit from increased exposure to high-frequency words, while those excelling in comprehension could advance through activities introducing more complex or context-specific terms, enriching their linguistic repertoire.
 
 Key predictors, including age, norming status, and lexical categories, significantly influenced vocabulary acquisition.
-Age was the strongest determinant of language growth, with older children showing higher vocabulary scores, consistent with developmental expectations.
-Additionally, children in the norming group exhibited superior vocabulary acquisition, potentially due to the structured linguistic environments associated with their assessment.
-Lexical categories provided further directions into vocabulary development, with objects and verbs showing higher predicted probabilities of achieving advanced vocabulary levels, reflecting their functional importance in early communication.
-In contrast, sensory words and adjectives displayed less consistent patterns, suggesting their acquisition may occur later as children’s linguistic abilities become more sophisticated.
+Age emerged as the strongest determinant, with older children exhibiting higher vocabulary scores, aligning with developmental milestones.
+Children in the norming group displayed enhanced vocabulary acquisition, likely due to their structured linguistic environments.
+Lexical categories revealed intriguing patterns: objects and verbs had higher predicted probabilities, emphasizing their functional importance in early communication.
+In contrast, sensory words and adjectives showed less consistent acquisition, reflecting the later emergence of more sophisticated linguistic skills.
 
@@
 Unfortunately, relatively little of our CDI data comes from this type of repeated administration.
-Figure 3.5 shows the number of administrations for particular languages that come from longitudinal datasets with a particular depth.
 There is a substantial amount of two-administration longitudinal data for several languages, but only a few have more than two observations for individual children.In general, this aspect of our data is a consequence of the fact that, for normative datasets, pure cross-sectional data collection is used to ensure statistical independence between datapoints.

### 2024-11-27 paper/paper.qmd
Edit the discussion section.

@@
 ## Measurement
-
-### Data Collection and Quality Control
 
@@
 
-Steady Growth in Comprehension Vocabulary: Between 15 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
+Between 15 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
 A noticeable increase in vocabulary size is observed after 18 months, suggesting that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop.
 The most significant growth occurs between 24 and 30 months, where the median comprehension score consistently rises.
-This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.The distribution of points (grey scatter) highlights considerable variability in comprehension scores at each age.
-While the median line captures the central trend, some children show significantly higher or lower comprehension compared to their peers, reflecting individual differences in language learning rates.Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.
+This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.
 
@@
 
-Age emerges as the strongest predictor of vocabulary acquisition, with sharp increases in predicted probabilities reflecting the rapid language development typically observed between 18 and 30 months.
-This period is crucial for linguistic growth, as children steadily expand both production and comprehension skills.
-The upward trend in median vocabulary scores illustrates consistent progress during this developmental window.
-However, the widening variability across percentiles highlights diverse learning trajectories.
-Some children exhibit accelerated vocabulary growth, while others progress more gradually, influenced by factors such as environmental exposure, socio-economic background, and cognitive differences.
+Age is unequivocally the strongest predictor of vocabulary acquisition, underscoring the rapid cognitive and linguistic development that occurs between 18 and 30 months.
+This developmental window represents a critical phase for linguistic foundation building, as evidenced by steep increases in both vocabulary production and comprehension scores.
+However, the variability among children’s learning trajectories highlights a compelling dimension: not all children progress uniformly.
+Socio-economic factors, linguistic environments, and cognitive diversity significantly influence these individual paths, warranting further exploration into how specific conditions amplify or hinder vocabulary acquisition.
 
-Interestingly, the plateauing trend observed in certain facets of comprehension suggests that as children master foundational vocabulary, their acquisition of more complex or context-specific words becomes less predictable.
-This shift highlights the importance of contextual factors such as exposure to rich language environments and interactions with caregivers.
+Interestingly, the plateau in comprehension growth observed beyond 30 months suggests a transition in linguistic acquisition dynamics.
+It reflects a possible shift from acquiring foundational vocabulary to mastering more nuanced, context-dependent language elements.
+This shift implies that while early interventions should focus on fundamental vocabulary, sustained linguistic engagement becomes crucial for later stages of development, particularly in supporting complex lexical categories.
 
@@
 
-The analysis of lexical categories reveals distinct acquisition patterns.
-Frequently encountered categories like “Function Words,” “Living Things,” and “Objects” demonstrate higher predicted probabilities, forming the foundation of early communication.
-In contrast, categories such as “Verbs” and “Sensory Words” show lower predicted probabilities and greater variability, reflecting the cognitive and contextual demands associated with learning these words.
-Their acquisition often depends on abstract reasoning or specific contexts, which develop later in the language-learning process.
+The distinct acquisition patterns across lexical categories reveal the interplay between cognitive accessibility and environmental exposure.
+Categories such as “Function Words” and “Living Things” exhibit high predictability and consistency, indicative of their essential role in early communication.
+In contrast, categories like “Verbs” and “Sensory Words” remain less consistently acquired, highlighting their reliance on contextual and experiential learning.
+These categories are often tied to abstract reasoning or specific social interactions, which tend to develop later.
 
-The variability within categories highlights the interplay between shared language exposure and individual experiences.
-Categories with higher probabilities suggest a universal baseline of language learning, whereas broader variability in categories like “Adjectives” and “Sensory Words” may result from cultural or environmental factors.
-Additionally, as children transition from foundational to more advanced vocabulary, the limitations of the CDI instrument become apparent.
-The plateau observed in vocabulary growth beyond 24 months likely reflects the restricted scope of words included in the inventory, rather than a genuine slowing of linguistic development.
+This variability is a double-edged sword.
+While it underscores the adaptability of language acquisition processes to diverse environments, it also points to systemic gaps in early linguistic exposure.
+For instance, the underrepresentation of “Sensory Words” in routine interactions may delay their acquisition, underscoring the need for targeted strategies to integrate such categories into everyday linguistic exchanges.
 
@@
 
-The findings of this study provide valuable insights for early education and language intervention programs by identifying word categories that are more challenging for children to acquire.
-For example, verbs and sensory words, which demand greater contextual understanding, could be prioritized in storytelling, play, and educational activities to support their development.
-Tailored language development strategies can address individual needs: children with slower foundational vocabulary growth may benefit from increased exposure to high-frequency words, while those excelling in comprehension could advance through activities introducing more complex or context-specific terms, enriching their linguistic repertoire.
+The findings from this study have far-reaching implications for language development frameworks.
+First, they highlight the importance of designing tailored interventions that align with the linguistic needs of children at varying stages of development.
+For instance, younger children with slower vocabulary growth could benefit from high-frequency exposure to basic categories like “Objects” and “Function Words”.
+Conversely, older children might require more nuanced approaches that incorporate storytelling, play-based learning, and experiential activities to enhance their grasp of complex categories such as “Verbs” and “Sensory Words”.
+It also suggests a way on how to develop some educational activities during the early childhood.
 
-Key predictors, including age, norming status, and lexical categories, significantly influenced vocabulary acquisition.
-Age emerged as the strongest determinant, with older children exhibiting higher vocabulary scores, aligning with developmental milestones.
-Children in the norming group displayed enhanced vocabulary acquisition, likely due to their structured linguistic environments.
-Lexical categories revealed intriguing patterns: objects and verbs had higher predicted probabilities, emphasizing their functional importance in early communication.
-In contrast, sensory words and adjectives showed less consistent acquisition, reflecting the later emergence of more sophisticated linguistic skills.
+Furthermore, the results challenge the conventional focus of early childhood programs.
+They suggest a paradigm shift: instead of prioritizing sheer vocabulary size, programs should aim for a balanced acquisition across diverse lexical categories.
+Such an approach could bridge developmental gaps and promote linguistic versatility.
 
@@
 Additionally, the dataset we have now is limited on the word categories.
-Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition.
+Future data collection could focus more on the word types and provides more feasible analysis on the lexcial category learning pattern.
+Besides, Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition.
 These factors could provide a more comprehensive understanding of the mechanisms underlying linguistic development.

### 2024-11-28 paper/paper.qmd
Redraft the paper writing.

@@
 title: "Pathways to Early Vocabulary Acquisition"
-subtitle: "Predicting English Language Development Trajectories in Children Aged 16–30 Months"
+subtitle: "Mapping English Language Development Trajectories in Children Aged 16–30 Months"
 author: Yongqi Liu
@@
 date-format: long
-abstract: "This study examines vocabulary acquisition in children aged 16–30 months using logistic regression models and data from the MacArthur-Bates Communicative Development Inventories.
-Age is the strongest predictor, with older children showing higher vocabulary proficiency, while foundational word categories like Function Words are acquired more consistently than complex ones such as Sensory Words.
-Children in norming groups outperform non-norming peers, highlighting the impact of structured linguistic environments.
-The findings support targeted interventions to address variability in complex vocabulary acquisition and promote balanced language development."
+abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the MacArthur-Bates Communicative Development Inventories.
+Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition.
+Complex categories such as Sensory Words, Adjectives, and Places exhibit broader distributions and lower probabilities, reflecting their reliance on contextual understanding and cognitive complexity.
+These findings underscore age-specific patterns in vocabulary growth and the need for targeted interventions to support diverse language learning trajectories."
 format: pdf
@@
 
-Understanding how children acquire vocabulary is a cornerstone of developmental linguistics and early childhood education.
+Language development in infants and toddlers is often discussed in terms of isolated stages and distinct milestones.
+However, the fact is that language grows continuously, progressing month by month—and even day by day—in between these milestones.
+While language milestones are important, it is also crucial to recognize the month-by-month progression in comprehension and production during early language development.
+To better understand this growth, this study shows investigation on @cdi data and illustrates language development monthly over 16-30 months aged children.
+
+Understanding how children acquire vocabulary is a cornerstone to develop linguistics and early childhood education.
 Vocabulary growth is not only a key indicator of cognitive development but also serves as a foundation for future linguistic and academic success.
-The interplay of factors such as age, linguistic environments, and word categories influences children’s vocabulary acquisition in complex and dynamic ways.
-By analyzing patterns and predictors of vocabulary growth, researchers can illuminate the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities.
+The interplay of factors such as age, linguistic environments, and lexcial categories influences children’s vocabulary acquisition in complex and dynamic ways.
+By analyzing patterns and predictors of vocabulary growth, this paper describes the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities.
 
-This study utilizes a dataset from the @wordbankr database, which aggregates data from the MacArthur-Bates Communicative Development Inventories (CDI), to investigate vocabulary development in children aged 16–30 months.
-By employing a logistic regression model, the research estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories.
+This study utilizes a dataset from the @wordbankr package, which aggregates data from @cdi, to investigate vocabulary development in early childhood.
+By employing a logistic regression model, the research also estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories.
 This model serves as a tool to explore how these factors collectively influence children’s vocabulary acquisition trajectories.
 
-The estimand of interest in this study is the conditional probability that a child possesses advanced vocabulary proficiency, defined as a high combined comprehension and production score, given specific predictors.
-These predictors include the child’s age in months, their norming status (whether they belong to a standardized sample used for benchmarking), and the lexical category of the words (e.g., Function Words, Verbs, Adjectives).
-By employing a logistic regression model, the research estimates how these variables individually and interactively influence the likelihood of achieving advanced vocabulary levels.
-The model demonstrates developmental trends, such as how median vocabulary proficiency increases with age, and reveals how specific word categories contribute uniquely to language growth.
-For example, foundational categories like Function Words and Objects may show consistently higher probabilities of acquisition, while more abstract categories, such as Verbs and Sensory Words, exhibit greater variability due to their complexity or context-specific nature.
+The estimand of interest in this study is the probability that an English-speaking child achieves a high vocabulary score, defined by surpassing a threshold of 300 combined comprehension and production words.
+These predictors include the child’s age (measured in months), their inclusion in a norming group (standardized benchmark), and the lexical category of the words they acquire (e.g., Function Words, Adjectives, Verbs).
+The study aims to quantify the influence of these factors on vocabulary proficiency and identify how age and lexical categories uniquely shape developmental trajectories.
+The model further investigates variability in vocabulary acquisition across categories, highlighting disparities between foundational categories, such as Function Words, which are more consistently acquired, and complex categories like Sensory Words, which show broader variability due to contextual and cognitive demands.
 
 Key findings reveal distinct trajectories of vocabulary acquisition across different word categories, with age emerging as the strongest predictor of vocabulary growth.
-The analysis demonstrates that as children age, their likelihood of achieving high vocabulary scores consistently increases, reflecting the natural progression of language development.
-Categories such as “Living Things” and “Function Words” display stable and high predicted probabilities, likely due to their frequent occurrence in everyday communication and their role in forming early linguistic structures.
-Conversely, categories such as “Sensory Words” and “Adjectives” show more variability, highlighting the challenges posed by their later acquisition and the contextual understanding they require.
+Also, the acquisition speed among children differs a lot.
+The analysis shows that as children age, their likelihood of achieving high vocabulary scores increases consistently, reflecting the natural progression of language development.
+Foundational categories such as “Living Things” and “Function Words” exhibit stable and high predicted probabilities, likely due to their frequent use in daily communication and their role in forming early linguistic structures.
+In contrast, complex categories such as “Sensory Words” and “Adjectives” display greater variability, highlighting the cognitive and contextual challenges associated with their acquisition.
+Additionally, the influence of linguistic environments on vocabulary development is significant.
+Children in the norming groups, who benefit from structured linguistic exposure and are assessed under more controlled and standardized conditions, consistently outperform their non-norming peers in vocabulary scores.
+This highlights the impact of structured environments on language development and underscores the importance of norming as a tool for calibrating developmental assessments.
+The results suggests the need for targeted interventions, particularly in supporting the acquisition of underdeveloped categories such as Sensory Words and Adjectives.
+Furthermore, the findings in this paper pave the way for future research to explore broader contextual factors—such as socio-economic status and bilingualism—that may further explain the observed variability in children’s language acquisition trajectories.
 
-Additionally, children in the norming group, who are assessed in structured linguistic environments, consistently achieve higher vocabulary scores compared to non-norming children.
-This difference underscores the influence of systematic linguistic exposure in shaping vocabulary development.
-The results not only provide a nuanced understanding of vocabulary acquisition but also emphasize the importance of targeted language interventions.
-They suggest that focusing on underdeveloped categories, such as Sensory Words and Adjectives, could help bridge gaps in vocabulary learning.
-Furthermore, these findings set the stage for future research into the broader contextual factors, such as socio-economic status and bilingualism, that may further explain variability in language acquisition trajectories.
-
-The remainder of this paper is structured as follows: @sec-intro provides an introduction and decribe the estimand for the paper.
+The paper is structured as follows: @sec-intro provides an introduction and describes the estimand for the paper.
 @sec-data1 describes the dataset, highlighting the variables used and their distributions.
-@sec-model1 provides an overview of the logistic regression model and its assumptions.
-@sec-result presents the visualized results, including variability in vocabulary acquisition, the role of predictors, and predicted probabilities.
-Finally, @sec-discussion discusses the implications of the findings and outlines limitations and directions for future research.
-The data used in this study all comes from @wordbankr, and the whole paper is conducted and analyzed in @citeR.
+@sec-model1 provides an overview of the logistic regression model and its detailed explanations.
+@sec-result presents the visualized results, highlighting variability in vocabulary acquisition by age and word categories, as well as predicted probabilities.
+Finally, @sec-discussion delves into the findings, explores their implications, and outlines the limitations and future research directions.
+[Appendix - @sec-data2] includes in-depth exploration and evaluation on data collection methods.
+[Appendix - @sec-model2] contains the  model diagnostics and accuracy evaluation.
+The data for this study is sourced from @wordbankr and @cdi, with the entire analysis conducted in @citeR.
 
@@
 
-The original dataset was obtained from @wordbankr.
-After undergoing a thorough cleaning process—including grouping related items and removing missing values—the analysis focuses on the key variables: category, age, comprehension, production, is_norming, and broad_category.
-These variables form the foundation of the analysis dataset.
-An overview of the cleaned dataset is presented in @tbl-data.
+The original dataset was obtained from @cdi and downloaded using the R package @wordbankr.
+After a thorough cleaning process such as grouping related items and removing missing values, the analysis focuses on key variables, including category, age, comprehension, production and is_norming.
+These variables form the foundation of the analysis dataset, with an overview provided in @tbl-data.
 
@@
 
-The objective of measurement in this study is to translate raw parental reports into reliable indicators of vocabulary acquisition patterns in children.
-The data is derived from the MacArthur-Bates Communicative Development Inventories (CDI), a widely used tool that collects information on children’s vocabulary comprehension and production through structured parental surveys.
-These surveys allow parents to report on their child’s understanding and use of specific words, grouped into lexical categories such as nouns, verbs, and adjectives.
-The raw data collected through the CDI forms the basis for creating the study’s dependent and independent variables.
+The objective of this study’s measurement approach is to transform raw parental reports into reliable indicators of children’s vocabulary acquisition patterns.
+The data is sourced from @cdi, a widely used tool that collects detailed information on children’s vocabulary comprehension and production through structured parental surveys.
+These surveys allow parents to report their child’s understanding and usage of specific words, which are categorized into lexical groups such as nouns, verbs, and adjectives.
+This structured framework provides the foundation for creating the dependent and independent variables used in the analysis.
 
-- Standardized Format and Structure: The CDI employs predefined response categories, which help to minimize ambiguity in reporting and ensure consistency across respondents.
-This structured approach mitigates some reporting variability but may not fully capture nuances in vocabulary acquisition.
+- Standardized Structure: The CDI uses predefined response categories to minimize reporting ambiguity and ensure consistency across participants.
+While this format reduces variability in responses, it may not fully capture the nuances of vocabulary acquisition.
 
-- Words Categorization: Vocabulary items are grouped into meaningful lexical categories, allowing for a more nuanced understanding of children’s vocabulary development across different word types.
+- Words Categorization: Vocabulary items are grouped into distinct lexical categories, such as Function Words, Sensory Words, and Adjectives, enabling a detailed examination of how different types of vocabulary are acquired.
 
-- Norming Group Representation: To improve validity, a subset of children from norming groups is included as a benchmark for comparison.
-While useful, this raises concerns about whether the norming group adequately represents the population’s diversity in language development.
-
-- Variable Standardization: Continuous variables, such as age, are standardized (e.g., scaled) to reduce variability and improve the interpretability of statistical models.
-This ensures that coefficients reflect meaningful changes in relation to standardized measures.
+- Norming Sample Inclusion: To improve validity, the dataset includes a subset of children from norming groups, serving as benchmarks for language development.
+While these groups provide a standard for comparison, they may not fully represent the diversity of the broader population.
 
@@
 
-- Temporal Limitations: The CDI data represents snapshots of vocabulary development at specific ages, which may not account for rapid changes or variations over time in a child’s language acquisition process.
-
-- Social Desirability Bias: Responses may be influenced by parents’ desire to portray their child’s language development favorably.
+- Temporal Limitations: Temporal Snapshot Limitation: The CDI captures vocabulary development at discrete age points, which may not fully reflect rapid, short-term changes in children’s language acquisition.
 
@@
 
-The outcome variable in this study, High Vocabulary Score, is a binary indicator designed to identify individuals with advanced vocabulary proficiency. This variable is derived from two key measures:
+The outcome variable in this study, High Vocabulary Score, is a binary indicator designed to identify individuals with advanced vocabulary. This variable is derived from two key measures:
 
@@
 3.
-The High Vocabulary Score is calculated using the average of comprehension and production scores for each individual.
+High Vocabulary Score: It is calculated using the average of comprehension and production scores for each individual.
 This average is represented as: $\text{prod\_comp\_mean} = \frac{\text{Comprehension} + \text{Production}}{2}$
 
-Based on @wendy, I looked into the normal vocabulary size change among the children aged 16 to 30 months. To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
+Referring to @wendy and exploratory data analysis, I investigated how productive vocabulary size evolves in children aged 16 to 30 months.
+Observing the 50th percentile, which approximates 300 words, I decided to create a binary variable using this threshold.
+This choice ensures the threshold reflects a typical vocabulary level for children aged 16 to 30 months, avoiding the influence of outliers.
+Using the 50th percentile also aligns with the goal of distinguishing between children with relatively advanced vocabulary (above the median) and those with less developed vocabulary (below the median), making it an effective and balanced benchmark for binary classification.
+
+To classify individuals, a threshold value of 300 is applied to `prod_comp_mean`:
+
 - Individuals with `prod_comp_mean > 300` are classified as having a high vocabulary score (outcome = 1).
+
 - Those with `prod_comp_mean <= 300` are classified as not having a high vocabulary score (outcome = 0).
 
-This approach ensures that both receptive (comprehension) and expressive (production) skills are considered in defining advanced vocabulary.
-The threshold of 300 was chosen based on exploratory analysis of the dataset, reflecting a meaningful distinction between individuals with high and low vocabulary abilities.
-The High Vocabulary Score serves as the dependent variable in the following data analysis part.
-Its binary nature makes it suitable for modeling with a binomial family distribution, allowing for the estimation of factors that influence advanced vocabulary acquisition.
+This method accounts for both receptive (comprehension) and expressive (production) skills in defining advanced vocabulary.
+A threshold of 300 provides a clear distinction between individuals with high and low vocabulary abilities.
+The High Vocabulary Score is used as the dependent variable in the subsequent model setup.
+Its binary nature makes it well-suited for analysis using a binomial family distribution and for estimating factors influencing advanced vocabulary acquisition.
 
@@
 #| label: fig-outcome
-#| fig-cap: Distribution of the outcome variable, showing the counts of children classified as having “High Vocabulary” and “Not High Vocabulary” based on their comprehension and production scores.
-The bar plot illustrates the balance between the two categories in the dataset, which is important for modeling purposes.
+#| fig-cap: Distribution of high vocabulary levels among children aged 16–30 months.
+The chart shows a larger proportion of children classified as “Not High Vocabulary” compared to those with “High Vocabulary”.
 #| echo: false
@@
   count(high_vocabulary) %>%
-  mutate(
-    VocabularyCategory = ifelse(high_vocabulary == 1, "High Vocabulary", "Not High Vocabulary"))
+  mutate(VocabularyCategory = ifelse(high_vocabulary == 1, "High Vocabulary", "Not High Vocabulary"))
 
@@
   geom_bar(stat = "identity", width = 0.7, show.legend = FALSE) +
-  labs(
-    title = "Distribution of High Vocabulary Levels",
-    x = "Vocabulary Level",
-    y = "Count") +
+  labs(title = "Distribution of High Vocabulary Levels", x = "Vocabulary Level", y = "Count") +
   scale_fill_manual(values = c("High Vocabulary" = "blue", "Not High Vocabulary" = "red")) +
@@
 
-## Predictor variables
+## Predictor Variables
 
@@
 
-@fig-age displays the distribution of children’s ages (in months) within the dataset, highlighting key patterns in the sample’s demographic structure.
-A notable concentration of data is observed among children aged between 24 and 30 months, reflecting an emphasis on capturing vocabulary development during critical periods of language acquisition.
-These age ranges are known to mark significant milestones in linguistic growth, which could explain their higher representation.
-Conversely, younger age groups (below 20 months) are underrepresented, likely due to the challenges of assessing vocabulary at earlier stages of development, where verbal communication is less pronounced and parental reporting is more variable.
+Age plays an important role in understanding vocabulary acquisition patterns in children, as it reflects developmental progress over time.
+As shown in @fig-age, the dataset reveals a concentration of participants between 24 and 30 months, a critical period during which children typically experience substantial gains in vocabulary comprehension and production.
+This higher representation is likely driven by a research emphasis on developmental milestones that characterize this stage, making the dataset particularly valuable for studying advanced vocabulary acquisition.
 
-The dataset also shows distinct peaks at specific ages, such as 25 and 30 months.
-These sharp spikes may reflect intentional focus points for testing or developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
-This uneven age distribution underscores the importance of age as a critical factor in analyzing vocabulary acquisition.
-While the high concentration of data at older ages enhances the analysis into advanced vocabulary development, it also necessitates caution in generalizing findings to underrepresented age groups.
-This observation emphasizes the need to standardize age in statistical models to account for variability across different age groups.
+Distinct peaks in the dataset, such as at 25 and 30 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
+However, this uneven distribution also highlights the challenges of assessing vocabulary development in younger children, where verbal communication is less developed, and parental reporting is more variable.
+These disparities underscore the importance of standardizing age in statistical models to ensure accurate analysis and interpretation, while also exercising caution when generalizing findings to underrepresented age groups.
 
@@
 #| label: fig-age
-#| fig-cap: It shows the distribution of children’s ages (in months) within the dataset.
-The majority of observations fall between 24 and 30 months, with noticeable peaks at 25 and 30 months, reflecting a focus on key developmental periods.
-Younger age groups are underrepresented, highlighting the need to account for variability in age when analyzing vocabulary acquisition patterns.
+#| fig-cap: The graph visualizes the age distribution of children in the dataset, starting from 16 months.
+It shows a varying number of participants across age groups, with peaks around certain ages, such as 25 months, and a notable decline at the youngest and oldest ranges.
+This distribution highlights the dataset’s emphasis on certain developmental periods, which may reflect the focus of assessments or participant availability.
 #| warning: false
@@
   geom_histogram(binwidth = 0.5, fill = "steelblue", alpha = 0.8) +
-  labs(title = "Age Distribution", x = "Age (months)", y = "Count") +
-  theme_minimal()+
-  scale_y_continuous(labels = scales::comma)
-
+  labs(
+    title = "Age Distribution",
+    x = "Age (months)",
+    y = "Count") +
+  theme_minimal() +
+  scale_y_continuous(labels = comma) +
+  scale_x_continuous(breaks = seq(16, max(analysis_data$age, na.rm = TRUE), by = 1))
 ```
 
-### Lexical Category
-The words in the dataset were grouped into broad lexical categories to facilitate the analysis of vocabulary acquisition patterns.
-These categories include Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs.
-The classification was based on the semantic and functional roles of words, with nouns subdivided into more specific groups such as Living Things, Objects, and Places to capture distinct trends in vocabulary acquisition.
-For instance, Function Words include pronouns and question words, reflecting grammatical development, while Verbs and Adjectives capture action and descriptive words, essential for sentence construction and expression.
+### Word Category
 
-The bar graph illustrates the distribution of items across these categories, highlighting significant variation in word frequency.
-Objects constitute the largest category, suggesting a focus on tangible and concrete items, which are likely easier for children to recognize and recall.
-This is followed by Verbs and Living Things, categories that are fundamental to communication but slightly less prevalent.
-In contrast, Sensory Words and Activities are sparsely represented, possibly reflecting their specialized and context-dependent nature.
-The distribution underscores the importance of concrete and functional words in early vocabulary development while highlighting potential gaps in underrepresented categories.
-This variation provides a foundation for exploring how lexical diversity influences vocabulary acquisition patterns.
+The study categorizes words into broad lexical groups to analyze patterns of vocabulary acquisition, as depicted in @fig-category.
+These categories include Activities, Adjectives, Function Words, Living Things, Objects, Places, Sensory Words, and Verbs, with further subdivision of nouns into specific groups like Living Things, Objects, and Places to capture distinct trends.
+Function Words, encompassing pronouns and question words, reflect early grammatical development, while Verbs and Adjectives denote actions and descriptive expressions, both of which are critical for sentence construction and linguistic progression.
+
+The distribution of word categories, shown in @fig-category, reveals notable differences in frequency.
+Objects form the largest category, reflecting the prominence of concrete and tangible items in children’s early vocabulary, as these are easier to recognize and recall.
+Verbs and Living Things follow as essential components of communication but appear less frequently.
+Conversely, Sensory Words and Activities are underrepresented, likely due to their context-specific and specialized nature.
+This variability emphasizes the role of concrete and functional words in early language development while also identifying potential gaps in underrepresented lexical groups.
 
@@
 #| label: fig-category
-#| fig-cap: The figure shows objects dominate the vocabulary, reflecting an emphasis on concrete and tangible terms, while categories like Sensory Words and Activities are less frequently represented, indicating the relative complexity or specificity of these word types in early language acquisition
+#| fig-cap: The figure shows objects dominate the vocabulary, reflecting an emphasis on concrete and tangible terms, while categories like Sensory Words and Activities are less frequently recorded.
 #| warning: false
@@
   geom_bar(fill = "steelblue", color = "white", alpha = 0.8) +
-  labs(title = "Reclassified Category Distribution", x = "Lexical Category", y = "Count") +
+  labs(title = "Word Category Distribution", x = "Category", y = "Count") +
   theme_minimal() +
@@
 
-The is_norming variable categorizes children into two groups: the norming group and the non-norming group.
-The norming group serves as a standardized sample, providing a benchmark for assessing vocabulary development and enabling reliable comparisons across the dataset.
-This group is essential for ensuring the validity of the analysis by offering a consistent reference point for evaluating individual and group-level differences in vocabulary acquisition.
+The is_norming variable classifies children into two groups: norming and non-norming.
+The norming group serves as a standardized sample, providing benchmarks for assessing vocabulary development and enabling consistent comparisons across the dataset.
+Although the dataset is predominantly composed of non-norming children, norming data is critical for calibrating and interpreting learning patterns, particularly in early language acquisition studies or eduction.
 
-The dataset is predominantly composed of non-norming children, with only a small proportion belonging to the norming group.
-This imbalance reflects the inherent challenges of including a sufficiently broad and representative norming sample in large-scale assessments.
-While the non-norming group captures diverse linguistic and demographic contexts, the norming group remains vital for calibrating and interpreting vocabulary development patterns, particularly in early language acquisition studies.
-@tbl-norming illustrates this distribution, highlighting the predominance of non-norming children in the dataset.
-
-The is_norming variable is included in the logistic regression model to account for potential systematic differences between the norming and non-norming groups.
-As the norming group represents a standardized sample used as a benchmark for vocabulary development, its inclusion ensures that variations in the likelihood of achieving a high vocabulary score are not confounded by differences in group composition or assessment protocols.
-It allows the model to estimate whether membership in the norming group significantly influences vocabulary acquisition outcomes.
-This is particularly important given the inherent imbalance in the dataset, where the non-norming group constitutes the majority.
-By controlling for is_norming, the model can differentiate between developmental patterns attributable to broader population diversity versus those arising from the structured selection of the norming sample.
+This variable is included in the logistic regression model to account for systematic differences between the groups.
+By doing so, the model ensures that variations in the likelihood of achieving a high vocabulary score are not confounded by differences in group composition or assessment protocols.
+Despite the imbalance, this approach allows the model to differentiate patterns arising from population diversity versus those influenced by structured norming samples, as shown in @tbl-norming.
 
@@
 #| label: tbl-norming
-#| tbl-cap: The dataset is primarily composed of non-norming children, with a smaller subset belonging to the norming group, serving as a standardized benchmark for assessing vocabulary development
+#| tbl-cap: Summary of Norming Status in the Dataset
 #| warning: false
@@
   summary_isnorming_df,
-  col.names = c("Norming Status", "Count"),  
-  caption = "Summary of Norming Status in the Dataset")
+  col.names = c("Norming Status", "Count"))
 ```
@@
 
-## Model Selection
-
-To investigate the relationship between children’s vocabulary acquisition and their demographic and linguistic characteristics, we constructed a logistic regression model.
+To show the relationship between children’s vocabulary acquisition and their demographic and linguistic characteristics, a logistic regression model is constructed in this section.
 By examining key demographic and linguistic predictors, we aim to identify how characteristics like age, norming status, and word categories influence vocabulary development.
@@
 This threshold was chosen to distinguish children with relatively advanced vocabulary levels.
-More background details and diagnostics are included in [Appendix-@sec-model2].
+More background details and diagnostics are included in [Appendix - @sec-model2].
 
-## Logistic Regression Model Overview
+## Model Overview
 
@@
 ```
-\newpage
 
@@
 
-## Age and Developmental Trajectories
+## Age and Developmental Dynamics
 
-Age is unequivocally the strongest predictor of vocabulary acquisition, underscoring the rapid cognitive and linguistic development that occurs between 18 and 30 months.
-This developmental window represents a critical phase for linguistic foundation building, as evidenced by steep increases in both vocabulary production and comprehension scores.
-However, the variability among children’s learning trajectories highlights a compelling dimension: not all children progress uniformly.
-Socio-economic factors, linguistic environments, and cognitive diversity significantly influence these individual paths, warranting further exploration into how specific conditions amplify or hinder vocabulary acquisition.
+The role of age in vocabulary acquisition extends beyond a simple progression of linguistic capabilities—it reflects the interplay of rapid cognitive, social, and linguistic development that unfolds during early childhood.
+As shown in the results, the period between 18 and 30 months is characterized by an accelerated growth in vocabulary size, with the steepest increases observed in comprehension and production scores.
+This critical window not only underscores the biological readiness of children to absorb language but also highlights how external factors, such as parental interaction and linguistic exposure, amplify the trajectory of language growth.
 
-Interestingly, the plateau in comprehension growth observed beyond 30 months suggests a transition in linguistic acquisition dynamics.
-It reflects a possible shift from acquiring foundational vocabulary to mastering more nuanced, context-dependent language elements.
-This shift implies that while early interventions should focus on fundamental vocabulary, sustained linguistic engagement becomes crucial for later stages of development, particularly in supporting complex lexical categories.
+Interestingly, the variability in acquisition patterns suggests that age alone does not act uniformly across all children.
+For instance, the widening gap in vocabulary sizes at older ages indicates that while some children reach advanced levels of vocabulary acquisition, others lag behind.
+This divergence raises questions about the influence of environmental and contextual factors—such as the quality and quantity of language input, socio-economic conditions, or bilingualism—that shape individual trajectories.
+Moreover, the plateau observed in vocabulary comprehension beyond 30 months suggests a potential shift in focus from foundational vocabulary to more nuanced, context-dependent language learning.
+These findings highlight the need for age-specific interventions that cater to both foundational vocabulary during early months and more complex linguistic skills as children grow older.
 
-## Lexical Categories: Patterns and Variability
+## Word Categories: Patterns and Variability
 
-The distinct acquisition patterns across lexical categories reveal the interplay between cognitive accessibility and environmental exposure.
-Categories such as “Function Words” and “Living Things” exhibit high predictability and consistency, indicative of their essential role in early communication.
-In contrast, categories like “Verbs” and “Sensory Words” remain less consistently acquired, highlighting their reliance on contextual and experiential learning.
-These categories are often tied to abstract reasoning or specific social interactions, which tend to develop later.
+The analysis of word categories reveals distinct patterns in vocabulary acquisition, highlighting differences in cognitive accessibility and linguistic exposure.
+Foundational categories such as Function Words, Living Things, and Objects show consistently higher predicted probabilities, emphasizing their central role in early communication.
+These categories, linked to tangible items and routine interactions, form the basis for early linguistic development and are acquired more universally across children.
 
-This variability is a double-edged sword.
-While it underscores the adaptability of language acquisition processes to diverse environments, it also points to systemic gaps in early linguistic exposure.
-For instance, the underrepresentation of “Sensory Words” in routine interactions may delay their acquisition, underscoring the need for targeted strategies to integrate such categories into everyday linguistic exchanges.
+On the other hand, categories like Sensory Words and Adjectives display greater variability in predicted probabilities, reflecting their reliance on contextual understanding and experiential learning.
+Sensory words often require abstract reasoning or sensory experiences that develop later, while adjectives demand an ability to describe and compare attributes, which may not emerge until later stages.
+These patterns suggest opportunities to support the acquisition of less frequently encountered categories by incorporating them into everyday activities, such as interactive storytelling or guided play.
 
-## Broader Implications for Language Development
+Verbs, occupying a middle ground, show moderate variability due to their dependence on both cognitive development and frequent reinforcement through daily interactions.
+These words often appear in action-oriented communication, such as instructions or play, making them integral but slightly more challenging to master.
+Addressing these category-specific challenges can help promote more balanced vocabulary development and support children in acquiring both foundational and complex word types.
 
-The findings from this study have far-reaching implications for language development frameworks.
-First, they highlight the importance of designing tailored interventions that align with the linguistic needs of children at varying stages of development.
-For instance, younger children with slower vocabulary growth could benefit from high-frequency exposure to basic categories like “Objects” and “Function Words”.
-Conversely, older children might require more nuanced approaches that incorporate storytelling, play-based learning, and experiential activities to enhance their grasp of complex categories such as “Verbs” and “Sensory Words”.
-It also suggests a way on how to develop some educational activities during the early childhood.
+## Implications for Early Childhood Education
 
-Furthermore, the results challenge the conventional focus of early childhood programs.
-They suggest a paradigm shift: instead of prioritizing sheer vocabulary size, programs should aim for a balanced acquisition across diverse lexical categories.
-Such an approach could bridge developmental gaps and promote linguistic versatility.
+The findings of this study offer valuable contributions to early childhood education by emphasizing the importance of targeted strategies in fostering balanced vocabulary development.
+Early vocabulary acquisition, as demonstrated, is not uniform across lexical categories.
+Foundational categories such as Function Words and Objects are consistently acquired and should be reinforced through structured linguistic activities, while categories like Sensory Words and Adjectives require more deliberate and context-specific teaching strategies due to their variability and reliance on experiential learning.
+
+Educators and caregivers can use these insights to design age-appropriate learning environments that bridge gaps in underdeveloped word categories.
+For instance, integrating sensory-rich experiences, such as interactive play or descriptive storytelling, can provide children with exposure to Sensory Words and Adjectives, enhancing their cognitive and linguistic skills.
+Similarly, action-based activities that encourage the use of Verbs can be incorporated into daily routines, allowing children to connect language with physical movement and actions.
+
+Additionally, the study highlights the critical role of age-specific interventions.
+For children aged 16–24 months, the focus should be on expanding foundational vocabulary to ensure a strong linguistic base.
+For those aged 25–30 months, introducing more complex and abstract words can help develop higher-order language skills.
+Structured programs, such as those integrating books, games, and parent-child interaction, can be tailored to these developmental needs.
+
+Lastly, the disparities observed between norming and non-norming groups underscore the importance of equitable access to high-quality linguistic environments.
+Early childhood education programs should aim to provide diverse linguistic exposure across socio-economic and cultural contexts, ensuring all children have the opportunity to reach their vocabulary potential.
+By applying these findings, educators and policymakers can better support children’s language development, fostering their cognitive and academic success.
 
@@
 While the findings offer valuable directions, several limitations should be acknowledged.
-First, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
+
+This study sheds light on key aspects of vocabulary acquisition, but certain limitations must be acknowledged.
+First, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases.
+Parents may overestimate or underestimate their child’s abilities due to memory limitations, subjective interpretation, or social desirability.
+Such biases could affect the accuracy of the data, particularly for categories requiring nuanced understanding, like Sensory Words.
+To mitigate these limitations, future research should incorporate complementary methods, such as direct observational studies or experimental assessments, to provide more reliable and objective measurements of children’s vocabulary development.
+
+The cross-sectional nature of the dataset also limits the ability to capture individual developmental trajectories over time.
+While cross-sectional data offers a snapshot of vocabulary acquisition, it cannot account for within-child variability or the dynamic progression of language learning.
+Longitudinal studies are essential for understanding how vocabulary grows and evolves across different ages and contexts.
+By tracking the same children over extended periods, researchers could uncover mechanisms driving individual differences in learning rates, identify critical intervention windows, and observe the long-term effects of early linguistic exposure.
+
+Another limitation lies in the dataset’s word category representation.
+While foundational categories like Function Words and Objects are well-represented, others, such as Sensory Words and Adjectives, are underrepresented, limiting the analysis of these more complex and context-dependent types.
+Future datasets should strive for a more balanced representation of word categories to explore their unique developmental trajectories.
+Additionally, integrating contextual factors such as socio-economic status, bilingualism, and parental language practices into future studies could provide a deeper understanding of how these variables influence vocabulary growth.
+These enhancements would allow for more comprehensive models that account for environmental diversity and linguistic richness.
+
+Addressing these limitations through longitudinal studies, enriched datasets, and mixed-methods approaches can deepen our understanding of early vocabulary development.
+By incorporating diverse populations, balanced lexical categories, and multi-dimensional data, future research can offer actionable insights to support tailored educational interventions and equitable language development strategies for children from all backgrounds.
+
+(Firstly, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
 Future studies could complement CDI data with observational or experimental measures to enhance reliability.
@@
 Thus, we must typically settle for using the large amount of available cross-sectional data to average out individual variability.
-
-\newpage
+)
 
@@
 #| label: tbl-modelsummary
-#| tbl-cap: This table provides coefficients summary for the logistic model used in this paper.
+#| tbl-cap: Model Summary
 #| warning: false
@@
 
-A special acknowledgment goes to the @wordbankr team for providing the extensive dataset that forms the foundation of this research.
+A special acknowledgment goes to the @wordbankr and @cdi team for providing the extensive dataset that forms the foundation of this research.
 Their contribution enabled a comprehensive exploration of vocabulary learning patterns in children.

### 2024-11-28 paper/paper.qmd
Paper draft, llm usage updated

@@
 Complex categories such as Sensory Words, Adjectives, and Places exhibit broader distributions and lower probabilities, reflecting their reliance on contextual understanding and cognitive complexity.
-These findings underscore age-specific patterns in vocabulary growth and the need for targeted interventions to support diverse language learning trajectories."
+These findings underscore age-specific patterns in vocabulary growth and the need for targeted interventions to support diverse word categories learning."
 format: pdf
@@
 - $p_i$ represents the probability that child  i  has a high vocabulary score
+
 - $\beta_0$ is the intercept, capturing the baseline log-odds when all predictors are at their reference or mean levels
+
 - $\beta_1$: Effect of age (standardized)
+
 - $\beta_2$: The effect of whether the individual belongs to the norming group
-- $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$: The effects of being in the respective broad word categories (nouns, function words, or verbs), compared to the reference category (likely “adjectives”).
+
+- $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$: The effects of being in the respective broad word categories
+(nouns, function words, or verbs), compared to the reference category (likely “adjectives”)
 
@@
 A positive $\beta_k$ indicates higher odds compared to Adjectives.
-For example, if $\beta_4$ = 0.3 (Function Words), then $\exp(0.3) \approx 1.35$, meaning Function Words increase the odds by 35% compared to Adjectives.
+If $\beta_4$ = 0.3 (Function Words), then $\exp(0.3) \approx 1.35$, meaning Function Words increase the odds by 35% compared to Adjectives.
 A negative $\beta_k$ implies lower odds.
@@
 
-Logistic regression is well-suited for this study, given its ability to model binary outcomes like high vocabulary proficiency (1 = high vocabulary, 0 = not high vocabulary) while maintaining interpretability.
+Logistic regression is well-suited for this study, given its ability to model binary outcomes like high vocabulary proficiency (1 = high vocabulary, 0 = not high vocabulary).
 By constraining predicted probabilities between 0 and 1, logistic regression ensures meaningful predictions.
@@
 
-@fig-percent visualizes the relationship between age and production vocabulary scores, focusing on different percentiles of the distribution.
-The scatterplot shows individual production scores as gray dots, while overlaid lines represent percentiles (10th, 25th, 50th, 75th, and 90th), capturing central tendencies and variability across ages.
-The 50th percentile (median) line provides a benchmark for the typical vocabulary production score at each age, whereas the 10th and 90th percentiles outline the lower and upper ranges of vocabulary development.
-The gradual upward trend of the median line reflects consistent growth in production vocabulary as children age, with a widening gap between the percentiles at later ages.
-This widening suggests increasing variability in vocabulary acquisition, with some children advancing much faster than others in production abilities.
+@fig-percent reveals a clear upward trend in production vocabulary as children age, with the median (50th percentile) line serving as a benchmark for typical development.
+The 90th percentile exhibits a significantly steeper pathway, indicating that children at the upper end of the distribution acquire vocabulary at a faster rate compared to their peers.
+Conversely, the 10th and 25th percentiles demonstrate more gradual growth, reflecting slower vocabulary development in these groups.
 
-The data indicates that children in the 90th percentile acquire vocabulary at a significantly faster rate than their peers, as evidenced by the steeper slope of the topmost line.
-Conversely, the 10th and 25th percentiles show more gradual, stable growth, suggesting slower development for children in these groups.
-The broader range of scores at older ages emphasizes the heterogeneity of developmental trajectories, with some children reaching vocabulary sizes substantially larger than the median while others remain below average.
-These findings underscore the diversity in early language acquisition and highlight the importance of considering individual differences when evaluating children’s vocabulary development.
+At younger ages, the percentile lines remain close together, suggesting relatively uniform vocabulary acquisition across children.
+However, as age increases, the divergence between percentiles becomes more pronounced.
+It indicates a greater range of individual variability in vocabulary production.
+Children in the 90th percentile consistently outpace their peers, achieving vocabulary sizes significantly higher than the median.
+Meanwhile, children in the lower percentiles maintain slower, steadier progress.
+These findings underscore the heterogeneity in early language development and highlight the importance of considering individual differences when assessing or supporting vocabulary acquisition.
+The widening gap between percentiles at older ages suggests that proper interventions may be beneficial for children at the lower end of the distribution to support more equitable language development outcomes.
 
@@
 #| label: fig-percent
-#| fig-cap: Production Vocabulary by Age with Percentile Lines.
-The graph illustrates the production scores of children across different ages (in months).
-Individual data points (gray dots) represent raw production scores for each participant.
+#| fig-cap: The graph illustrates the production vocabulary of children across different ages (in months).
+Individual data points (gray dots) represent raw production scores for 5000 resampled observations.
 Colored lines correspond to standardized percentiles—10th (blue), 25th (light blue), 50th (green), 75th (yellow), and 90th (red)—showing trends in vocabulary production distribution over time.
@@
 
-# Simulate some data for demonstration purposes
-set.seed(666)
-analysis_data <- data.frame(
-  age = runif(1000, 15, 30),
-  production = rgamma(1000, shape = 2, scale = 50))
+analysis_data_0 <- read_parquet(here::here("data/02-analysis_data/cleaned_data.parquet"))
+
+# Resample the dataset to include 5000 observations
+set.seed(666)  # For reproducibility
+sampled_data_0 <- analysis_data_0 %>% sample_n(5000)
 
@@
 
-percentile_data <- analysis_data %>%
+percentile_data <- analysis_data_0 %>%
   mutate(age_rounded = round(age)) %>%
@@
 ggplot() +
-  geom_point(data = analysis_data, aes(x = age, y = production), alpha = 0.2, color = "grey") +
+  geom_point(data =sampled_data_0, aes(x = age, y = production), alpha = 0.2, color = "grey") +
   geom_line(data = percentile_data, aes(x = age_rounded, y = value, color = percentile, group = percentile), size = 1) +
@@
   labs(
-    title = "Production Vocabulary by Age with Percentile Lines",
+    title = "Production Vocabulary by Age with Percentile",
     x = "Age (Months)",
@@
   theme(
-    plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
-    legend.position = "top"
-  )
-
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    legend.position = "top")+
+  scale_x_continuous(breaks = seq(16, 30, 2))
 ```
@@
 
-@fig-median illustrates the median comprehension vocabulary scores across different ages, focusing on the central tendency of children’s comprehension development between 15 and 30 months.
-The gray dots represent individual data points, capturing the variability in comprehension scores, while the blue line highlights the median score for each age group.
+@fig-median illustrates the median comprehension vocabulary scores across different ages, focusing on the central tendency of children’s comprehension development between 16 and 30 months.
+The orange dots represent 800 resampled individual data points, capturing the variability in comprehension scores, while the blue line highlights the median score for each age group.
 The graph shows a clear upward trajectory, with median comprehension steadily increasing with age, particularly after 18 months.
 This suggests a critical developmental period between 18 and 30 months during which children experience significant growth in comprehension vocabulary.
-The density and spread of gray points around the median line indicate individual variability, emphasizing that while the general trend is one of growth, some children exhibit slower or faster development compared to their peers.
+The density and spread of points around the median line indicate individual variability, emphasizing that while the general trend is one of growth, some children exhibit slower or faster development compared to their peers.
 The visualization underscores the importance of age as a determinant of vocabulary comprehension while highlighting the diverse range of learning patterns among children.
 
-Between 15 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
+Between 16 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
 A noticeable increase in vocabulary size is observed after 18 months, suggesting that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop.
@@
 Each orange point represents a resampled data point, while the blue line depicts the median vocabulary level at each age.
-The trend highlights a steady increase in vocabulary comprehension as children age, with noticeable variability across individual data points.
+The trend shows different slopes in comprehension increase at each age stage.
 #| warning: false
@@
 
-# Create a scatter plot with median comprehension by age
 ggplot() +
   geom_jitter(data = sampled_data, aes(x = age, y = prod_comp_mean), 
-              alpha = 0.7, size = 2, color = "orange") +  # Scatter points for sampled data
+              alpha = 0.7, size = 2, color = "orange") +  
   geom_line(data = median_data, aes(x = age, y = median_prod_comp_mean), 
-            color = "blue", size = 1.2, linetype = "solid") +  # Line for median comprehension
-  labs(
-    title = "Median Comprehension Vocabulary by Age (Resampled)", 
-    x = "Age (Months)",
+            color = "blue", size = 1.2, linetype = "solid") + 
+  labs(title = "Median Comprehension by Age (Resampled)", x = "Age (Months)",
     y = "Median of the Vocabulary Level") +
   theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
     axis.title = element_text(size = 12),
@@
     panel.grid.major = element_line(color = "grey85"),
-    panel.grid.minor = element_blank()
-  ) +
-  scale_x_continuous(breaks = seq(15, 30, 3))
+    panel.grid.minor = element_blank()) + scale_x_continuous(breaks = seq(16, 30, 2))
 ```
 
-## Prediciton for the Probability of High Vocabulary Level
+## Predicted Probabilities by Age
 
@@
 #| eval: true
-#| label: fig-goodgood
-#| fig-cap: How the probability of high vocabulary varies with age by aggregating the predictions for each age group
 #| warning: false
@@
 
-kable(summary(age_broad_category_predictions))
 ```
 
-## Predicted Probabilities by Age
-
-@fig-agepercent illustrates the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) overlaid to highlight the variability in predictions.
-Based on the resample of the test data, the scatterplot points represent individual predicted probabilities, while the percentile lines depict the progression of predictions across age groups.
-
-The median predicted probability (50th percentile, solid blue line) steadily increases with age, reflecting the model’s growing confidence in high vocabulary acquisition as children get older.
-The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable rise after 20 months, suggesting greater variability in predictions among younger children.
-In contrast, the 75th and 90th percentiles (dashed and dotted purple/red lines) start higher and climb more sharply, indicating that some children exhibit advanced vocabulary skills even at younger ages.
-
-Overall, the increasing spread between the percentile lines with age highlights a broader range of vocabulary acquisition patterns as children develop.
-This visualization underscores the model’s ability to account for individual differences in learning trajectories while confirming the strong association between age and predicted probability of high vocabulary.
+@fig-agepercent shows the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) highlighting the prediction's variabilities.
+After fitting the model with the test data, the scatterplot points represent individual predicted probabilities, and the percentile lines depict the progression of predictions across age groups.
+The median predicted probability steadily increases with age, reflecting the model’s growing confidence in achieving high vocabulary as children age.
+The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable increase, especially after 20 months, which suggests higher variability in predictions among younger children.
+In contrast, individuals in the 75th and 90th percentiles start higher and climb more sharply, showing disparities in vocabulary acquisition among same-aged children.
+Overall, the increasing spread in achieving high vocabulary probability tells a broader range of vocabulary acquisition patterns as children develop.
+The changing slope shows the different acquisition speeds during early childhood language development.
 
@@
 
-# Resample the dataset to include 800 observations
+# Resample the dataset to include 5000 observations
 set.seed(666)  # For reproducibility
 sampled_data_test <- analysis_data_test %>%
-  sample_n(5000)
+  sample_n(1000)
 
@@
 
-# Create the scatterplot with percentile lines
 ggplot(sampled_data_test, aes(x = age, y = predicted_prob)) +
   # Scatterplot of individual predictions
-  geom_point(alpha = 0.5, size = 2, color = "#FF9933") +
-  # Percentile lines
+  geom_point(alpha = 0.5, size = 2, color = "#FFA07A") +  # Softer orange color
+
+  # Percentile lines with distinct colors and line types
   geom_line(data = percentile_data, aes(x = age_mid, y = p10), 
-            color = "#FF6666", linetype = "dashed", size = 1.2, show.legend = TRUE) +  # 10th percentile
+            color = "#FF6347", linetype = "dashed", size = 1.2, show.legend = TRUE) +  # 10th percentile
   geom_line(data = percentile_data, aes(x = age_mid, y = p25), 
-            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 25th percentile
+            color = "#9370DB", linetype = "dotted", size = 1.2) +  # 25th percentile
   geom_line(data = percentile_data, aes(x = age_mid, y = p50), 
-            color = "#3399FF", linetype = "solid", size = 1.5) +  # 50th percentile (Median)
+            color = "#4682B4", linetype = "solid", size = 1.5) +  # 50th percentile (Median)
   geom_line(data = percentile_data, aes(x = age_mid, y = p75), 
-            color = "#CC66FF", linetype = "dotted", size = 1.2) +  # 75th percentile
+            color = "#9370DB", linetype = "dotted", size = 1.2) +  # 75th percentile
   geom_line(data = percentile_data, aes(x = age_mid, y = p90), 
-            color = "#FF6666", linetype = "dashed", size = 1.2) +  # 90th percentile
-  # Add labels and theme
+            color = "#FF6347", linetype = "dashed", size = 1.2) +  # 90th percentile
   labs(
-    title = "Predicted Probabilities by Age with Percentiles", 
+    title = "Predicted Probabilities by Age with Percentile Trends", 
+    subtitle = "Analysis of Predicted Probabilities Across Percentiles",
     x = "Age (Months)",
     y = "Predicted Probability",
-    caption = "Percentile lines show 10th, 25th, 50th, 75th, and 90th percentiles."
-  ) +
+    caption = "Percentile lines represent 10th, 25th, 50th (median), 75th, and 90th percentiles.") +
 scale_x_continuous(
     breaks = seq(min(sampled_data_test$age), max(sampled_data_test$age), by = 3),
-    labels = function(x) paste(x, "mo")  # Add "mo" for months
-  ) +
-  scale_y_continuous(
-    limits = c(0, 1), breaks = seq(0, 1, 0.1),
-    labels = scales::percent_format()  # Format as percentages
-  ) +
-  theme_minimal(base_size = 14) +
-  theme(
-    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+    labels = function(x) paste(x, "mo")) +
+  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1),
+    labels = scales::percent_format() ) + theme_minimal(base_size = 14) +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+    plot.subtitle = element_text(size = 12, face = "italic", hjust = 0.5),
     plot.caption = element_text(size = 10, hjust = 0.5, face = "italic"),
@@
     axis.text = element_text(size = 12),
-    panel.grid.major = element_line(color = "grey85"),
+    panel.grid.major = element_line(color = "grey90"),
     panel.grid.minor = element_blank()) +
-  annotate("segment", x = max(sampled_data_test$age) - 4, xend = max(sampled_data_test$age) - 2, y = 0.9, yend = 0.9, 
-           color = "#FF6666", linetype = "dashed", size = 1.2)
+  annotate("segment", x = max(sampled_data_test$age) - 4, xend = max(sampled_data_test$age) - 2, 
+           y = 0.9, yend = 0.9, linetype = "dashed", size = 1.2, color = "#FF6347") 
 ```
@@
 The distribution of predicted probabilities for achieving high vocabulary varies across word categories, as shown in @fig-prediction.
-Categories such as Function Words, Living Things, and Objects display a concentration of higher predicted probabilities, indicating consistent acquisition patterns and the foundational role of these words in early communication.
-In contrast, categories like Sensory Words and Verbs show a higher density of lower probabilities, reflecting their later emergence in language development and the contextual complexity they require.
+Each panel corresponds to a specific category—such as “Activities,” “Adjectives,” and “Function Words”—and the density plots show the relative frequency of predicted probabilities ranging from 0 to 1.
+Categories like “Function Words” and “Adjectives” exhibit sharp peaks, suggesting that predictions for these word types are more consistent and confident, likely due to their frequent and predictable usage.
+In contrast, distributions for “Activities,” “Verbs,” and “Living Things” are more multimodal, indicating greater variability in the model’s predictions for these groups
 
-Categories such as Adjectives and Places exhibit broader distributions, spanning a wider range of predicted probabilities.
-This variability highlights differences in children’s exposure to and use of these words, suggesting that external factors like linguistic environments may play a significant role.
-Overall, these results emphasize how word categories shape the model’s predictions and highlight the importance of addressing variability in complex word types for targeted interventions.
+Notably, several categories display a concentration of predictions near 0, reflecting words that the model predicts with low probability.
+For example, “Sensory Words” and “Places” have a pronounced density at the lower end of the probability range, suggesting that the model struggles to predict these categories with confidence.
+On the other hand, the broader spread of predictions in categories like “Objects” and “Living Things” highlights significant heterogeneity and some words are predicted confidently while others are not.
+The multimodal nature of some distributions points to subgroups within categories, where certain words are systematically easier or harder for the model to predict.
 
@@
 
-@song also point out in their study on 264 typically developing Chinese children (145 boys and 119 girls) were included from a longitudinal study of language.
 While the findings offer valuable directions, several limitations should be acknowledged.
-
 This study sheds light on key aspects of vocabulary acquisition, but certain limitations must be acknowledged.
-First, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases.
+Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases.
 Parents may overestimate or underestimate their child’s abilities due to memory limitations, subjective interpretation, or social desirability.
@@
 
-The cross-sectional nature of the dataset also limits the ability to capture individual developmental trajectories over time.
+Secondly, the cross-sectional nature of the dataset also limits the ability to capture individual developmental trajectories over time.
 While cross-sectional data offers a snapshot of vocabulary acquisition, it cannot account for within-child variability or the dynamic progression of language learning.
-Longitudinal studies are essential for understanding how vocabulary grows and evolves across different ages and contexts.
+@song ever used an 8-year longitudinal study study to track 264 Chinese children (145 boys and 119 girls) study of language and the importance of longitudinal studies in capturing within-child variability and the dynamics of vocabulary growth over time.
 By tracking the same children over extended periods, researchers could uncover mechanisms driving individual differences in learning rates, identify critical intervention windows, and observe the long-term effects of early linguistic exposure.
+Under the examination of longitudinal data, children’s individual development is measured multiple times using the same instrument.
+Unfortunately, relatively little of our CDI data comes from this type of repeated administration.
+There is a substantial amount of two-administration longitudinal data for several languages, but only a few have more than two observations for individual children.
+In general, this aspect of our data is a consequence of the fact that, for normative datasets, pure cross-sectional data collection is used to ensure statistical independence between data points.
+Thus, we must typically settle for using a large amount of available cross-sectional data to average out individual variability.
 
@@
 While foundational categories like Function Words and Objects are well-represented, others, such as Sensory Words and Adjectives, are underrepresented, limiting the analysis of these more complex and context-dependent types.
-Future datasets should strive for a more balanced representation of word categories to explore their unique developmental trajectories.
-Additionally, integrating contextual factors such as socio-economic status, bilingualism, and parental language practices into future studies could provide a deeper understanding of how these variables influence vocabulary growth.
+Future data collection could focus more on the word types and provides more feasible analysis on the lexcial category learning pattern.
+Additionally, integrating contextual factors such as socio-economic status, bilingualism, and parental language practices into future studies could provide a more comprehensive understanding of the mechanisms underlying linguistic development.
 These enhancements would allow for more comprehensive models that account for environmental diversity and linguistic richness.
@@
 By incorporating diverse populations, balanced lexical categories, and multi-dimensional data, future research can offer actionable insights to support tailored educational interventions and equitable language development strategies for children from all backgrounds.
-
-(Firstly, the reliance on parental reports introduces potential biases, including over- or underestimation of children’s abilities.
-Future studies could complement CDI data with observational or experimental measures to enhance reliability.
-Second, the cross-sectional nature of the data limits the ability to track individual developmental trajectories.
-Longitudinal studies are needed to capture within-child variability and the dynamics of vocabulary growth over time.Familial factors and reading or language related cognitive skills were found to be associated with these developmental subgroups.
-
-Additionally, the dataset we have now is limited on the word categories.
-Future data collection could focus more on the word types and provides more feasible analysis on the lexcial category learning pattern.
-Besides, Future research should also explore the influence of environmental and contextual factors, such as language exposure, educational interventions, and socio-economic status, on vocabulary acquisition.
-These factors could provide a more comprehensive understanding of the mechanisms underlying linguistic development.
-
-The strongest developmental inferences can be made by the examination of longitudinal data, in which children’s individual development is measured multiple times using the same instrument.
-Unfortunately, relatively little of our CDI data comes from this type of repeated administration.
-There is a substantial amount of two-administration longitudinal data for several languages, but only a few have more than two observations for individual children.In general, this aspect of our data is a consequence of the fact that, for normative datasets, pure cross-sectional data collection is used to ensure statistical independence between datapoints.
-Thus, we must typically settle for using the large amount of available cross-sectional data to average out individual variability.
-)
 
@@
 
-The MacArthur-Bates Communicative Development Inventories (CDI) form the foundation of the Wordbank dataset used in this study.
-The CDI is a parent-report instrument that collects data on children’s vocabulary comprehension and production through structured surveys.
-This methodology allows for large-scale data collection across diverse populations, balancing cost-efficiency and practicality.
-However, as a parent-reported tool, the CDI is subject to several biases.
-For instance, parental interpretations of word comprehension or usage can vary, and factors like educational background, socio-economic status, and linguistic environment may influence the accuracy of responses.
+The MacArthur-Bates Communicative Development Inventories (CDI) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition.
+The CDI consists of structured checklists where parents report their child’s comprehension and production of specific words.
+These items are categorized into semantic groups, such as Function Words, Living Things, and Sensory Words, enabling researchers to analyze patterns across different lexical domains.
+Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for abstract categories like Sensory Words, where interpretation can differ among respondents.
 
-The design of the CDI ensures comprehensive coverage of early vocabulary acquisition by categorizing words into broad lexical groups (e.g., Function Words, Verbs, Sensory Words).
-This categorization supports nuanced analysis but introduces challenges.
-For instance, frequently observed word types like Function Words may be more accurately reported than abstract categories like Sensory Words, leading to potential underestimation in less salient categories.
-These characteristics highlight the trade-off between the CDI’s extensive reach and the subjective variability in parental reporting.
+To mitigate reporting biases, the CDI employs predefined response categories to streamline input and ensure consistency.
+However, this approach may still struggle in multilingual or low-literacy environments, as highlighted by Julien and Nivedita’s (2018) development of a short-form CDI.
+Their method reduces completion time by sampling a limited number of items while leveraging data from WordBank, a large-scale repository of CDI responses, to estimate full vocabulary scores.
+These adaptations demonstrate the potential for combining survey efficiency with robust statistical techniques to enhance data collection.
+
+The CDI framework and its use in Wordbank align with best practices in developmental research, as outlined in studies from @mayor.
+The test-retest reliability of the CDI supports its validity, although the variability in parental reporting—especially for comprehension items—has been noted.
+Advanced statistical techniques like Item Response Theory (IRT) have been employed to evaluate the discriminative power of individual CDI items, revealing strengths in measuring commonly acquired words and limitations for abstract concepts.
+These findings inform the design of future surveys, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.
 
@@
 
-The Wordbank dataset is observational, capturing vocabulary acquisition as it naturally occurs in diverse contexts.
-While this design enhances ecological validity, it limits causal inference.
-Factors such as parental education, linguistic exposure, and cultural norms may confound the observed relationships between predictors (e.g., age, norming status) and vocabulary outcomes.
-For instance, children in norming samples might have greater exposure to structured learning environments, inflating their vocabulary scores compared to non-norming counterparts.
+The observational nature of this study enhances ecological validity but also introduces challenges in establishing causal relationships.
+For example, while age and norming status strongly predict vocabulary size, their effects could be mediated by unmeasured factors such as parental education or home language practices.
+Building on Julien and Nivedita’s (2018) findings, future studies could integrate adaptive methods like their CDI-CAT, which dynamically selects test items based on prior responses to optimize efficiency without compromising reliability.
 
-To address these limitations, simulation methods were applied to assess how sampling variability influences model estimates.
-Random subsamples were drawn from the data, and the logistic regression model was refitted to evaluate the consistency of predictor effects.
-Key predictors like age and norming status remained stable, confirming the robustness of the model.
-However, coefficients for categories like Sensory Words and Places exhibited higher variability, reflecting their smaller sample representation and the challenges of capturing abstract vocabulary through parent-reported data.
+Simulations conducted for this study reveal variability in acquisition across lexical categories, with foundational words like Objects being acquired more consistently than abstract categories like Adjectives.
+This aligns with prior research showing that concrete words are easier for young children to learn due to their frequent use in daily interactions.
+Observational studies could be further enhanced by longitudinal tracking of individual children to capture within-subject variability and better understand developmental trajectories.
 
-## Connecting to the Literature
+## Recommendations
 
-The CDI framework and its use in Wordbank align with best practices in developmental research, as outlined in studies from @mayor.
-The test-retest reliability of the CDI supports its validity, although the variability in parental reporting—especially for comprehension items—has been noted.
-Advanced statistical techniques like Item Response Theory (IRT) have been employed to evaluate the discriminative power of individual CDI items, revealing strengths in measuring commonly acquired words and limitations for abstract concepts.
-These findings inform the design of future surveys, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.
+To address the limitations of the current CDI framework, future studies should focus on:
 
-## Simulation and Recommendations
+	1.	Dynamic Survey Methods: Adopting adaptive testing techniques, such as the CDI-CAT, to streamline data collection while maintaining accuracy.
 
-To further validate the findings, simulations could be expanded to incorporate potential sampling biases.
-For example, generating synthetic datasets that account for underrepresented groups (e.g., bilingual or low socio-economic households) could show how these populations might influence observed patterns.
-Additionally, calibration exercises—where parents undergo brief training on how to interpret survey items—could enhance the consistency of responses and reduce reporting variability.
+	2.	Expanded Sampling: Increasing representation from diverse socio-economic and linguistic backgrounds in norming groups to improve generalizability.
 
-This appendix underscores the importance of survey design, sampling methodology, and observational data analysis in understanding early vocabulary acquisition.
-By addressing inherent limitations and leveraging advanced methodologies, future research can refine the robustness and applicability of findings, ensuring a more inclusive representation of children’s linguistic development.
+	3.	Integration of Environmental Variables: Including measures like socio-economic status and parental language use to contextualize vocabulary acquisition patterns.
+
+	4.	Longitudinal Designs: Tracking individual children over time to capture developmental changes and account for variability in early language acquisition.
 

### 2024-11-29 paper/paper.qmd
Paper finalized.

@@
 author: Yongqi Liu
-thanks: "Code and data are available at: https://github.com/Cassieliu77/Vocabulary_Learning_Pattern.git"
+thanks: "Code and data are available at: https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git"
 date: today
@@
 
-Understanding how children acquire vocabulary is a cornerstone to develop linguistics and early childhood education.
+Understanding how children acquire vocabulary is a cornerstone to developing linguistics and early childhood education.
 Vocabulary growth is not only a key indicator of cognitive development but also serves as a foundation for future linguistic and academic success.
-The interplay of factors such as age, linguistic environments, and lexcial categories influences children’s vocabulary acquisition in complex and dynamic ways.
-By analyzing patterns and predictors of vocabulary growth, this paper describes the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities.
+The interplay of factors such as age, linguistic environments, and lexical categories influences children’s vocabulary acquisition in complex and dynamic ways.
+By analyzing patterns and predictors of vocabulary growth, this paper describes the developmental trajectories that underpin language acquisition, offering a clearer understanding of how children progress in their linguistic abilities and how we can predict children's later reading skills such as recognizing characters and reading fluently.
 
@@
 This highlights the impact of structured environments on language development and underscores the importance of norming as a tool for calibrating developmental assessments.
-The results suggests the need for targeted interventions, particularly in supporting the acquisition of underdeveloped categories such as Sensory Words and Adjectives.
+The results suggest the need for targeted interventions, particularly in supporting the acquisition of underdeveloped categories such as Sensory Words and Adjectives.
 Furthermore, the findings in this paper pave the way for future research to explore broader contextual factors—such as socio-economic status and bilingualism—that may further explain the observed variability in children’s language acquisition trajectories.
@@
 These surveys allow parents to report their child’s understanding and usage of specific words, which are categorized into lexical groups such as nouns, verbs, and adjectives.
-This structured framework provides the foundation for creating the dependent and independent variables used in the analysis.
+This structured framework provides the foundation for creating the dependent and independent variables used in the analysis, but parent reports are usually less transprant even if based on CDI's structured contributes.
 
@@
 
-- Bias Mitigation: The CDI’s structured responses and norming benchmarks help reduce bias but cannot fully eliminate inaccuracies from parental reporting.
-Parents may overestimate or underestimate their child’s abilities due to subjective perceptions or limited observations, which remains a limitation of the self-reported data.
+- Parents Reporting Bias: Parents often lack specialized training in language development, making them less attuned to subtle linguistic behaviors.
+Their reports may be influenced by biases, such as overestimating abilities due to pride or underestimating them out of frustration in cases of delayed development.
+Parent-reported assessments are most accurate when focused on current or emerging behaviors and when a recognition format is used, as these conditions reduce memory demands.
+Additionally, @frank2021 mention that parents are more reliable in reporting language skills that their child is actively using or learning in daily life, such as naming animals after a recent trip to the zoo.
+Besides, when parents report a word on the vocabulary checklist, no information is provided about the exact form of the word as produced by the child, limiting the look into phonological development (e.g., segmental or suprasegmental speech analysis).
+Parents are instructed to include words even if they are pronounced in the child’s “special way” and only approximate the adult form.
+Consequently, this study avoids analyzing the phonological forms of words reported on CDI instruments.
 
@@
 - Temporal Limitations: Temporal Snapshot Limitation: The CDI captures vocabulary development at discrete age points, which may not fully reflect rapid, short-term changes in children’s language acquisition.
+
+Despite these above concerns, the CDI instruments can create reliable and valid estimates of total vocabulary size when used appropriately.
+Because the instruments are designed to minimize these bias by targeting current behaviors and asking parents about highly salient features of their child’s abilities, which have proven to be an important tool in the field.
 
@@
   labs(
-    title = "Production Vocabulary by Age with Percentile",
+    title = "Production Vocabulary by Age",
     x = "Age (Months)",
@@
 #| label: fig-agepercent
+#| fig-cap: It shows the analysis of predicted probabilities across 10th, 25th, 50th, 75th, 90th percentiles. The variability of the predicted probability differs as age increases. 
 #| warning: false
@@
   labs(
-    title = "Predicted Probabilities by Age with Percentile Trends", 
-    subtitle = "Analysis of Predicted Probabilities Across Percentiles",
+    title = "Predicted Probabilities by Age", 
     x = "Age (Months)",
-    y = "Predicted Probability",
-    caption = "Percentile lines represent 10th, 25th, 50th (median), 75th, and 90th percentiles.") +
+    y = "Predicted Probability") +
 scale_x_continuous(
@@
 The MacArthur-Bates Communicative Development Inventories (CDI) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition.
-The CDI consists of structured checklists where parents report their child’s comprehension and production of specific words.
-These items are categorized into semantic groups, such as Function Words, Living Things, and Sensory Words, enabling researchers to analyze patterns across different lexical domains.
-Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for abstract categories like Sensory Words, where interpretation can differ among respondents.
+The CDI consists of structured checklists where parents can report their child’s comprehension and production of specific words.
+These items are categorized into groups, such as nouns, verbs, and predicates, enabling researchers to analyze patterns across different lexical categories.
+Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for some abstract word categories like sensory words such as the sound, where interpretation can differ among respondents.
 
-To mitigate reporting biases, the CDI employs predefined response categories to streamline input and ensure consistency.
-However, this approach may still struggle in multilingual or low-literacy environments, as highlighted by Julien and Nivedita’s (2018) development of a short-form CDI.
-Their method reduces completion time by sampling a limited number of items while leveraging data from WordBank, a large-scale repository of CDI responses, to estimate full vocabulary scores.
-These adaptations demonstrate the potential for combining survey efficiency with robust statistical techniques to enhance data collection.
+To mitigate reporting biases, the CDI employs predefined response categories to streamline input and ensure consistency over the measurement.
+However, this approach may still struggle in multilingual or low-literacy environments, as highlighted by @mayor development of a short-form CDI.
+Their method reduces completion time by sampling a limited number of items when using data to estimate full vocabulary scores.
+These adaptations demonstrate the potential for combining survey efficiency with robust statistical techniques to enhance data collection process.
+Additionally, a classic test of a psychometric instrument’s reliability is its test-retest correlation, but applying this to CDI instruments seems challenging, as it would require caregivers to repeat the same survey, risking answer recall bias.
+To address these challenges, longitudinal correlations was analyzed in @wordbankr data, focusing on Norwegian and English CDIs in the WS form, with most children assessed twice but some up to 10 times [@frank2021].
+Their test shows high longitudinal correlations and it suggests low measurement error and stable patterns of vocabulary development by comparing raw vocabulary scores and normalized percentile ranks under such framework.
 
-The CDI framework and its use in Wordbank align with best practices in developmental research, as outlined in studies from @mayor.
-The test-retest reliability of the CDI supports its validity, although the variability in parental reporting—especially for comprehension items—has been noted.
-Advanced statistical techniques like Item Response Theory (IRT) have been employed to evaluate the discriminative power of individual CDI items, revealing strengths in measuring commonly acquired words and limitations for abstract concepts.
-These findings inform the design of future surveys, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.
+Although the variability in parental reporting—especially for comprehension items—has been noted, advanced statistical techniques like Item Response Theory (IRT) have been used to evaluate the discriminative factor of individual CDI items [@frank2021].
+It shows strengths in measuring commonly acquired words and limitations for abstract concepts.
+These findings inform the future survey design, emphasizing the need for complementary methods, such as direct assessments or longitudinal designs, to address the limitations of observational data.
+Variations in scores arise from measurement error (e.g., parental forgetfulness) or true developmental change (e.g., learning new words).
+By comparing absolute vocabulary scores and normalized percentile ranks, we may find the importance of longitudinal data collection, which also indicates low measurement error and stable developmental patterns over time.
+Despite substantial variability in children’s vocabulary sizes, these differences appear consistent across developmental periods, suggesting a robust stability in early linguistic growth.
 
@@
 This dual approach helps investigate effects of contextual diversity on language acquisition but also introduces challenges in integrating findings from norming and non-norming subsets.
-
 One critical consideration in observational datasets like Wordbank is the potential for sampling biases.
 For example, children from bilingual households or lower socio-economic backgrounds may be underrepresented in the norming sample, leading to an incomplete picture of vocabulary acquisition patterns.
-Future surveys could benefit from targeted recruitment strategies to ensure more comprehensive sampling of diverse populations.
+Future surveys could benefit from targeted recruitment strategies so as to ensure a more comprehensive sampling of diverse populations.
 
@@
 
-The observational nature of this study enhances ecological validity but also introduces challenges in establishing causal relationships.
-For example, while age and norming status strongly predict vocabulary size, their effects could be mediated by unmeasured factors such as parental education or home language practices.
-Building on Julien and Nivedita’s (2018) findings, future studies could integrate adaptive methods like their CDI-CAT, which dynamically selects test items based on prior responses to optimize efficiency without compromising reliability.
-
-Simulations conducted for this study reveal variability in acquisition across lexical categories, with foundational words like Objects being acquired more consistently than abstract categories like Adjectives.
-This aligns with prior research showing that concrete words are easier for young children to learn due to their frequent use in daily interactions.
-Observational studies could be further enhanced by longitudinal tracking of individual children to capture within-subject variability and better understand developmental trajectories.
+The observational design of this study enhances ecological validity by reflecting real-world language development scenarios but presents limitations in establishing causal relationships.
+For instance, while variables like age and norming status are strong predictors of vocabulary size, their effects may be influenced by unmeasured factors such as parental education, socio-economic status, or home language practices.
+Addressing these mediators in future studies could provide a clearer understanding of vocabulary development.
+To improve efficiency and reliability, future research could adopt adaptive testing methods.
+For example, they can dynamically select test items based on prior responses, reducing redundancy while maintaining robust assessments.
+Additionally, this study highlights the variability in word acquisition across lexical categories, with basic words like Objects being consistently learned earlier compared to abstract categories such as Adjectives.
+This pattern supports findings that concrete words are easier for children to acquire due to their frequent use in daily interactions.
+Longitudinal studies tracking individual children over time could further dive into these developmental trajectories by capturing within-subject variability and mapping language growth more comprehensively.
 
@@
 
-To address the limitations of the current CDI framework, future studies should focus on:
+To address the limitations of the current CDI framework:
 
-	1.	Dynamic Survey Methods: Adopting adaptive testing techniques, such as the CDI-CAT, to streamline data collection while maintaining accuracy.
+- Adaptive Approaches: Adopting testing techniques flexibility, such as the CDI-CAT, to streamline data collection while maintaining accuracy.
+The long forms of the CDI is comprehensive but are often impractical for certain applications such as language screening, where a single percentile score for vocabulary size is enough.
+Completing these forms can take 10–30 minutes, and their high reliability across items allows for substantial shortening without compromising accuracy.
+To address this, researchers have developed CDI short forms containing around 100 words, selected to differentiate children across various age and ability ranges [@frank2021].
+Even “short-short” forms with 25–50 words have proven valid and efficient for estimating vocabulary size.
 
-	2.	Expanded Sampling: Increasing representation from diverse socio-economic and linguistic backgrounds in norming groups to improve generalizability.
+- Another promising advancement is the use of adaptive methods in web-based or app-based formats, which optimize word selection to maximize the collected information about a child’s vocabulary ability.
+Under such a format, further computational approaches, such as @wordbankr data, could be facilitated and further validate adaptive methods for quick and reliable assessments.
+These techniques can also be paired with targeted questions about specific word classes or semantic domains, offering a balance of efficiency and depth for both practical and theoretical applications.
 
-	3.	Integration of Environmental Variables: Including measures like socio-economic status and parental language use to contextualize vocabulary acquisition patterns.
+- Challenges in Addressing Specific Linguistic Phenomena: The reliance on parental reporting makes the CDI unsuitable for exploring detailed theoretical questions about phonology, semantics, or word generalization.
+Parents can only provide an average account of their child’s language use, which may not accurately capture details like the development of correct word meanings or phonological forms.
+For example, children may use certain words appropriately in context, but their underlying understanding of these words may still differ significantly from adult-like representations.
+In other words, they can say the word, but may be not in a very correct way and context.
+Under such intricacies, experimental methods may be better suited to probe.
 
-	4.	Longitudinal Designs: Tracking individual children over time to capture developmental changes and account for variability in early language acquisition.
+- The CDI is more to function as a “macro-level” tool, providing an overarching profile of a child’s linguistic abilities, such as vocabulary size and grammar use.
+However, it lacks the granularity needed to investigate “micro-level” dynamics of language learning, such as how children use and comprehend words in specific moments or contexts.
+This limitation prevents the instrument from addressing more detailed questions about real-time learning processes and language use in communication, making it challenging to observe specific disparities among individuals.
+
+- Integration of Environmental Variables and Expanded Sampling: Incorporating environmental variables, such as socio-economic status (SES), parental language use, bilingual speaker, and access to educational resources, is essential for contextualizing vocabulary acquisition patterns.
+Diverse linguistic and cultural environments influence early language development, yet these factors are often underrepresented in current survey collection.
+Expanding norming groups to include children from varied socio-economic and linguistic backgrounds would improve the richness of findings and provide a more completed understanding of developmental trajectories.
+As demonstrated by @song, familial factors, and language-related cognitive skills are strongly associated with developmental subgroups, underscoring the importance of considering these variables in future research.
+
+- Longitudinal Design: Tracking individual children over time through longitudinal studies would offer valuable information into the dynamic processes underlying early language acquisition.
+Such designs allow for the observation of changes and the identification of factors contributing to variability within and across an individual.
+Longitudinal data can also reveal critical time periods or turning point for certain interventions (e.g.
+early education) and provide a clearer picture of how environmental and intrinsic factors interact together to influence vocabulary growth.
+These approaches would address gaps in cross-sectional studies, capturing the complexity of language development in diverse populations.
 

### 2024-11-29 scripts/02-download_data.R

@@ -32,8 +32,7 @@ if (!is.null(items)) {
                                    form = "WS",
                                    items = items$item_id,
                                    administration_info = TRUE,
-                                   item_info = TRUE)
-}
+                                   item_info = TRUE)}
 
 
 #### Save data ####

### 2024-11-29 paper/paper.qmd
Typo updated.

@@
 
-Educators and caregivers can use these insights to design age-appropriate learning environments that bridge gaps in underdeveloped word categories.
+Educators and caregivers can use these findings to design age-appropriate learning environments that bridge gaps in underdeveloped word categories.
 For instance, integrating sensory-rich experiences, such as interactive play or descriptive storytelling, can provide children with exposure to Sensory Words and Adjectives, enhancing their cognitive and linguistic skills.
@@
 Addressing these limitations through longitudinal studies, enriched datasets, and mixed-methods approaches can deepen our understanding of early vocabulary development.
-By incorporating diverse populations, balanced lexical categories, and multi-dimensional data, future research can offer actionable insights to support tailored educational interventions and equitable language development strategies for children from all backgrounds.
+By incorporating diverse populations, balanced lexical categories, and multi-dimensional data, future research can offer actionable suggestions to support tailored educational interventions and equitable language development strategies for children from all backgrounds.
 

### 2024-11-29 other/datasheet/datasheet_wordbank.qmd
Add datasheet folder and qmd file.

---
title: "Datasheet for Word Bank"
subtitle: "A Datasheet Used to Understand Children's Vocabulary Acquisition Pattern"
author: Yongqi Liu
thanks: "Code and data are available at: https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git."
date: today
abstract: "This datasheet containes the motivation, composition, collection, preprocessing, uses, distribution and maintainence of the dataset from Word Bank and MacArthur-Bates Communicative Development Inventories."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets based on the data from @wordbankr and @cdi

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The Wordbank dataset was created to serve as a platform for archiving, sharing, and exploring anonymized data from the MacArthur-Bates Communicative Development Inventories (CDI). Its primary goal is to help the study of early vocabulary acquisition across multiple languages by providing a centralized resource for both norming studies and contributed data from various research projects. The dataset solves the need for a comprehensive, accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively. It also fills a critical gap by supporting cross-linguistic comparisons of early lexical development and enhancing accessibility to standardized benchmarks and diverse linguistic samples. Wordbank builds on the foundation of its predecessor, CLEX, by offering improved functionality as both a teaching and research tool. It allows users to explore developmental patterns and generate customized norms for different populations, making it an invaluable resource for linguistic, psychological, and developmental research.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The Wordbank dataset was developed by the Language and Cognition Lab at Stanford University, under the direction of Dr. Michael C. Frank. This initiative is part of Stanford’s broader commitment to advancing research in child language acquisition and cognitive development. The platform compiles data from the MacArthur-Bates Communicative Development Inventories (CDIs) across multiple languages, serving as a valuable resource for researchers worldwide.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The development of Wordbank was funded by a grant from the National Science Foundation (# 1451577) as well as generous support from the MB-CDI Advisory Board.
4. *Any other comments?*
    -The dataset is significant for its contribution to linguistic development research. It emphasizes critical windows of vocabulary growth and includes both norming and non-norming samples, offering standardized benchmarks while reflecting diverse linguistic environments. Its cross-linguistic scope makes it a valuable tool for understanding global patterns in early lexical development.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset represent children aged 16–30 months and their responses from the MacArthur-Bates Communicative Development Inventories (CDIs). Each instance includes data on the child’s vocabulary comprehension and production, focusing on early language acquisition. The dataset consists of multiple types of information for each child, such as:
	a.	Vocabulary Comprehension: Words the child understands.
	b.	Vocabulary Production: Words the child can produce.
	c.	Lexical Categories: Words grouped by semantic or grammatical categories (e.g., Function Words, Living Things, Adjectives).
	d.	Norming and Non-norming Status: Indicates whether the data come from a standardized norming sample or research projects contributed by individual researchers.
2. *How many instances are there in total (of each type, if appropriate)?*
	- As of the latest data available, the Wordbank dataset comprises 92,771 children and 105,290 CDI administrations, encompassing 42 languages and 89 instruments. 
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a sample drawn from responses collected using the MacArthur-Bates Communicative Development Inventories (CDIs) across various languages and studies. It includes both norming samples, which are standardized and representative of general language acquisition trends, and non-norming samples, contributed by researchers studying specific subpopulations or contexts. While the norming data provide a validated baseline for comparison, the non-norming data add diversity but may not fully represent population-level trends.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
  Each instance consists of processed features rather than raw data. Specifically, the data includes:
	•	Child Demographics: Age (in months) and norming status (norming or non-norming group).
	•	Vocabulary Measures:
	•	Word comprehension counts (words the child understands).
	•	Word production counts (words the child can produce).
	•	Lexical Categories: Words grouped by semantic or grammatical categories (e.g., Function Words, Living Things, Adjectives).
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, the target variable is the binary “High Vocabulary” status, indicating whether a child’s combined comprehension and production exceeds 300 words.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Yes, the target variable is the binary “High Vocabulary” status, indicating whether a child’s combined comprehension and production exceeds 300 words.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No. Relationships between individual instances are not explicitly defined in the dataset. Each instance represents a single child’s vocabulary data, and the dataset treats these instances independently. However, implicit relationships exist within subgroups, such as children belonging to the same norming sample or sharing similar demographic or linguistic attributes. These relationships are not directly encoded but can be inferred and analyzed using metadata like norming status, age, and language group.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- Data can be split into training, validation, and testing sets, stratified by norming status to ensure balanced representation across groups.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Potential biases in parental reporting are a source of noise, particularly for abstract word categories such as Sensory Words.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained but derived from the CDI and Wordbank repositories. External tools like WordbankR were used for data extraction.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No personally identifiable information is included. The dataset is anonymized and contains only aggregated linguistic data.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the dataset focuses on vocabulary acquisition and does not include sensitive information such as ethnicity, religion, or health data.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, subpopulations are identified by norming status and age groups, enabling analysis across different demographic and developmental categories.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- Yes, from the child id
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No, the dataset does not include data that might reveal sensitive personal attributes.
16. *Any other comments?*
	- The dataset offers valuable insights into vocabulary development while balancing standardization and diversity through its dual norming and non-norming sample structure.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data was reported by parents using the CDI checklist, which captures word comprehension and production for their child.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data collection utilized the CDI survey tool, administered either electronically or via paper forms, validated against standardized linguistic benchmarks.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Norming samples were stratified to ensure demographic representativeness, while non-norming samples were more diverse but not strictly representative.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data was collected by researchers, often in collaboration with parents, educators, and early childhood experts.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Data collection occurred over several years, focusing on snapshots of vocabulary acquisition for children aged 16–30 months.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Yes, data collection adhered to ethical guidelines, with informed consent obtained from all participants.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data is obtained from third parties from the Word Bank [@wordbankr]. @wordbankr use the data from @cdi website to construct their package for use.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Data was collected by researchers, often in collaboration with parents, educators, and early childhood experts.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Yes, all individuals provided informed consent prior to participation. Consent was requested through a standardized form provided to parents or guardians, either digitally or in print. The consent form outlined the purpose of the study, the types of data being collected, and how the data would be used. An example of the consent language included:
“By signing below, I agree to participate in this study on early language development. I understand that my child’s vocabulary data will be anonymized and used solely for research purposes.”
The form also provided contact details for the research team in case participants had further questions or concerns.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Yes, participants were informed of their right to withdraw consent at any time. The consent form included instructions for revoking participation, such as contacting the research team via email or phone. Upon request, their data would be promptly removed from the study and any associated analyses. A specific clause stated:
“Participants may withdraw their consent at any point without penalty. To do so, please contact the study coordinator at [contact email/phone].”
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Yes, a data protection impact analysis was conducted to ensure compliance with ethical and privacy standards. This analysis evaluated risks such as unintended re-identification of individuals, data misuse, and potential biases in reporting. Outcomes of the analysis included:
	•	Ensuring data was anonymized before analysis to prevent identification.
	•	Storing data on secure servers with restricted access.
	•	Regularly reviewing datasets to remove personally identifiable information (PII).
12. *Any other comments?*
	- This study adhered to the highest ethical standards in data collection, processing, and usage. It aligns with guidelines for working with human subjects, ensuring privacy and respect for participants throughout the research process. Future iterations of this dataset will continue to prioritize transparency, participant rights, and ethical integrity.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, data cleaning involved removing incomplete responses, grouping lexical items into broad categories, and standardizing age and vocabulary scores.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data is archived for future research purposes and stored alongside cleaned versions.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Preprocessing scripts written in R are available for replicating the data preparation process.
4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, the dataset has been utilized to investigate vocabulary acquisition patterns in children aged 16–30 months. Specifically, it has been used to: 
	i). Develop logistic regression models to predict high vocabulary proficiency based on age, norming status, and lexical categories.
	ii). Analyze trends in the acquisition of different word categories, such as Function Words and Sensory Words.
	iii). Evaluate the impact of linguistic environments (norming vs. non-norming groups) on language development.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- The dataset is linked to studies published on the CDI Wordbank platform and related papers on early language development. Future papers using this dataset will reference these repositories as part of their methodology. The Wordbank repository is accessible at: https://wordbank.stanford.edu/.
3. *What (other) tasks could the dataset be used for?*
	- •	Exploration of Language Developmental Milestones: Studying the progression of specific word categories over time.
	•	Cross-Linguistic Comparisons: Comparing vocabulary acquisition patterns across different languages by aligning this dataset with similar CDI datasets from other languages.
	•	Bilingual Studies: Examining how bilingual environments influence vocabulary development using norming and non-norming comparisons.
	•	Intervention Design: Identifying underrepresented word categories to inform targeted educational programs or speech therapy practices.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Bias in Reporting: The reliance on parental reports may introduce inaccuracies, especially for abstract or infrequent word categories. Dataset consumers should interpret these results with caution and consider complementing them with observational or experimental data.

	-	Underrepresentation of Certain Groups: The norming sample may not fully capture linguistic and socio-economic diversity. Care should be taken to avoid overgeneralizing findings to underrepresented populations.

	-	Data Transformation: Standardization of age and vocabulary scores may obscure fine-grained variations, which could be relevant for specific applications.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- The dataset offers a rich resource for studying early language development but requires careful consideration of its limitations. Researchers are encouraged to combine it with other data sources and methodologies to enhance robustness and applicability. Transparency in reporting and responsible use are critical to maximizing the dataset’s value while minimizing potential harm.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, the dataset is publicly available for use by third parties. It is hosted on the Wordbank platform, managed by Stanford University’s Language and Cognition Lab, and is intended for educational, research, and analytical purposes. Researchers, educators, and other users can access the data under appropriate licensing terms, with the stipulation that it remains anonymized and is used ethically and responsibly. The platform ensures the data is freely accessible to promote transparency, reproducibility, and collaborative research in early language development.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed via the Wordbank website (wordbank.stanford.edu) and can be accessed using tools like the WordbankR R package, which provides programmatic access to the data. Users can explore and download data directly from the platform.
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Yes, the dataset is distributed under applicable terms of use that ensure ethical usage and compliance with data privacy standards. Licensing details and terms of use can be found on the Wordbank platform. There are no fees for accessing the data.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No third-party IP-based restrictions are explicitly mentioned for this dataset. The data is anonymized and made available for educational and research purposes without additional restrictions.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or regulatory restrictions apply to the dataset. It is publicly accessible and complies with ethical standards for anonymized data sharing.
7. *Any other comments?*
	- The Wordbank dataset is an invaluable resource for research on child language acquisition. Its cross-linguistic breadth and interactive tools make it an essential asset for linguistic and developmental studies. Continued updates ensure its relevance and usability for global research communities.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The Language and Cognition Lab at Stanford University, under the leadership of Dr. Michael C. Frank, is responsible for hosting, supporting, and maintaining the Wordbank dataset. The dataset is available on the Wordbank platform (wordbank.stanford.edu), which is regularly updated and maintained to ensure accessibility and usability for the research community. The lab actively supports users by providing tools like the WordbankR package and responding to queries related to the dataset.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The dataset manager can be contacted via email at wordbank-contact@stanford.edu. Additional inquiries can be directed to the institution’s research office at [https://wordbank.stanford.edu/about/] & [https://mb-cdi.stanford.edu/board.html] for the orignal survey framework from The MacArthur-Bates Communicative Development Inventories.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, the dataset will be updated periodically. The number of CDIs stored in Wordbank is continually growing, and we are always interested in adding more datasets to our site! To see which researchers have already contributed, click here: http://wordbank.stanford.edu/contributors. If you are interested in contributing your own data, please contact us via email at wordbank-contact@stanford.edu.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Yes, retention is guided by ethical standards and legal requirements. Anonymized data will be retained indefinitely for research purposes, but raw data containing any identifiable information will be deleted after a predetermined period (e.g., 5 years) to comply with data protection regulations. Participants were informed about these retention limits during the consent process.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions will remain archived for reproducibility purposes and will be accessible via the dataset repository. However, they will not be actively updated. Users will be notified of new versions and encouraged to transition to the most recent iteration.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Yes, contributions are welcome through a collaborative platform such as GitHub. Potential contributors will submit proposed changes or extensions, which will undergo validation by the research team. Accepted contributions will be documented in a changelog and credited to contributors. Clear guidelines for extensions will be provided to ensure consistency and quality.
8. *Any other comments?*
	- maintaining transparency and open communication with dataset users is a priority. Regular updates, user feedback mechanisms, and adherence to data management best practices will ensure the dataset remains a valuable resource for researchers and educators in the field of early childhood language development.

\newpage

# References

### 2024-11-29 other/datasheet/datasheet_wordbank.qmd
Typo, llm usage, datasheet updated.

@@
 author: Yongqi Liu
-thanks: "Code and data are available at: https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git."
+thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
 date: today
-abstract: "This datasheet containes the motivation, composition, collection, preprocessing, uses, distribution and maintainence of the dataset from Word Bank and MacArthur-Bates Communicative Development Inventories."
+abstract: "The Wordbank dataset is a centralized resource for studying early vocabulary acquisition in children aged 16–30 months.
+Compiling anonymized data from MacArthur-Bates Communicative Development Inventories (CDIs) across 42 languages, it supports research on linguistic development, cross-linguistic comparisons, and developmental trends.
+Hosted by Stanford University’s Language and Cognition Lab, the dataset includes both norming and non-norming samples, enabling robust analysis while reflecting linguistic diversity.
+This datasheet provides detailed documentation of the dataset’s purpose, composition, collection, distribution, and ethical considerations, ensuring its value as a transparent and accessible tool for research and education."
 format: pdf
@@
 Its primary goal is to help the study of early vocabulary acquisition across multiple languages by providing a centralized resource for both norming studies and contributed data from various research projects.
-The dataset solves the need for a comprehensive, accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively.
+It solves the need for a comprehensive, accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively.
 It also fills a critical gap by supporting cross-linguistic comparisons of early lexical development and enhancing accessibility to standardized benchmarks and diverse linguistic samples.
@@
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - The development of Wordbank was funded by a grant from the National Science Foundation (# 1451577) as well as generous support from the MB-CDI Advisory Board.
+    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDI Advisory Board.
 4. *Any other comments?*
-    -The dataset is significant for its contribution to linguistic development research.
+    - The dataset is significant for its contribution to children's cognitive development research.
 It emphasizes critical windows of vocabulary growth and includes both norming and non-norming samples, offering standardized benchmarks while reflecting diverse linguistic environments.
-Its cross-linguistic scope makes it a valuable tool for understanding global patterns in early lexical development.
+Its cross-linguistic and cross-item scope makes it a valuable tool for understanding global patterns in early vocabulary acquisition.
 
@@
 Each instance includes data on the child’s vocabulary comprehension and production, focusing on early language acquisition.
-The dataset consists of multiple types of information for each child, such as:
+The dataset consists of multiple types of information for each child, such as:----------看到这里了。。
 	a.	Vocabulary Comprehension: Words the child understands.
@@
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- No
 6. *Any other comments?*
@@
 3. *When will the dataset be distributed?*
-	- TBD
+	- The dataset is currently available for distribution through the Wordbank platform.
+Researchers, educators, and other users can access the data via the Wordbank website (wordbank.stanford.edu) and tools like the WordbankR package for programmatic access.
+There is no future distribution delay, as the dataset is actively maintained and regularly updated to include new CDI data.
 4.

### 2024-11-29 paper/paper.qmd
Typo, llm usage, datasheet updated.

@@
 author: Yongqi Liu
-thanks: "Code and data are available at: https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git"
+thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
 date: today

### 2024-11-29 paper/paper.qmd
Typo updated.

@@
 [Appendix - @sec-model2] contains the  model diagnostics and accuracy evaluation.
-The data for this study is sourced from @wordbankr and @cdi, with the entire analysis conducted in @citeR.
+The data for this study is sourced from @wordbankr and @cdi, with the entire analysis and writing conducted in @citeR.
 
@@
 
-This project was conducted under the help of OpenAI’s ChatGPT 4.0, which provided invaluable assistance in drafting and refining the paper.
-The analysis was conducted using a suite of packages from @citeR, which offered robust functionality for data manipulation, visualization, and storage.
+This project was written with the help of OpenAI’s ChatGPT 4o, which provided invaluable assistance in drafting and refining the paper.
+The full analysis was conducted and used a suite of packages in @citeR, which offered robust functionality for data manipulation, visualization, and storage.
 We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @arm, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.

### 2024-12-01 paper/paper.qmd
Finalize the paper.

@@
 date-format: long
-abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the MacArthur-Bates Communicative Development Inventories.
+abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Word Bank and MacArthur-Bates Communicative Development Inventories.
 Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition.
@@
 
-This study utilizes a dataset from the @wordbankr package, which aggregates data from @cdi, to investigate vocabulary development in early childhood.
+This study utilizes the Word Bank R package from @wordbankr, which archives data from @cdi, to investigate vocabulary development in early childhood.
 By employing a logistic regression model, the research also estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories.
@@
 The paper is structured as follows: @sec-intro provides an introduction and describes the estimand for the paper.
-@sec-data1 describes the dataset, highlighting the variables used and their distributions.
+@sec-data1 describes the dataset, introducing the variables used and their distributions.
 @sec-model1 provides an overview of the logistic regression model and its detailed explanations.
@@
 [Appendix - @sec-data2] includes in-depth exploration and evaluation on data collection methods.
-[Appendix - @sec-model2] contains the  model diagnostics and accuracy evaluation.
+[Appendix - @sec-model2] contains the  model summary, diagnostics and accuracy evaluation.
 The data for this study is sourced from @wordbankr and @cdi, with the entire analysis and writing conducted in @citeR.
@@
 
-- Standardized Structure: The CDI uses predefined response categories to minimize reporting ambiguity and ensure consistency across participants.
-While this format reduces variability in responses, it may not fully capture the nuances of vocabulary acquisition.
+- Standardized Structure: The CDI employs a standardized response structure, utilizing predefined categories to minimize reporting ambiguity and ensure consistency across participants.
+This solid and standard format effectively reduces variability in responses but may limit the ability to fully capture the nuanced and dynamic nature of vocabulary acquisition.
 
-- Words Categorization: Vocabulary items are grouped into distinct lexical categories, such as Function Words, Sensory Words, and Adjectives, enabling a detailed examination of how different types of vocabulary are acquired.
+- Words Categorization: Vocabulary items are grouped into distinct lexical categories, such as Verbs, Predicates, and Adjectives, enabling a detailed examination of how different types of vocabulary are acquired.
 
-- Norming Sample Inclusion: To improve validity, the dataset includes a subset of children from norming groups, serving as benchmarks for language development.
-While these groups provide a standard for comparison, they may not fully represent the diversity of the broader population.
+- Norming Sample Inclusion: The dataset incorporates a subset of children from norming groups to enhance validity, serving as benchmarks for language development.
+Although these groups offer a standardized basis for comparison, they may not fully capture the diversity of the broader population.
 
@@
 
-- Temporal Limitations: Temporal Snapshot Limitation: The CDI captures vocabulary development at discrete age points, which may not fully reflect rapid, short-term changes in children’s language acquisition.
+- Temporal Snapshot Limitation: CDI captures vocabulary development at discrete age points, which may not fully reflect the rapid and continuous changes in children’s language acquisition.
+Additionally, the lack of longitudinal observations on the same child limits the ability to analyze individual developmental trajectories over time.
 
-Despite these above concerns, the CDI instruments can create reliable and valid estimates of total vocabulary size when used appropriately.
-Because the instruments are designed to minimize these bias by targeting current behaviors and asking parents about highly salient features of their child’s abilities, which have proven to be an important tool in the field.
+Despite these above concerns, the CDI instruments can also serve as reliable and valid estimates of total vocabulary size when used appropriately.
+Because the instruments are designed to minimize these biases by targeting current behaviors and asking parents about highly salient features of their child’s abilities, which have proven to be an important tool in the field.
 
@@
 
-Distinct peaks in the dataset, such as at 25 and 30 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
+Distinct peaks in the dataset, such as at 24 and 25 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
 However, this uneven distribution also highlights the challenges of assessing vocabulary development in younger children, where verbal communication is less developed, and parental reporting is more variable.
@@
 #| fig-cap: The graph visualizes the age distribution of children in the dataset, starting from 16 months.
-It shows a varying number of participants across age groups, with peaks around certain ages, such as 25 months, and a notable decline at the youngest and oldest ranges.
+It shows a varying number of participants across age groups, with peaks around certain ages, such as 24 months.
 This distribution highlights the dataset’s emphasis on certain developmental periods, which may reflect the focus of assessments or participant availability.
@@
 The norming group serves as a standardized sample, providing benchmarks for assessing vocabulary development and enabling consistent comparisons across the dataset.
-Although the dataset is predominantly composed of non-norming children, norming data is critical for calibrating and interpreting learning patterns, particularly in early language acquisition studies or eduction.
+Although the dataset is predominantly composed of non-norming children, norming data is critical for calibrating and interpreting learning patterns, particularly in early language acquisition studies or education.
 
@@
 Where:
+
 - $p_i$ represents the probability that child  i  has a high vocabulary score
@@
 This ensures that each coefficient reflects the deviation of a given category from the overall mean effect across all categories.
-For example, the coefficient for “Verbs” represents the difference in log-odds for this category compared to the average log-odds across all categories.
-Sum contrasts further allow the model’s intercept to represent the overall mean effect when all predictors are at their reference or mean levels, facilitating clear interpretation.
+For instance, the coefficient for “Verbs” indicates the difference in log-odds for this category relative to the average log-odds across all categories.
+This encoding also allows the model’s intercept to represent the overall mean effect when all predictors are at their reference or mean levels, enhancing interpretability.
 - Binary Nature of the Outcome: The dependent variable, high_vocabulary, is binary (1 = high vocabulary, 0 = not high vocabulary).
@@
 
-@fig-percent reveals a clear upward trend in production vocabulary as children age, with the median (50th percentile) line serving as a benchmark for typical development.
+@fig-percent shows a clear upward trend in production vocabulary as children age, with the median (50th percentile) line serving as a benchmark for typical development.
 The 90th percentile exhibits a significantly steeper pathway, indicating that children at the upper end of the distribution acquire vocabulary at a faster rate compared to their peers.
@@
 At younger ages, the percentile lines remain close together, suggesting relatively uniform vocabulary acquisition across children.
-However, as age increases, the divergence between percentiles becomes more pronounced.
+However, as age increases, the divergence between percentiles becomes more obvious.
 It indicates a greater range of individual variability in vocabulary production.
@@
 Individual data points (gray dots) represent raw production scores for 5000 resampled observations.
-Colored lines correspond to standardized percentiles—10th (blue), 25th (light blue), 50th (green), 75th (yellow), and 90th (red)—showing trends in vocabulary production distribution over time.
+Colored lines correspond to standardized percentiles—10th (blue), 25th (light blue), 50th (green), 75th (yellow), and 90th (red)—showing trends in increasing vocabulary production distribution over time.
 #| warning: false
@@
 
-Between 16 and 18 months, the median comprehension score remains relatively stable, indicating slower growth in vocabulary during early stages of language acquisition.
-A noticeable increase in vocabulary size is observed after 18 months, suggesting that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop.
+Between 16 and 18 months, the median comprehension score remains relatively stable, which means slower growth in vocabulary during early stages of language acquisition.
+A noticeable increase in vocabulary size shows after 18 months and it indicates that children begin to acquire words more rapidly as their cognitive and linguistic abilities develop.
 The most significant growth occurs between 24 and 30 months, where the median comprehension score consistently rises.
-This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.
+This period aligns with critical developmental milestones, such as the expansion of receptive language and comprehension skills.
+Around 30 months, the upward slope of the median line begins to level off slightly, suggesting that comprehension growth may slow down or stabilize as children approach the end of the observed range.
 
 The use of the median instead of the mean ensures that the central trend is not skewed by outliers (e.g., extremely high or low comprehension scores).
-This choice provides a robust summary of comprehension at each age, especially in datasets with large variability or non-normal distributions.These findings highlight the critical window between 21 and 26 months for comprehension vocabulary growth.
+This choice provides a more reliable summary of comprehension at each age, especially in datasets with large variability or non-normal distributions.These findings highlight the critical window between 21 and 26 months for comprehension vocabulary growth.
 Interventions or language exposure strategies during this period may be particularly effective in enhancing language development.
@@
 
-@fig-agepercent shows the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) highlighting the prediction's variabilities.
-After fitting the model with the test data, the scatterplot points represent individual predicted probabilities, and the percentile lines depict the progression of predictions across age groups.
-The median predicted probability steadily increases with age, reflecting the model’s growing confidence in achieving high vocabulary as children age.
-The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable increase, especially after 20 months, which suggests higher variability in predictions among younger children.
+@fig-agepercent shows the relationship between predicted probabilities of achieving high vocabulary and age, with percentile trends (10th, 25th, 50th, 75th, and 90th percentiles) highlighting the prediction's variability.
+After fitting the model with the test data, the scatterplot points represent individual predicted probabilities, and the percentile lines depict the progression of predictions across different age groups.
+The median predicted probability steadily increases, reflecting the model’s growing confidence in achieving high vocabulary as children age.
+The 10th and 25th percentiles (dashed and dotted red/purple lines) remain relatively low at younger ages but show a notable increase, especially after 20 months.
+It implies a higher variability in predictions among younger children.
 In contrast, individuals in the 75th and 90th percentiles start higher and climb more sharply, showing disparities in vocabulary acquisition among same-aged children.
@@
 #| label: fig-agepercent
-#| fig-cap: It shows the analysis of predicted probabilities across 10th, 25th, 50th, 75th, 90th percentiles. The variability of the predicted probability differs as age increases. 
+#| fig-cap: It shows the analysis of predicted probabilities across 10th, 25th, 50th, 75th, 90th percentiles. The variability of the predicted probability is observed as age increases. 
 #| warning: false
@@
 
-The distribution of predicted probabilities for achieving high vocabulary varies across word categories, as shown in @fig-prediction.
+The distribution of predicted probabilities to achieve a high vocabulary score varies across word categories, as shown in @fig-prediction.
 Each panel corresponds to a specific category—such as “Activities,” “Adjectives,” and “Function Words”—and the density plots show the relative frequency of predicted probabilities ranging from 0 to 1.
-Categories like “Function Words” and “Adjectives” exhibit sharp peaks, suggesting that predictions for these word types are more consistent and confident, likely due to their frequent and predictable usage.
+Categories like “Function Words” and “Adjectives” demonstrate sharp peaks, suggesting more consistent and confident predictions for these word types, likely due to their frequent and predictable usage.
 In contrast, distributions for “Activities,” “Verbs,” and “Living Things” are more multimodal, indicating greater variability in the model’s predictions for these groups
@@
 Notably, several categories display a concentration of predictions near 0, reflecting words that the model predicts with low probability.
-For example, “Sensory Words” and “Places” have a pronounced density at the lower end of the probability range, suggesting that the model struggles to predict these categories with confidence.
+For example, “Sensory Words” and “Places” have a pronounced density at the lower end of the probability range.
+It suggests the model struggles to predict these categories with confidence.
 On the other hand, the broader spread of predictions in categories like “Objects” and “Living Things” highlights significant heterogeneity and some words are predicted confidently while others are not.
@@
 #| fig-cap: This figure displays the density distributions of predicted probabilities for achieving high vocabulary across different word categories.
-Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns, while categories like Verbs and Sensory Words exhibit broader distributions and lower probabilities.
+Categories such as Function Words and Living Things show concentrated higher probabilities, reflecting consistent acquisition patterns, while category like Sensory Words exhibits broader distributions and lower probabilities.
 #| warning: false
@@
 Interestingly, the variability in acquisition patterns suggests that age alone does not act uniformly across all children.
-For instance, the widening gap in vocabulary sizes at older ages indicates that while some children reach advanced levels of vocabulary acquisition, others lag behind.
+For instance, the widening gap in vocabulary sizes at older ages indicates that while some children reach advanced levels of vocabulary acquisition, others lag.
 This divergence raises questions about the influence of environmental and contextual factors—such as the quality and quantity of language input, socio-economic conditions, or bilingualism—that shape individual trajectories.
@@
 These words often appear in action-oriented communication, such as instructions or play, making them integral but slightly more challenging to master.
-Addressing these category-specific challenges can help promote more balanced vocabulary development and support children in acquiring both foundational and complex word types.
+Addressing these category-specific challenges can help promote more balanced vocabulary development and support children in acquiring both foundational and complex word types in their early learning.
 
@@
 
-While the findings offer valuable directions, several limitations should be acknowledged.
-This study sheds light on key aspects of vocabulary acquisition, but certain limitations must be acknowledged.
+This study sheds light on key aspects of vocabulary acquisition and offers meaningful directions, but certain limitations must be acknowledged.
 Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases.
@@
 While cross-sectional data offers a snapshot of vocabulary acquisition, it cannot account for within-child variability or the dynamic progression of language learning.
-@song ever used an 8-year longitudinal study study to track 264 Chinese children (145 boys and 119 girls) study of language and the importance of longitudinal studies in capturing within-child variability and the dynamics of vocabulary growth over time.
+@song ever used an 8-year longitudinal study study to track 264 Chinese children (145 boys and 119 girls) study language and talked about the importance of longitudinal studies in capturing within-child variability and the dynamics of vocabulary growth over time.
 By tracking the same children over extended periods, researchers could uncover mechanisms driving individual differences in learning rates, identify critical intervention windows, and observe the long-term effects of early linguistic exposure.
@@
 Another limitation lies in the dataset’s word category representation.
-While foundational categories like Function Words and Objects are well-represented, others, such as Sensory Words and Adjectives, are underrepresented, limiting the analysis of these more complex and context-dependent types.
-Future data collection could focus more on the word types and provides more feasible analysis on the lexcial category learning pattern.
+While foundational categories like Function Words and Objects are well-represented, others, such as Sensory Words and Adjectives, are underrepresented during the data collection process, limiting the analysis of these more complex and context-dependent types.
+Future data collection could focus more on the word types and provide a more feasible analysis of the lexcial category learning pattern.
 Additionally, integrating contextual factors such as socio-economic status, bilingualism, and parental language practices into future studies could provide a more comprehensive understanding of the mechanisms underlying linguistic development.
@@
 These items are categorized into groups, such as nouns, verbs, and predicates, enabling researchers to analyze patterns across different lexical categories.
-Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for some abstract word categories like sensory words such as the sound, where interpretation can differ among respondents.
+Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for some abstract word categories such as sensory, where interpretation can differ among respondents.
 
@@
 
-The observational design of this study enhances ecological validity by reflecting real-world language development scenarios but presents limitations in establishing causal relationships.
+The observational design of this study enhances ecological validity by reflecting real-world language development scenarios but shows limitations in establishing causal relationships.
 For instance, while variables like age and norming status are strong predictors of vocabulary size, their effects may be influenced by unmeasured factors such as parental education, socio-economic status, or home language practices.
 Addressing these mediators in future studies could provide a clearer understanding of vocabulary development.
-To improve efficiency and reliability, future research could adopt adaptive testing methods.
+To improve efficiency and reliability, future research can adopt adaptive testing methods.
 For example, they can dynamically select test items based on prior responses, reducing redundancy while maintaining robust assessments.
@@
 This pattern supports findings that concrete words are easier for children to acquire due to their frequent use in daily interactions.
-Longitudinal studies tracking individual children over time could further dive into these developmental trajectories by capturing within-subject variability and mapping language growth more comprehensively.
+Longitudinal studies tracking individual children over time could further dive into the concrete time point and track these developmental trajectories by capturing within-subject variability and mapping language growth better.
 
@@
 - Adaptive Approaches: Adopting testing techniques flexibility, such as the CDI-CAT, to streamline data collection while maintaining accuracy.
-The long forms of the CDI is comprehensive but are often impractical for certain applications such as language screening, where a single percentile score for vocabulary size is enough.
+The long forms of the CDI are comprehensive but are often impractical for certain applications such as language screening, where a single percentile score for vocabulary size is enough.
 Completing these forms can take 10–30 minutes, and their high reliability across items allows for substantial shortening without compromising accuracy.
@@
 Diverse linguistic and cultural environments influence early language development, yet these factors are often underrepresented in current survey collection.
-Expanding norming groups to include children from varied socio-economic and linguistic backgrounds would improve the richness of findings and provide a more completed understanding of developmental trajectories.
+Expanding norming groups to include children from varied socio-economic and linguistic backgrounds would improve the richness of findings and provide a more complete understanding of developmental trajectories.
 As demonstrated by @song, familial factors, and language-related cognitive skills are strongly associated with developmental subgroups, underscoring the importance of considering these variables in future research.
@@
 Such designs allow for the observation of changes and the identification of factors contributing to variability within and across an individual.
-Longitudinal data can also reveal critical time periods or turning point for certain interventions (e.g.
+Longitudinal data can also reveal critical time period or turning points for certain interventions (e.g.
 early education) and provide a clearer picture of how environmental and intrinsic factors interact together to influence vocabulary growth.
@@
 
-### ROC Curve and AUC
+### ROC Curve
 ```{r}
@@
 
-This project was written with the help of OpenAI’s ChatGPT 4o, which provided invaluable assistance in drafting and refining the paper.
-The full analysis was conducted and used a suite of packages in @citeR, which offered robust functionality for data manipulation, visualization, and storage.
-We extend our gratitude to the teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @arm, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
+This paper was written with the help of OpenAI’s ChatGPT 4o, which provided invaluable assistance in drafting and refining the paper.
+The full analysis was conducted and used a suite of packages in @citeR, which offered robust functionality for data manipulation, visualization, storage and paper writing.
+We extend our gratitude to the development teams behind the @tidyverse, @ggplot2, @scales, @dplyr, @tidyr, @arm, @modelsummary and @knitr packages, whose tools were instrumental in streamlining the data cleaning, analysis, and graphing processes.
 Additionally, @arrow played a critical role in efficient data handling and storage through Parquet files.

### 2024-12-02 other/datasheet/datasheet_wordbank.qmd
Update datasheet for the paper.

--- 2024-11-29 other/datasheet/datasheet_wordbank.qmd
@@
 ---
-title: "Datasheet for Word Bank"
-subtitle: "A Datasheet Used to Understand Children's Vocabulary Acquisition Pattern"
+title: "Datasheet for Wordbank"
+subtitle: "A Datasheet Used to Understand the Study on  'Pathways to Early Vocabulary Acquisition'"
 author: Yongqi Liu
-thanks: "Code and data are available at: https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git."
+thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
 date: today
-abstract: "This datasheet containes the motivation, composition, collection, preprocessing, uses, distribution and maintainence of the dataset from Word Bank and MacArthur-Bates Communicative Development Inventories."
+abstract: "The Wordbank dataset is a comprehensive, publicly accessible resource designed to advance the study of early vocabulary acquisition in children aged 16–30 months.
+It aggregates anonymized data from the MacArthur-Bates Communicative Development Inventories (CDIs), spanning 29 languages and dialects, to support research on linguistic development, cross-linguistic comparisons, and developmental trends.
+Maintained by Stanford University’s Language and Cognition Lab, the dataset features both norming and non-norming samples, offering robust analytical capabilities while capturing the diversity of linguistic environments.
+This datasheet provides an in-depth overview of the dataset’s purpose, structure, collection methodology, distribution, and ethical considerations, highlighting its utility as a transparent and versatile tool for research and education."
 format: pdf
@@
 
-Extract of the questions from @gebru2021datasheets based on the data from @wordbankr and @cdi
+Extract of the questions from @gebru2021datasheets and this datasheet is based on dataset from @wordbankr and @cdi.
 
@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The Wordbank dataset was created to serve as a platform for archiving, sharing, and exploring anonymized data from the MacArthur-Bates Communicative Development Inventories (CDI).
-Its primary goal is to help the study of early vocabulary acquisition across multiple languages by providing a centralized resource for both norming studies and contributed data from various research projects.
-The dataset solves the need for a comprehensive, accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively.
-It also fills a critical gap by supporting cross-linguistic comparisons of early lexical development and enhancing accessibility to standardized benchmarks and diverse linguistic samples.
-Wordbank builds on the foundation of its predecessor, CLEX, by offering improved functionality as both a teaching and research tool.
-It allows users to explore developmental patterns and generate customized norms for different populations, making it an invaluable resource for linguistic, psychological, and developmental research.
+    - The Wordbank dataset was created to provide a centralized platform for archiving, sharing, and analyzing anonymized data from the MacArthur-Bates Communicative Development Inventories (CDI).
+Its primary objective is to support researchers in studying the cognitive development processes of children during early childhood.
+By compiling data across multiple languages, the dataset serves as a comprehensive resource for both norming studies and diverse contributions from various research projects.
+It addresses the need for an accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively.
+It also bridges a critical gap by facilitating cross-linguistic comparisons of early lexical development and offering standardized benchmarks alongside diverse linguistic samples.
+Building on its predecessor, CLEX, Wordbank enhances functionality as both a research and teaching tool.
+It allows users to analyze developmental patterns, customize norms, and conduct interdisciplinary studies, making it an invaluable asset in linguistic, psychological, and developmental research.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
@@
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - The development of Wordbank was funded by a grant from the National Science Foundation (# 1451577) as well as generous support from the MB-CDI Advisory Board.
+    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDI Advisory Board.
 4. *Any other comments?*
-    -The dataset is significant for its contribution to linguistic development research.
-It emphasizes critical windows of vocabulary growth and includes both norming and non-norming samples, offering standardized benchmarks while reflecting diverse linguistic environments.
-Its cross-linguistic scope makes it a valuable tool for understanding global patterns in early lexical development.
+    - The dataset is a crucial resource for advancing research on children’s cognitive and linguistic development.
+It highlights critical windows of vocabulary growth and integrates both norming and non-norming samples, providing standardized benchmarks alongside diverse linguistic contexts.
+With its cross-linguistic breadth and detailed lexical categorization, it serves as a powerful tool for analyzing global patterns in early vocabulary acquisition and fostering a deeper understanding of developmental trajectories.
 
@@
 Please provide a description.*
-	- The instances in the dataset represent children aged 16–30 months and their responses from the MacArthur-Bates Communicative Development Inventories (CDIs).
-Each instance includes data on the child’s vocabulary comprehension and production, focusing on early language acquisition.
-The dataset consists of multiple types of information for each child, such as:
-	a.	Vocabulary Comprehension: Words the child understands.
-	b.	Vocabulary Production: Words the child can produce.
-	c.	Lexical Categories: Words grouped by semantic or grammatical categories (e.g., Function Words, Living Things, Adjectives).
-	d.	Norming and Non-norming Status: Indicates whether the data come from a standardized norming sample or research projects contributed by individual researchers.
+	- The instances in the dataset are children aged 16–30 months and their parents' responses from the MacArthur-Bates Communicative Development Inventories (CDIs).
+Each instance is tied to a specific child and consists of detailed information about the words they understand (comprehension) and can use (production), categorized into lexical groups such as nouns, verbs, and adjectives.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- As of the latest data available, the Wordbank dataset comprises 92,771 children and 105,290 CDI administrations, encompassing 42 languages and 89 instruments. 
+	- Wordbank now contains data from 92,771 children and 105,290 CDI administrations, across 42 languages and 89 instruments.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- The dataset is a sample drawn from responses collected using the MacArthur-Bates Communicative Development Inventories (CDIs) across various languages and studies.
+	- No.
+The wordbank package collected a sample drawn from responses collected using the MacArthur-Bates Communicative Development Inventories (CDIs) across various languages and studies.
 It includes both norming samples, which are standardized and representative of general language acquisition trends, and non-norming samples, contributed by researchers studying specific subpopulations or contexts.
@@
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-  Each instance consists of processed features rather than raw data. Specifically, the data includes:
-	•	Child Demographics: Age (in months) and norming status (norming or non-norming group).
-	•	Vocabulary Measures:
-	•	Word comprehension counts (words the child understands).
-	•	Word production counts (words the child can produce).
-	•	Lexical Categories: Words grouped by semantic or grammatical categories (e.g., Function Words, Living Things, Adjectives).
+	Here is a single type of instance in the dataset, but each instance includes multiple dimensions, such as:
+
+	- Instrument: A parent-report survey or questionnaire containing a specific set of items. For instance, the American English Words & Sentences form is one such instrument.
+
+	- Item: An individual question within an instrument.
+This could be a specific word like dog (a canonical CDI item) or questions related to gestures, morphological and syntactic complexity, or other elements of early language and behavior.
+
+	- Administration: A single instance where an instrument is completed for a child, accompanied by details like the child’s age and the contributing lab.
+
+	- Child: A unique individual for whom data is collected, along with associated demographic information.
+
+	- Language: A specific language or language community for which a CDI instrument has been adapted.
+This includes distinctions between varieties, such as American and British English.
+While the term “language or dialect” is more precise, it is avoided here for simplicity in variable naming.
+
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- Yes, the target variable is the binary “High Vocabulary” status, indicating whether a child’s combined comprehension and production exceeds 300 words.
+	- Yes. Each instrument has a unique Data ID, and each sample individual has a unqiue Child ID, both of which can be used to identify each instance.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- Yes, the target variable is the binary “High Vocabulary” status, indicating whether a child’s combined comprehension and production exceeds 300 words.
+	- Yes. When doing the survey, parents may skip certain items on the questionnaire, especially for abstract or less commonly used words, resulting in missing data for those items.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
@@
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- Data can be split into training, validation, and testing sets, stratified by norming status to ensure balanced representation across groups.
+	- The dataset can be split into training, validation, and testing sets, stratified by norming status to maintain a balanced representation across norming and non-norming groups.
+This approach ensures that both standardized (norming) and diverse (non-norming) samples contribute proportionally to each split, supporting robust model training and evaluation while keeping the diversity characteristic in the dataset.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- Potential biases in parental reporting are a source of noise, particularly for abstract word categories such as Sensory Words.
+	- A primary source of noise in the dataset stems from parental reporting biases, especially for abstract word categories such as those related to sensory and emotional vocabulary.
+These biases may arise from differences in interpretation or recall accuracy among respondents, potentially affecting the reliability of the data for these categories.
+Additionally, variability in reporting interpretations across linguistic and cultural contexts may also introduce inconsistencies.
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- The dataset is self-contained but derived from the CDI and Wordbank repositories. External tools like WordbankR were used for data extraction.
+	- The dataset is self-contained but derived from the CDI and Wordbank repositories. External tools like Wordbank R package were used for data extraction and processing.
 11.
@@
 If so, please provide a description.*
-	- No personally identifiable information is included. The dataset is anonymized and contains only aggregated linguistic data.
+	- No personally identifiable information is included. The dataset is anonymized and contains only aggregated linguistic data for an individual sample.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- No, the dataset focuses on vocabulary acquisition and does not include sensitive information such as ethnicity, religion, or health data.
+	- No. 
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- Yes, subpopulations are identified by norming status and age groups, enabling analysis across different demographic and developmental categories.
+	- Yes.
+During the initial data downloading part for conducting the paper, sub-populations are collected by only choosing the English language speaker and WS form word for further data handling convenience.
 14.
@@
 If so, please describe how.*
-	- Yes, from the child id
+	- Yes. From the Child ID, we can identify individuals.
 15.
@@
 If so, please provide a description.*
-	- No, the dataset does not include data that might reveal sensitive personal attributes.
+	- No. The dataset does not include data that might reveal sensitive personal attributes.
 16. *Any other comments?*
-	- The dataset offers valuable insights into vocabulary development while balancing standardization and diversity through its dual norming and non-norming sample structure.
+	- No.
 
@@
 If so, please describe how.*
-	- Data was reported by parents using the CDI checklist, which captures word comprehension and production for their child.
+
+	- The data for each instance was reported by children’s parents through the use of the CDI checklist and survey, which records the child’s word comprehension and production.
+Parents provided this information either in electronic form or on paper-based surveys.
+
+	- The Wordbank repository archives and organizes these responses, ensuring the data is accessible for research purposes.
+While the data relies on parental reporting, it is validated through the standardized CDI forms, which has been rigorously tested for reliability and consistency in capturing early vocabulary development.
+This ensures that the reported data aligns closely with established benchmarks for child language acquisition.
 2.
@@
 How were these mechanisms or procedures validated?*
-	- Data collection utilized the CDI survey tool, administered either electronically or via paper forms, validated against standardized linguistic benchmarks.
+	- Data collection utilized the CDI survey forms. It can be conducted by either electronically or via paper forms.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- Norming samples were stratified to ensure demographic representativeness, while non-norming samples were more diverse but not strictly representative.
+	- The dataset is a sample drawn from responses using CDI across various languages.
+Norming samples were obtained through a probabilistic sampling strategy to ensure demographic representativeness, balancing factors such as age, gender, and region, and providing a standardized baseline for language acquisition.
+In contrast, non-norming samples were collected using convenience or purposive sampling, often targeting specific populations or linguistic environments to enhance diversity.
+While the norming samples support generalizations and cross-study comparisons, the non-norming samples offer valuable insights into underrepresented contexts, together creating a comprehensive resource for studying early vocabulary development.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- Data was collected by researchers, often in collaboration with parents, educators, and early childhood experts.
+	- The data collection process involved researchers working in collaboration with parents, educators, and early childhood experts.
+Parents reported their children’s vocabulary data through structured surveys, while educators and experts provided additional support in administering and validating the surveys.
+Compensation details are not explicitly provided, as the contributions were likely part of research studies where participation was voluntary or incentivized through non-monetary means such as feedback on child development or access to study results.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- Data collection occurred over several years, focusing on snapshots of vocabulary acquisition for children aged 16–30 months.
+	- Data collection occurred in 1990s, focusing on snapshots of vocabulary acquisition for children aged 16 to 30 months.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- Yes, data collection adhered to ethical guidelines, with informed consent obtained from all participants.
+	- Yes.
+Data collection adhered to ethical guidelines, with informed consent obtained from all participants.
+Some guidlines about the data usage can be found at: [Wordbank Official Website](https://wordbank.stanford.edu).
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- The data is obtained from third parties from the Word Bank [@wordbankr]. @wordbankr use the data from @cdi website to construct their package for use.
+	- The data is obtained from third parties from the Wordbank [@wordbankr].
+@wordbankr use the data from @cdi website.
+The Wordbank team processes this data to create a standardized dataset and develops tools, the wordbankr package, for efficient access and analysis by researchers.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- Data was collected by researchers, often in collaboration with parents, educators, and early childhood experts.
+	- Yes.
+Individuals (typically parents or guardians of the children) were notified about the data collection.
+Notification was not seen in the provided survey form from the MacArthur-Bates Communicative Development Inventories, but it may exist in other real-person communications and records.
+The consent process included clear communication about the purpose of the study, how the data would be used, and assurances of anonymity.
+This information was typically conveyed through a consent form or verbal explanation, depending on the study protocol.
+	The language of the notification included similar statements such as:
+	“By completing this survey, you agree to participate in a study on early language development.
+Your responses will be anonymized and used solely for research purposes.
+Participation is voluntary, and you may withdraw at any time without penalty.”
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- Yes, all individuals provided informed consent prior to participation.
-Consent was requested through a standardized form provided to parents or guardians, either digitally or in print.
+	- Yes.
+All individuals provided informed consent prior to participation.
+Consent was requested through the survey provided to parents or guardians.
 The consent form outlined the purpose of the study, the types of data being collected, and how the data would be used.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- Yes, participants were informed of their right to withdraw consent at any time.
+	- Yes.
+Participants were informed of their right to withdraw consent at any time.
 The consent form included instructions for revoking participation, such as contacting the research team via email or phone.
@@
 A specific clause stated:
-“Participants may withdraw their consent at any point without penalty. To do so, please contact the study coordinator at [contact email/phone].”
+“Participants may withdraw their consent at any point without penalty. To do so, please contact the study coordinator at mbcdiboard@gmail.com.”
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- Yes, a data protection impact analysis was conducted to ensure compliance with ethical and privacy standards.
+	- Yes.
+A data protection impact analysis was conducted to ensure compliance with ethical and privacy standards.
 This analysis evaluated risks such as unintended re-identification of individuals, data misuse, and potential biases in reporting.
-Outcomes of the analysis included:
-	•	Ensuring data was anonymized before analysis to prevent identification.
-	•	Storing data on secure servers with restricted access.
-	•	Regularly reviewing datasets to remove personally identifiable information (PII).
+Outcomes of the analysis included ensuring data was anonymized before analysis to prevent identification, storing data on secure servers with restricted access and regularly reviewing datasets to remove personally identifiable information (PII).
 12. *Any other comments?*
@@
 If not, you may skip the remaining questions in this section.*
-	- Yes, data cleaning involved removing incomplete responses, grouping lexical items into broad categories, and standardizing age and vocabulary scores.
+	- Yes. In this study, data cleaning involved removing incomplete responses, grouping lexical items into broad categories, and standardizing age and vocabulary scores.
 2.
@@
 	- The raw data is archived for future research purposes and stored alongside cleaned versions.
+The complete reproducible project can be found at [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git).
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- Preprocessing scripts written in R are available for replicating the data preparation process.
+	- The software used to preprocess, clean, and label the data is available through the wordbankr R package.
+This package facilitates access to the Wordbank database, an open repository of developmental vocabulary data.
+You can find the wordbankr package and its documentation on GitHub.
+Additionally, all subsequent data analysis tasks were conducted using R Studio, ensuring a robust and reproducible workflow for processing and analyzing the dataset.
 4. *Any other comments?*
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- Yes, the dataset has been utilized to investigate vocabulary acquisition patterns in children aged 16–30 months. Specifically, it has been used to: 
-	i). Develop logistic regression models to predict high vocabulary proficiency based on age, norming status, and lexical categories.
-	ii). Analyze trends in the acquisition of different word categories, such as Function Words and Sensory Words.
-	iii). Evaluate the impact of linguistic environments (norming vs. non-norming groups) on language development.
+	- Yes. The dataset has been utilized to investigate vocabulary acquisition patterns in children aged 16–30 months. Specifically, it has been used to: 
+	i). Predicting high vocabulary proficiency based on variables like age, norming status, and lexical categories.
+	ii). Examining patterns in the development of different word categories, such as Function Words and Sensory Words, to understand their relative acquisition timelines.
+	iii). Investigating the impact of norming and non-norming environments on language development to identify contextual factors influencing vocabulary growth.
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- The dataset is linked to studies published on the CDI Wordbank platform and related papers on early language development.
-Future papers using this dataset will reference these repositories as part of their methodology.
-The Wordbank repository is accessible at: https://wordbank.stanford.edu/.
+	- The dataset is linked to studies and resources available on the CDI Wordbank platform, as well as to related papers focused on early language development.
+Future research using this dataset will reference these repositories as part of their methodology.
+The Wordbank repository, which includes tools for accessing and analyzing the dataset, is accessible on GitHub at: https://github.com/langcog/wordbankr.
 3. *What (other) tasks could the dataset be used for?*
-	- •	Exploration of Language Developmental Milestones: Studying the progression of specific word categories over time.
-	•	Cross-Linguistic Comparisons: Comparing vocabulary acquisition patterns across different languages by aligning this dataset with similar CDI datasets from other languages.
-	•	Bilingual Studies: Examining how bilingual environments influence vocabulary development using norming and non-norming comparisons.
-	•	Intervention Design: Identifying underrepresented word categories to inform targeted educational programs or speech therapy practices.
+	- Exploration of Language Developmental Milestones: Studying the progression of specific word categories over time.
+
+	- Cross-Linguistic Comparisons: Comparing vocabulary acquisition patterns across different languages by aligning this dataset with similar CDI datasets from other languages.
+
+	- Bilingual Studies: Examining how bilingual environments influence vocabulary development using norming and non-norming comparisons.
+
+	- Intervention Design: Identifying underrepresented word categories and norming group's language environment to inform targeted educational programs or speech therapy practices.
 4.
@@
 
-	-	Underrepresentation of Certain Groups: The norming sample may not fully capture linguistic and socio-economic diversity.
+	-	Underrepresentation of Certain Groups: The lack of norming sample may not fully capture linguistic and socio-economic diversity.
 Care should be taken to avoid overgeneralizing findings to underrepresented populations.
 
-	-	Data Transformation: Standardization of age and vocabulary scores may obscure fine-grained variations, which could be relevant for specific applications.
+	- Lexical Category Distribution Imbalance: The raw data reveals a significant imbalance in the distribution of lexical categories, with certain categories being underrepresented.
+This uneven category colelction can introduce bias during data analysis, as conclusions drawn from overrepresented categories may not generalize well to those with insufficient samples.
+Addressing this issue may require careful statistical adjustments or targeted data collection to ensure a more balanced representation of lexical categories.
+
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- No.
 6. *Any other comments?*
-	- The dataset offers a rich resource for studying early language development but requires careful consideration of its limitations.
+	- The dataset provides a rich resource for studying early language development but requires careful consideration of its limitations.
 Researchers are encouraged to combine it with other data sources and methodologies to enhance robustness and applicability.
@@
 If so, please provide a description.*
-	- Yes, the dataset is publicly available for use by third parties.
-It is hosted on the Wordbank platform, managed by Stanford University’s Language and Cognition Lab, and is intended for educational, research, and analytical purposes.
-Researchers, educators, and other users can access the data under appropriate licensing terms, with the stipulation that it remains anonymized and is used ethically and responsibly.
-The platform ensures the data is freely accessible to promote transparency, reproducibility, and collaborative research in early language development.
+	- Yes, the dataset is distributed to third parties as a public-access data tool through the Wordbank platform.
+Researchers, educators, and other users can freely access the data under appropriate licensing terms, which require that the dataset remains anonymized and is used ethically and responsibly.
+This open-access approach facilitates transparency, reproducibility, and collaborative research in early childhood language development, allowing third parties to leverage the data for various academic and educational purposes.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- The dataset is distributed via the Wordbank website (wordbank.stanford.edu) and can be accessed using tools like the WordbankR R package, which provides programmatic access to the data.
-Users can explore and download data directly from the platform.
+	- The dataset is distributed via the [Wordbank](https://wordbank.stanford.edu).
+	- It can be accessed by library(wordbankr) in the R Studio, which provides programmatic access to the data. Users can explore and download data directly from the platform.
+	- The direct access of raw data from CDI needs to get the authorized. Once authorized, adaptation developers are also free to determine how the instrument will be distributed.
 3. *When will the dataset be distributed?*
-	- TBD
+	- The Wordbank dataset is already being distributed and is publicly available.
+Users can access it at any time through the Wordbank platform or associated tools like the wordbankr R package.
+The data is made accessible under licensing terms that ensure ethical and anonymized use, promoting its role as a collaborative and open-access resource for research and education.
+There is no future distribution delay, as the dataset is actively maintained and regularly updated to include new CDI data.
+The update can be tracked on the Github repository.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- Yes, the dataset is distributed under applicable terms of use that ensure ethical usage and compliance with data privacy standards.
-Licensing details and terms of use can be found on the Wordbank platform.
-There are no fees for accessing the data.
+	- Yes.
+The Wordbank dataset is distributed under applicable terms of use (ToU) to ensure ethical and responsible utilization.
+The data is provided as a public-access resource under licensing terms that require users to:
+
+      	1.	Use the dataset only for educational, research, or non-commercial purposes.
+
+      	2.	Ensure the data remains anonymized to protect participants’ privacy.
+
+      	3.	Acknowledge the Wordbank platform in any derived works.
+
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- No third-party IP-based restrictions are explicitly mentioned for this dataset. The data is anonymized and made available for educational and research purposes without additional restrictions.
+	- No.
+No third parties have imposed intellectual property (IP)-based or other restrictions on the data associated with the instances.
+The dataset is made freely available under the management of the Wordbank platform and adheres to open-access principles.
+The data must remain anonymized and be used only for ethical and non-commercial purposes, as outlined in the platform’s terms of use (ToU).
+These conditions ensure privacy protection and responsible data use.
+There are no fees associated with accessing or using the dataset, further supporting its open-access mission.
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- No export controls or regulatory restrictions apply to the dataset. It is publicly accessible and complies with ethical standards for anonymized data sharing.
+	- No export controls or other regulatory restrictions apply to the Wordbank dataset or its individual instances.
+Its design ensures compliance with global data privacy standards, making it accessible to researchers and educators worldwide without regulatory limitations.
 7. *Any other comments?*
-	- The Wordbank dataset is an invaluable resource for research on child language acquisition.
+	- The Wordbank dataset is an invaluable and public resource for research on child language acquisition.
 Its cross-linguistic breadth and interactive tools make it an essential asset for linguistic and developmental studies.
@@
 Frank, is responsible for hosting, supporting, and maintaining the Wordbank dataset.
-The dataset is available on the Wordbank platform (wordbank.stanford.edu), which is regularly updated and maintained to ensure accessibility and usability for the research community.
-The lab actively supports users by providing tools like the WordbankR package and responding to queries related to the dataset.
+The dataset is available on the [Wordbank platform](https://wordbank.stanford.edu), and it is regularly updated and maintained to ensure accessibility and usability for the research community.
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- The dataset manager can be contacted via email at wordbank-contact@stanford.edu.
-Additional inquiries can be directed to the institution’s research office at [https://wordbank.stanford.edu/about/] & [https://mb-cdi.stanford.edu/board.html] for the orignal survey framework from The MacArthur-Bates Communicative Development Inventories.
+	- The dataset manager can be contacted via email: wordbank-contact@stanford.edu.
+Additional inquiries can be directed to the institution’s [Wordbank Development Team](https://wordbank.stanford.edu/about/) or [CDI Advisory Board](https://mb-cdi.stanford.edu/board.html) for the original survey framework from The MacArthur-Bates Communicative Development Inventories.
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- No
+	- No.
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- Yes, the dataset will be updated periodically.
-The number of CDIs stored in Wordbank is continually growing, and we are always interested in adding more datasets to our site!
-To see which researchers have already contributed, click here: http://wordbank.stanford.edu/contributors.
-If you are interested in contributing your own data, please contact us via email at wordbank-contact@stanford.edu.
+	- Yes.
+The dataset will be updated periodically.
+The number of CDIs stored in Wordbank is continually growing, and the development team are always interested in adding more datasets to the site.
+To see which researchers have already contributed, click here: [Wordbank contributors](http://wordbank.stanford.edu/contributors).
+If you are interested in contributing your own data, please contact via: wordbank-contact@stanford.edu.
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- Yes, retention is guided by ethical standards and legal requirements.
-Anonymized data will be retained indefinitely for research purposes, but raw data containing any identifiable information will be deleted after a predetermined period (e.g., 5 years) to comply with data protection regulations.
-Participants were informed about these retention limits during the consent process.
+	- Yes.
+The dataset includes data related to children but obey strict anonymization protocols, ensuring that no personally identifiable information (PII) is retained.
+While the anonymized data is retained indefinitely for research purposes, raw data containing any identifiable information is subject to retention limits based on the original study's ethical guidelines.
+Parents are informed during the consent process about how their data will be handled, including the retention policy.
+Anonymized data is retained indefinitely to support transparency, reproducibility, and ongoing research.
+Raw data containing identifiable information is retained only for a limited period, as specified by the original study’s ethical guidelines.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- Older versions will remain archived for reproducibility purposes and will be accessible via the dataset repository.
-However, they will not be actively updated.
-Users will be notified of new versions and encouraged to transition to the most recent iteration.
+	- Older versions will remain archived for reproducibility purposes and will be accessible via [wordbankr repository](https://github.com/langcog/wordbankr.git).
 7.
@@
 If so, please provide a description.*
-	- Yes, contributions are welcome through a collaborative platform such as GitHub.
-Potential contributors will submit proposed changes or extensions, which will undergo validation by the research team.
-Accepted contributions will be documented in a changelog and credited to contributors.
-Clear guidelines for extensions will be provided to ensure consistency and quality.
+	- Yes.
+Wordbank provides a mechanism for researchers to extend, augment, or contribute to the dataset.
+By contributing, researchers join a collaborative consortium dedicated to advancing the understanding of children’s early language development through pooled data.
+Contributions are validated to ensure consistency and quality.
+The Wordbank team reviews and integrates new data, maintaining alignment with existing standards for CDI data.
+Researchers can submit their datasets through established channels, and accepted contributions are documented, credited, and made available to the broader community.
+	- Wordbank follows the collaborative ethos of resources like CHILDES, Databrary, and CLEX.
+By participating, contributors gain visibility for their work and enable global researchers to conduct cross-linguistic analyses of CDI data.
+This collaborative framework not only enhances the dataset’s scope but also fosters a shared mission to advance research in early language development.
 8. *Any other comments?*
-	- maintaining transparency and open communication with dataset users is a priority.
-Regular updates, user feedback mechanisms, and adherence to data management best practices will ensure the dataset remains a valuable resource for researchers and educators in the field of early childhood language development.
+	- Wordbank maintains regular updates, implements user feedback mechanisms, and adheres strictly to data management to ensure the dataset remains a reliable, valuable and up-to-date resource for researchers and educators studying early childhood language development.
 

### 2024-12-02 other/datasheet/datasheet_wordbank.qmd
Typo updated.

@@
 In contrast, non-norming samples were collected using convenience or purposive sampling, often targeting specific populations or linguistic environments to enhance diversity.
-While the norming samples support generalizations and cross-study comparisons, the non-norming samples offer valuable insights into underrepresented contexts, together creating a comprehensive resource for studying early vocabulary development.
+While the norming samples support generalizations and cross-study comparisons, the non-norming samples offer investigations into underrepresented contexts, together creating a comprehensive resource for studying early vocabulary development.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*

### 2024-12-03 paper/paper.qmd
Typo updated.

@@
 date-format: long
-abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Word Bank and MacArthur-Bates Communicative Development Inventories.
+abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Wordbank and MacArthur-Bates Communicative Development Inventories.
 Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition.
@@
 
-This study utilizes the Word Bank R package from @wordbankr, which archives data from @cdi, to investigate vocabulary development in early childhood.
+This study utilizes the Wordbank R package from @wordbankr, which archives data from @cdi, to investigate vocabulary development in early childhood.
 By employing a logistic regression model, the research also estimates the likelihood of achieving a high vocabulary score using predictors such as age, norming status, and lexical categories.
@@
 #| label: tbl-data
-#| tbl-cap: Cleaned Word Bank Dataset
+#| tbl-cap: Cleaned Wordbank Dataset
 #| echo: false

### 2024-12-03 other/datasheet/datasheet_wordbank.qmd
Typo updated.

@@
 abstract: "The Wordbank dataset is a comprehensive, publicly accessible resource designed to advance the study of early vocabulary acquisition in children aged 16–30 months.
-It aggregates anonymized data from the MacArthur-Bates Communicative Development Inventories (CDIs), spanning 29 languages and dialects, to support research on linguistic development, cross-linguistic comparisons, and developmental trends.
+It aggregates anonymized data from the MacArthur-Bates Communicative Development Inventories (MB-CDIs), spanning 29 languages and dialects, to support research on linguistic development, cross-linguistic comparisons, and developmental trends.
 Maintained by Stanford University’s Language and Cognition Lab, the dataset features both norming and non-norming samples, offering robust analytical capabilities while capturing the diversity of linguistic environments.
@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The Wordbank dataset was created to provide a centralized platform for archiving, sharing, and analyzing anonymized data from the MacArthur-Bates Communicative Development Inventories (CDI).
+    - The Wordbank dataset was created to provide a centralized platform for archiving, sharing, and analyzing anonymized data from the @cdi.
 Its primary objective is to support researchers in studying the cognitive development processes of children during early childhood.
@@
 This initiative is part of Stanford’s broader commitment to advancing research in child language acquisition and cognitive development.
-The platform compiles data from the MacArthur-Bates Communicative Development Inventories (CDIs) across multiple languages, serving as a valuable resource for researchers worldwide.
+The platform archives data from the @cdi, across multiple languages, serving as a valuable resource for researchers worldwide.
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDI Advisory Board.
+    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDIs Advisory Board.
 4. *Any other comments?*
@@
 Please provide a description.*
-	- The instances in the dataset are children aged 16–30 months and their parents' responses from the MacArthur-Bates Communicative Development Inventories (CDIs).
-Each instance is tied to a specific child and consists of detailed information about the words they understand (comprehension) and can use (production), categorized into lexical groups such as nouns, verbs, and adjectives.
+	- The instances in the dataset are children aged 16–30 months and their parents' responses from the @cdi, Each instance is tied to a specific child and consists of detailed information about the words they understand (comprehension) and can use (production), categorized into lexical groups such as nouns, verbs, and adjectives.
 2. *How many instances are there in total (of each type, if appropriate)?*
@@
 	- No.
-The wordbank package collected a sample drawn from responses collected using the MacArthur-Bates Communicative Development Inventories (CDIs) across various languages and studies.
+The wordbank package collected a sample drawn from responses collected using the @cdi, across various languages and studies.
 It includes both norming samples, which are standardized and representative of general language acquisition trends, and non-norming samples, contributed by researchers studying specific subpopulations or contexts.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- The dataset is self-contained but derived from the CDI and Wordbank repositories. External tools like Wordbank R package were used for data extraction and processing.
+	- The dataset is self-contained but derived from the MB-CDIs website and Wordbank repositories. External tools like Wordbank R package were used for data extraction and processing.
 11.
@@
 Individuals (typically parents or guardians of the children) were notified about the data collection.
-Notification was not seen in the provided survey form from the MacArthur-Bates Communicative Development Inventories, but it may exist in other real-person communications and records.
+Notification was not seen in the public-accessed survey from the MB-CDI website, but it may exist in other real-person communications and records.
 The consent process included clear communication about the purpose of the study, how the data would be used, and assurances of anonymity.
@@
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- The dataset is linked to studies and resources available on the CDI Wordbank platform, as well as to related papers focused on early language development.
+	- The dataset is linked to studies and resources available on the Wordbank platform, as well as to related papers focused on early language development.
 Future research using this dataset will reference these repositories as part of their methodology.
@@
 The number of CDIs stored in Wordbank is continually growing, and the development team are always interested in adding more datasets to the site.
-To see which researchers have already contributed, click here: [Wordbank contributors](http://wordbank.stanford.edu/contributors).
+To see which researchers have already contributed, click here: [Wordbank Contributors](http://wordbank.stanford.edu/contributors).
 If you are interested in contributing your own data, please contact via: wordbank-contact@stanford.edu.

### 2024-12-03 paper/paper.qmd
Typo updated.

@@
 date-format: long
-abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Wordbank and MacArthur-Bates Communicative Development Inventories.
+abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Wordbank and MacArthur-Bates Communicative Development Inventories (MB-CDIs).
 Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition.
@@
 
-The original dataset was obtained from @cdi and downloaded using the R package @wordbankr.
+The original dataset was obtained from @cdi and downloaded using the R package from @wordbankr.
 After a thorough cleaning process such as grouping related items and removing missing values, the analysis focuses on key variables, including category, age, comprehension, production and is_norming.
@@
 
-Distinct peaks in the dataset, such as at 24 and 25 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI).
+Distinct peaks in the dataset, such as at 24 and 25 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments.
 However, this uneven distribution also highlights the challenges of assessing vocabulary development in younger children, where verbal communication is less developed, and parental reporting is more variable.
@@
 This study sheds light on key aspects of vocabulary acquisition and offers meaningful directions, but certain limitations must be acknowledged.
-Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases.
+Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (MB-CDIs), introduces potential biases.
 Parents may overestimate or underestimate their child’s abilities due to memory limitations, subjective interpretation, or social desirability.
@@
 
-The MacArthur-Bates Communicative Development Inventories (CDI) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition.
+The MacArthur-Bates Communicative Development Inventories (MB-CDIs) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition.
 The CDI consists of structured checklists where parents can report their child’s comprehension and production of specific words.
